### 2019-09-17 code/Mayo_RegressDiagnosis_4_TWAS.Rmd

---
title: Covariate and differential expression analysis of MAYO reprocessed counts
output: html_notebook
---
Date of analysis update: "`r date()`"

| *Synapse ID* | *File Name* |
|  -----------------------------------------  |   ---------                      |
| syn5223705.5 | | 
| syn8698211.1 | | 
| syn3817650.8 | | 
| syn8690904.1 | | 
| syn6126119.3 | | 
| syn8698214.1 | | 
| syn8449369.2 | | 
| syn6126114.3 | | 
| syn6145639.5 | | 
| syn8690799.1 | |

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, cache=FALSE}
##It is assumed your working directory is "~/ampad-DiffExp/gene_level_analysis"
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(plyr)
library(tidyverse)
library(psych)
library(limma)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(cqn)
library(glmnet)
library(synapser)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
library(knitr)
library(doParallel)
library(foreach)
library(sva)
cl = makeCluster(detectCores()-2)
registerDoParallel(cl)
options(xtable.type="html")
# Source modified utility functions from limma and sva package
source('./utilityFunctions/parallelDuplicateCorrelation.R')
source('./utilityFunctions/irwsva.build.R')
source('./utilityFunctions/f.pvalue.R')
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synLogin}
synLogin()
```

### Data download
Obtain count matrix and metadata from Synapse
```{r download.data}
# Get reprocessed sample counts for temporal cortex
COUNT_TC_ID <- 'syn8690799';
ALL_USED_IDs <- COUNT_TC_ID
COUNT_TC <- synGet(COUNT_TC_ID, version = 1)$path %>%
  read.table(header=T, sep='\t', check.names = F, row.names = 1)
# Get reprocessed sample counts for cerebellum
COUNT_CE_ID <- 'syn8690904';
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_CE_ID
COUNT_CE <- synGet(COUNT_CE_ID, version = 1)$path %>%
  read.table(header=T, sep='\t', check.names = F, row.names = 1)
# Combine TCX and CBE counts
COUNT = cbind(COUNT_TC, COUNT_CE[rownames(COUNT_TC),])
# Convert rownames of counts from tracking id to ensemble gene id
tmp = data.frame(Gene.ID = rownames(COUNT)) %>%
  dplyr::mutate(ID = Gene.ID) %>%
  tidyr::separate(ID, c('ensembl_gene_id', 'position'), sep = '\\.')
rownames(tmp) = tmp$Gene.ID
rownames(COUNT) = tmp[rownames(COUNT), 'ensembl_gene_id']
# Get clinical metadata
METADATA_TC_ID <- 'syn3817650'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TC_ID
METADATA_TC <- synGet(METADATA_TC_ID, version = 8)$path %>%
  read.table(sep=',',header=T, stringsAsFactors=F) %>%
  tidyr::separate(ID, c('Donor_ID', 'Tissue'), sep = '_') %>%
  dplyr::mutate(ID = paste(Donor_ID, Tissue, sep = '_'))
# Get clinical metadata
METADATA_CE_ID <- 'syn5223705'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_CE_ID
METADATA_CE <- synGet(METADATA_CE_ID, version = 5)$path %>%
  read.table(sep=',',header=T, stringsAsFactors=F) %>%
  tidyr::separate(SampleID, c('Donor_ID', 'Tissue')) %>%
  dplyr::mutate(ID = paste(Donor_ID, Tissue, sep = '_')) %>%
  dplyr::rename(Gender = Sex, FLOWCELL = Flowcell)
# Get picard metrics from synapse
METADATA_PICARD_CE_ID <- 'syn8698214';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_PICARD_CE_ID
METADATA_PICARD_CE <- synGet(METADATA_PICARD_CE_ID, version = 1)$path %>%
  read.table(sep='\t', header=T, stringsAsFactors=F) %>% 
  dplyr::rename(ID = sample)
# colnames(METADATA_PICARD_CE) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA_PICARD_CE))
# colnames(METADATA_PICARD_CE) = gsub('RnaSeqMetrics__','',colnames(METADATA_PICARD_CE))
METADATA_PICARD_TC_ID <- 'syn8698211';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_PICARD_TC_ID
METADATA_PICARD_TC <- synGet(METADATA_PICARD_TC_ID, version = 1)$path %>%
  read.table(sep='\t', header=T, stringsAsFactors=F) %>% 
  dplyr::rename(ID = sample)
# colnames(METADATA_PICARD_TC) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA_PICARD_TC))
# colnames(METADATA_PICARD_TC) = gsub('RnaSeqMetrics__','',colnames(METADATA_PICARD_TC))
# Merge all metadata
METADATA = list(METADATA_TC, METADATA_CE) %>%
  data.table::rbindlist(use.names = T, fill = T) %>%
  dplyr::inner_join(list(METADATA_PICARD_TC, METADATA_PICARD_CE) %>% 
                      data.table::rbindlist(use.names = T, fill = T)) %>%
  as.data.frame
colnames(METADATA) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA))
colnames(METADATA) = gsub('RnaSeqMetrics__','',colnames(METADATA))
# Get gene specific parameters from synapse
GENE.PARAM = synGet('syn8449369', version = 2)$path %>% 
  read.table(sep = '\t', header = T, stringsAsFactors = F)
ALL_USED_IDs = c(ALL_USED_IDs, 'syn8449369')
GENE.LEN = dplyr::select(GENE.PARAM, ensembl_gene_id, gene.length) %>% unique()
rownames(GENE.LEN) = GENE.LEN$ensembl_gene_id
GENE.GC = dplyr::select(GENE.PARAM, ensembl_gene_id, percentage_gc_content) %>% unique()
rownames(GENE.GC) = GENE.GC$ensembl_gene_id
# Samples excluded by MAYO investigators
SAMPLES.EXCLUDE = c(read.table(synGet('syn6126119', version = 3)$path, sep = '\t', header = T, stringsAsFactors = F)$Sample.Name,
                    read.table(synGet('syn6126114', version = 3)$path, sep = '\t', header = T, stringsAsFactors = F)$Sample.Name)
ALL_USED_IDs = c(ALL_USED_IDs, 'syn6126119', 'syn6126114')
```
Following `r length(SAMPLES.EXCLUDE)` samples were removed from analysis as original investigators tagged them as outliers for various reasons as described here syn6126119, syn6126114
Samples removed are `r paste(SAMPLES.EXCLUDE, collapse = ', ')`

### Data preprocessing
```{r preprocess.data, echo=TRUE}
# Filter metadata
METADATA = as_tibble(METADATA, .name_repair = "unique") %>%
  dplyr::filter(ID %in% colnames(COUNT)) %>%
  dplyr::filter(!is.na(RIN)) %>%
  dplyr::filter(!is.na(PCT_INTRONIC_BASES)) %>%
  dplyr::filter(!is.na(AgeAtDeath)) %>%
  dplyr::filter(!is.na(Source)) %>%
  dplyr::filter(!(ID %in% SAMPLES.EXCLUDE)) %>%
  as.data.frame()
# Fix Tissue
METADATA$Tissue = gsub('CER', 'CBE', METADATA$Tissue)
# Fix AgeAtDeath
METADATA = METADATA %>%
  dplyr::mutate(AgeAtDeath = gsub("_or_above", "", AgeAtDeath))
# Fix Diagnosis
METADATA = METADATA %>%
  dplyr::mutate(Source.Diagnosis = Diagnosis) %>%
  dplyr::mutate(Source.Diagnosis = factor(Source.Diagnosis,
                                          levels = c('AD', 'Control', 'Pathologic Aging', 'PSP'),
                                          labels = c('AD', 'CONTROL', 'PATH_AGE', 'PSP'))) %>%
  dplyr::mutate(Diagnosis = gsub("Pathologic Aging", "OTHER", Diagnosis),
                Diagnosis = gsub("PSP", "OTHER", Diagnosis),
                Diagnosis = factor(Diagnosis, 
                                   levels = c("AD", "Control", "OTHER"),
                                   labels = c('AD', 'CONTROL', 'OTHER'))) %>%
  dplyr::mutate(Tissue.Diagnosis = paste(Tissue, Diagnosis, sep = '.'),
                Tissue.SourceDiagnosis = paste(Tissue, Source.Diagnosis, sep = '.'))
# Fix Gender
METADATA = METADATA %>%
  dplyr::mutate(Sex = Gender,
                Sex = factor(Sex, levels = c('F','M'), labels = c('FEMALE','MALE')))
# Get RIN square
METADATA = METADATA %>%
  dplyr::mutate(RIN2 = RIN^2)
# Add apoe4 genotype (0, 1, 2)
METADATA$APOE4 = 0
METADATA$APOE4[METADATA$ApoE %in% c(24, 34)] = 1
METADATA$APOE4[METADATA$ApoE %in% c(44)] = 2
METADATA$Tissue.APOE4 = paste(METADATA$Tissue, METADATA$APOE4, sep = '.')
# Match covariates to expression data
indToRetain = intersect(METADATA$ID, colnames(COUNT))
indRemoved = setdiff(colnames(COUNT), METADATA$ID)
COUNT <- COUNT[, indToRetain]
rownames(METADATA) = METADATA$ID
METADATA = METADATA[indToRetain,]
```
`r dim(COUNT)[2]` samples from `r length(unique(METADATA$Donor_ID))` subjects were obtained from the Mayo cohorts in AMP-AD reprocessed rnaseq project.

Following samples were removed `r indRemoved`

### Predict missing PMI
```{r predict.pmi}
# Get PMI
PMI = METADATA[,c('PMI', 'ID', 'Donor_ID'), drop = F]
PMI$PMI = PMI$PMI
# Get expression (convert counts to cpm)
expr = voom(COUNT, design = NULL)$E
ind = which(rowSums(expr>=1)/dim(expr)[2] >= 0.5)
expr = expr[ind,]
# Combine expression from different regions of brain of an individual into one
expr = plyr::dlply(PMI, .(Donor_ID), .fun = function(x, expr){
  data.frame(value = rowMeans(expr[,x$ID,drop = F], na.rm = T)) %>%
    plyr::rename(c('value' = unique(x$Donor_ID)))
}, expr, .parallel = F) %>%
  do.call(cbind,.)
# Split data into training and prediction set
PMI = dplyr::select(PMI, Donor_ID, PMI) %>% unique()
rownames(PMI) = PMI$Donor_ID
PMI = PMI[colnames(expr),]
# Fit a logistic regression model
x.train = expr[,!is.na(PMI$PMI)]
x.predict = expr[,is.na(PMI$PMI)]
y.train = PMI[!is.na(PMI$PMI),]
y.train = y.train[colnames(x.train), 'PMI']
plr.model = cv.glmnet(t(x.train), y.train, family = 'gaussian')
coeff = as.matrix(coef.glmnet(plr.model, s = 'lambda.min'))
ssres = sum((predict(plr.model, t(x.train), s = 'lambda.min', type = 'response') - y.train)^2)
sstot = sum((y.train-mean(y.train, na.rm = T))^2, na.rm = T)
R2.allgenes = 1-(ssres/sstot)
y.predict.all = predict(plr.model, t(x.predict), s = 'lambda.min', type = 'response') %>%
  rownameToFirstColumn('Donor_ID') %>%
  plyr::rename(c('1' = 'PMI'))
# Get genes that are predictors of PMI and use them as control probes
CONTROL.PROBES_ID = 'syn6145639'
ALL_USED_IDs = c(ALL_USED_IDs, CONTROL.PROBES_ID)
CONTROL.PROBES = fread(synGet(CONTROL.PROBES_ID, version = 5)$path, data.table = F) %>%
  dplyr::mutate(ID = x) %>%
  tidyr::separate(ID, c('ensembl_gene_id', 'position'), sep = '\\.')
# Initial normalisation of gene expression
ind = rownames(expr) %in% CONTROL.PROBES$ensembl_gene_id 
expr = expr[ind,]
# Fit a logistic regression model
x.train = expr[,!is.na(PMI$PMI)]
x.predict = expr[,is.na(PMI$PMI)]
y.train = PMI[!is.na(PMI$PMI),]
y.train = y.train[colnames(x.train), 'PMI']
plr.model = cv.glmnet(t(x.train), y.train, family = 'gaussian')
coeff = as.matrix(coef.glmnet(plr.model, s = 'lambda.min'))
ssres = sum((predict(plr.model, t(x.train), s = 'lambda.min', type = 'response') - y.train)^2)
sstot = sum((y.train-mean(y.train, na.rm = T))^2, na.rm = T)
R2.smallgenes = 1-(ssres/sstot)
y.predict = predict(plr.model, t(x.predict), s = 'lambda.min', type = 'response') %>%
  rownameToFirstColumn('Donor_ID') %>%
  plyr::rename(c('1' = 'PMI'))
PMI = PMI %>% 
  dplyr::select(Donor_ID, PMI) %>%
  dplyr::filter(!is.na(PMI)) %>%
  list(y.predict) %>%
  data.table::rbindlist(use.names = T, fill = T)
PMI$PMI[PMI$PMI < 0] = 0
METADATA = METADATA %>% 
  dplyr::select(-PMI) %>%
  dplyr::left_join(PMI)
```
Fitted R2 values are, (i) with all gx features `r round(R2.allgenes, 3)`, (ii) with small set of inferred gx features `r round(R2.smallgenes,3)`

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering}
primaryVariable <- c("Tissue.SourceDiagnosis", "Tissue.Diagnosis", "Tissue.APOE4")
#_# ADD DIAGNOSIS HERE TO REGRESS FOR TWAS
FactorCovariates <- c("Donor_ID", "Source", "Sex", "FLOWCELL", "Tissue.SourceDiagnosis", "Tissue.Diagnosis", "Tissue.APOE4", "Diagnosis")
ContCovariates <- c("RIN", "RIN2", "AgeAtDeath","PMI", "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES",
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", "PCT_RIBOSOMAL_BASES")
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$ID
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=6}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm'))
```
### Explore metadata
```{r data.explore, fig.width = 12, fig.height = 8}
my.theme = theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = RIN)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme
# AgeAtDeath
p[[2]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(AgeAtDeath))) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeAtDeath') + my.theme
# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PMI))) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + my.theme
# Ribosomal bases
p[[4]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_RIBOSOMAL_BASES))) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Fraction of Ribosomal Bases') + my.theme
# Intronic bases
p[[5]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_INTRONIC_BASES))) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Fraction Intronic Bases') + my.theme
# Intergenic bases
p[[6]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_INTERGENIC_BASES))) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Intergenic Bases') + my.theme
multiplot(plotlist = p, cols = 3)
```

### Filter genes
* Remove genes that have less than 1 cpm counts in at least 50% of samples per Tissue x Diagnosis
* Remove genes with missing gene length and percentage GC content
```{r filter.genes}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('SampleID') %>%
  dlply(.(Tissue.Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
  }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$ensembl_gene_id[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$gene.length)]) %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
# Check gene biotype
## Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "dec2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
## Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "gene_biotype"),
                       filters = "ensembl_gene_id", 
                       values = PROCESSED_COUNTS$filteredExprMatrix$genes,
                       mart = mart)
summary(factor(Ensemble2HGNC$gene_biotype)) %>%
  rownameToFirstColumn('Biotype') %>%
  dplyr::rename(fraction = DF) %>%
  dplyr::mutate(fraction = fraction/dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]) %>%
  dplyr::filter(fraction >= 0.01) %>%
  kable
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

### Library Normalisation
Library normalisation is performed using cqn (conditional quantile normalisation)
```{r cqn}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gc_content'],
                          lengths = GENE.LEN[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

### Outlier Analysis
#### Sample outliers
Outlier analysis is performed before library normalisation with raw cpm counts
```{r outlier.analysis, cache = FALSE, fig.width = 10}
indToRemove = c("11311_CER", "1923_CER", "1923_TCX", "11396_TCX", "11294_TCX", "11408_TCX",
                '1950_TCX', '1950_CER', '1925_TCX', '1957_CER')
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(label = SampleID) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="top") 
p <- p + facet_grid(Tissue~.+Diagnosis, scales = 'free')
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove) %>%
             droplevels(),
           aes(x = logCPM)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Tissue~.+Diagnosis)
p
indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]
tmp = COVARIATES %>%
  dplyr::group_by(Tissue.Diagnosis) %>%
  dplyr::summarise(count = n()) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.') %>%
  tidyr::spread(Diagnosis, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

Distribution of samples are: `r kable(tmp)`

#### Gene outliers
Assign NA values to genes that are above and below 3 std deviation of its distribution
```{r winsorise.data}
# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(CQN.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)

  x[x < (mn-3*std.dev)] = NA
  x[x > (mn+3*std.dev)] = NA
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
CQN.GENE_EXPRESSION$E.no.na = CQN.GENE_EXPRESSION$E
CQN.GENE_EXPRESSION$E.no.na[is.na(CQN.GENE_EXPRESSION$E.no.na)] = 0
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Sample clustering
PCA based clustering of samples
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.no.na, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="top") + facet_grid(Tissue+.~Diagnosis, scales = 'free_y')
p
```

Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
COVARIATES.tmp = (COVARIATES[,c(FactorCovariates), drop = F])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E.no.na)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE, fig.height=10, fig.width=20}
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
p = ggplot(tmp1, aes(x = logCPM)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(Tissue+.~Diagnosis, scale = 'free')
p
```

Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(CQN.GENE_EXPRESSION$E.no.na))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E.no.na, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix.

NOTE:
1. Using a mixed effect model where random effect is chosen as Donor_ID
2. Adding Source, FLOWCELL and Gender a priori to variable selection
3. Primary variable of interest Tissue.Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Source', 'FLOWCELL', 'Sex');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, 
                         c(postAdjustCovars, primaryVariable, 'Donor_ID'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
notEstimable = c()
loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects and Donor_ID as random effect'))

  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

  # Estimate voom weights for dispersion control
  cnts = NEW.COUNTS; cnts[is.na(cnts)] = 0
  VOOM.GENE_EXPRESSION = voom(cnts, design=DM1$design, 
                              plot=F, na.rm = T)

  # Calculate correlation values of random effects
  VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$Donor_ID,
                                             method = 'lmer')
  # correlation = list(); correlation$cor = 0.3819258

  if (!is.nan(correlation$cor)){
    # Re-estimate voom weights
    VOOM.GENE_EXPRESSION = voom(cnts, 
                                design=DM1$design, plot=F, 
                                block = COVARIATES$Donor_ID, 
                                correlation = correlation$cor)

    # Fit linear model using new weights and new design
    VOOM.ADJUSTED.FIT = lmFit(CQN.GENE_EXPRESSION$E,
                              design = DM1$design,
                              weights = VOOM.GENE_EXPRESSION$weights,
                              block = COVARIATES$Donor_ID, 
                              correlation = correlation$cor)

    # Residuals after normalisation
    RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                  CQN.GENE_EXPRESSION$E)
  } else {
    notEstimable = c(notEstimable, postAdjustCovars[length(postAdjustCovars)])
    postAdjustCovars = postAdjustCovars[1:(length(postAdjustCovars)-1)]
  }

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates),
                         c(postAdjustCovars, primaryVariable, 'Donor_ID', notEstimable))

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  expr = RESIDUAL.GENE_EXPRESSION; expr[is.na(expr)] = 0
  residualSigCovars = runPCAandPlotCorrelations(expr, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  residCovars = setdiff(residualSigCovars$significantCovars, c(postAdjustCovars, primaryVariable, notEstimable))
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}
modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')
tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`

### Sanity check
```{r residual.adj, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(expr, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(expr))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
PCA of residual data
```{r decompse.normalise.data2.1, fig.height=10, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)
# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Region','Diagnosis'), sep = '\\.') %>% 
  dplyr::mutate(Region = factor(Region),
                Source = factor(Source))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Region~., scales = 'free_y')
p
```
Tree based clustering of residual data
```{r decompse.normalise.data2.2, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,FactorCovariates])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Adjust data with covariates for Network Analysis
Identified covariates are regressed out from the expression matrix for network analysis
```{r network.adjust}
# Get design matrix
DESIGN.NET = getDesignMatrix(COVARIATES[, postAdjustCovars, drop = F], Intercept = F)
DESIGN.NET = DESIGN.NET$design[,linColumnFinder(DESIGN.NET$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.NET.WEIGHTS = voom(cnts, design=DESIGN.NET, plot=F)
# Re-estimate voom weights
VOOM.NET.WEIGHTS = voom(cnts, 
                        design = DESIGN.NET, plot=F,
                        block = COVARIATES$Donor_ID, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.NET.FIT = lmFit(VOOM.NET.WEIGHTS$E,
                     weights = VOOM.NET.WEIGHTS$weights,
                     design = VOOM.NET.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)
# Residuals after normalisation
RESIDUAL.NET.GENE_EXPRESSION = residuals.MArrayLM(VOOM.NET.FIT,
                                                  CQN.GENE_EXPRESSION$E)
```

### SVA Adjustments for eQTL analysis
Conditioned on primary variable (Tissue.Diagnosis) and identified covariates estimate surrogate variables using SVA package
```{r sva.adjust}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep(primaryVariable[2],colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]
# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(FIT, CQN.GENE_EXPRESSION$E)
# Compute actual variance of all principal components
expr = RESIDUAL.SVA.GENE_EXPRESSION
expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)
# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])
# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.2)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p
# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$E.no.na,
                        MODEL1, MODEL0, n.sv = NUM.SV, 
                        B = 30, tol = 1e-20)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)
# Estimate voom weights for dispersion control
VOOM.SVA.WEIGHTS = voom(cnts, design=cbind(MODEL1, SURR.VAR), plot=F)
# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Donor_ID, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.SVA.FIT = lmFit(VOOM.SVA.WEIGHTS$E,
                     weights = VOOM.SVA.WEIGHTS$weights,
                     design = VOOM.SVA.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)
# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$E)
# Add variable of interest back to the residuals
varsToAddIn = grep("Tissue.Diagnosis", colnames(MODEL1), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```

### Save Environment and SVA.GENE_EXPRESSION
```{r SAVE and Wrap}
save.image("MAYO.RData")
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = "MAYO_SVA_Diagnosis_Corrected.tsv", row.names = T, col.names = T, sep ="\t", quote=F)

```