### 2019-09-17 code/ROSMAP_RegressDiagnosis_4_TWAS.Rmd
Full Store Implemented

---
title: Covariate and differential expression analysis of ROSMAP reprocessed counts Using Diagnosis as a covariate for TWAS
author: "JKG - Adapted from: Thanneer Perumal"
editor_options:
  chunk_output_type: console
---
Date of analysis update: "`r date()`"

syn3382527.7
syn4300313.1
syn8449369.2
syn7116000.1
syn8691134.1
syn3191087.3
syn8698240.2

### Load Libraries
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(plyr)
library(tidyverse)
library(psych)
library(limma)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(cqn)
library(synapser)
library(knitr)
library(doParallel)
library(foreach)
cl = makeCluster(detectCores()-2)
registerDoParallel(cl)
options(xtable.type="html")
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synLogin}
synLogin()
```

### Data download
Obtain count matrix and metadata from synapse
```{r download.data}
# Download expression data
COUNT_ID <- 'syn8691134';
ALL_USED_IDs <- COUNT_ID
COUNT_OBJ <- synGet(COUNT_ID, version = 1)
COUNT <- read.table(COUNT_OBJ$path, header=T, sep='\t', check.names = F, row.names = 1)
COUNT[,grep('150_120419', colnames(COUNT))[2]] = NULL
# Convert rownames of counts from tracking id to ensemble gene id
tmp = data.frame(Gene.ID = rownames(COUNT)) %>%
  dplyr::mutate(ID = Gene.ID) %>%
  tidyr::separate(ID, c('ensembl_gene_id', 'position'), sep = '\\.')
rownames(tmp) = tmp$Gene.ID
rownames(COUNT) = tmp[rownames(COUNT), 'ensembl_gene_id']
# Get clinical metadata
METADATA.CLINICAL_ID <- 'syn3191087'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.CLINICAL_ID
METADATA.CLINICAL_OBJ <- synGet(METADATA.CLINICAL_ID, version = 3)
METADATA.CLINICAL <- read.table(METADATA.CLINICAL_OBJ$path,sep=',',header=T)
# Get clinical metadata with uncensored ages
METADATA.CLINICAL_ID1 <- 'syn7116000'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.CLINICAL_ID1
METADATA.CLINICAL_OBJ1 <- synGet(METADATA.CLINICAL_ID1, version = 1)
METADATA.CLINICAL1 <- read.table(METADATA.CLINICAL_OBJ1$path,sep=',',header=T)
# Get technical covariates
METADATA.TECH_ID <- 'syn4300313'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.TECH_ID
METADATA.TECH_OBJ <- synGet(METADATA.TECH_ID, version = 1)
METADATA.TECH <- read.table(METADATA.TECH_OBJ$path,sep='\t',header=T)
# Get picard metrics from synapse
METADATA.PICARD_ID <- 'syn8698240';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_ID
METADATA.PICARD <- synGet(METADATA.PICARD_ID, version = 2)$path %>%
  data.table::fread() %>%
  dplyr::rename(Sampleid = sample)
# Fix error in technical covariates data
KEY_ID <- 'syn3382527'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = KEY_ID  
KEY <- synGet(KEY_ID, version = 7)$path %>%
  read.csv %>% 
  dplyr::filter(!is.na(rnaseq_id)) %>%
  dplyr::select(projid, rnaseq_id) %>%
  tidyr::separate(rnaseq_id, c('a','b','batch'), sep = '_') %>% 
  unite(Sampleid, a, b) %>%
  dplyr::select(-batch) %>%
  unique
# Match technical and clinical covariates
METADATA <- METADATA.TECH %>%
  dplyr::left_join(METADATA.PICARD) %>%
  dplyr::select(-projid) %>%
  dplyr::left_join(KEY) %>%
  dplyr::left_join(METADATA.CLINICAL) %>%
  dplyr::select(-age_first_ad_dx, -age_death, -age_at_visit_max) %>%
  dplyr::left_join(METADATA.CLINICAL1)
# Pick higher quality RIN batch for sample 492_120515
METADATA <- METADATA %>%
  dplyr::group_by(Sampleid) %>%
  dplyr::top_n(1, RINcontinuous)
colnames(METADATA) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA))
colnames(METADATA) = gsub('RnaSeqMetrics__','',colnames(METADATA))
# Get gene specific parameters from synapse
GENE.PARAM = synGet('syn8449369', version = 2)$path %>%
  data.table::fread(data.table = FALSE)
ALL_USED_IDs = c(ALL_USED_IDs, 'syn8449369')
GENE.LEN = dplyr::select(GENE.PARAM, ensembl_gene_id, gene.length) %>% 
  unique() 
rownames(GENE.LEN) = GENE.LEN$ensembl_gene_id
GENE.GC = dplyr::select(GENE.PARAM, ensembl_gene_id, percentage_gc_content) %>% 
  unique() 
rownames(GENE.GC) = GENE.GC$ensembl_gene_id 
```

### Data preprocessing
```{r preprocess.data, cache=TRUE}
# Remove samples with no cogdx, RIN, PMI scores and age_death
METADATA <- METADATA %>%
  ungroup %>%
  dplyr::filter(Sampleid %in% colnames(COUNT)) %>%
  dplyr::filter(!is.na(cogdx), !is.na(braaksc), !is.na(ceradsc)) %>%
  dplyr::filter(!is.na(RINcontinuous)) %>%
  dplyr::filter(!is.na(pmi)) %>%
  #dplyr::filter(!is.na(PCT_INTRONIC_BASES)) %>%
  dplyr::filter(!is.na(age_death)) %>%
  as.data.frame()
# Add harmonised case-control status
METADATA$Diagnosis = 'OTHER'
METADATA$Diagnosis[METADATA$cogdx == 1 & METADATA$braaksc <= 3 & METADATA$ceradsc >= 3] = 'CONTROL'
METADATA$Diagnosis[METADATA$cogdx == 4 & METADATA$braaksc >= 4 & METADATA$ceradsc <= 2] = 'AD'
# Add sex variable 
METADATA$Sex = 'FEMALE'
METADATA$Sex[METADATA$msex == 1] = 'MALE'
# Add apoe4 genotype (0, 1, 2)
METADATA$APOE4 = 0
METADATA$APOE4[METADATA$apoe_genotype %in% c(24, 34)] = 1
METADATA$APOE4[METADATA$apoe_genotype %in% c(44)] = 2
# METADATA$APOE4[is.na(METADATA$apoe_genotype)] = NA
# Get square of RIN
METADATA$RINcontinuous2 = METADATA$RINcontinuous^2
# Match covariates to expression data
indToRetain = intersect(METADATA$Sampleid, colnames(COUNT))
removedIDs = setdiff(colnames(COUNT), METADATA$Sampleid)
COUNT = COUNT[,indToRetain]
rownames(METADATA) = METADATA$Sampleid
METADATA = METADATA[indToRetain,]
```
Dorsolateral prefrontal cortex of `r dim(COUNT)[2]` subjects from the ROS and MAP cohorts are used for the analysis. Following sample are removed due to missing metadata `r paste(removedIDs, collapse = ',')`

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering, cache=TRUE}
primaryVariable <- c("cogdx", "Diagnosis", "APOE4")
FactorCovariates <- c("Batch", "Sex", "race", "spanish", "cogdx", "Diagnosis", "APOE4")
ContCovariates <- c("RINcontinuous", "RINcontinuous2", "age_death", "pmi", "educ", 
                    "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES",
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", 
                    "PCT_RIBOSOMAL_BASES")
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES <- data.frame(lapply(COVARIATES,function(x){x <- sapply(x,function(y){str_replace_all(as.character(y),'\\+','')})}))
rownames(COVARIATES) <- METADATA$Sampleid
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```

Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=9, fig.height=6}
COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm'))
```

### Explore metadata
```{r data.explore, fig.width = 10, fig.height = 14}
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Diagnosis, y = RINcontinuous)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')
# AgeAtDeath
p[[2]] = ggplot(COVARIATES, aes(x = Diagnosis, y = age_death)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + theme(legend.position = 'top')
# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Diagnosis, y = pmi)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + theme(legend.position = 'top')
# Education
p[[4]] = ggplot(COVARIATES, aes(x = Diagnosis, y = educ)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Education') + theme(legend.position = 'top')
# Intronic bases
p[[5]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_INTRONIC_BASES)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Fraction Intronic Bases') + theme(legend.position = 'top')
# Ribosomal bases
p[[6]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_RIBOSOMAL_BASES)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Ribosomal Bases') + theme(legend.position = 'top')
multiplot(plotlist = p, cols = 2)
```

### Filter genes
* Remove genes that have less than 1 cpm counts in at least 50% of samples per Diagnosis
* Remove genes with missing gene length and percentage GC content
```{r filter.genes}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
  }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$ensembl_gene_id[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$gene.length)]) %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
# Check gene biotype
## Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "dec2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
## Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "gene_biotype"),
                       filters = "ensembl_gene_id", 
                       values = PROCESSED_COUNTS$filteredExprMatrix$genes,
                       mart = mart)
summary(factor(Ensemble2HGNC$gene_biotype)) %>%
  rownameToFirstColumn('Biotype') %>%
  dplyr::rename(fraction = DF) %>%
  dplyr::mutate(fraction = fraction/dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]) %>%
  dplyr::filter(fraction >= 0.01) %>%
  kable
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

### Library Normalisation
Library normalisation is performed using cqn (conditional quantile normalisation)
```{r cqn}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gc_content'],
                          lengths = GENE.LEN[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

### Outlier Analysis
#### Sample outliers
Outlier analysis is performed before library normalisation with raw cpm counts
```{r outlier.analysis, cache = FALSE, fig.width = 10}
indToRemove = c('380_120503', '500_120515')
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- dplyr::left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(label = SampleID)
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Batch, shape=Diagnosis, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  dplyr::left_join(COVARIATES %>%
                     rownameToFirstColumn('SampleID'))
p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top')
p
indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]
tmp = COVARIATES %>%
  dplyr::group_by(Diagnosis, cogdx) %>%
  dplyr::summarise(count = n()) %>%
  tidyr::spread(Diagnosis, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

Distribution of samples are: `r kable(tmp)`

#### Gene outliers
Assign NA values to genes that are above and below 3 std deviation of its distribution
```{r winsorise.data}
# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(CQN.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)

  x[x < (mn-3*std.dev)] = NA
  x[x > (mn+3*std.dev)] = NA
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
CQN.GENE_EXPRESSION$E.no.na = CQN.GENE_EXPRESSION$E
CQN.GENE_EXPRESSION$E.no.na[is.na(CQN.GENE_EXPRESSION$E.no.na)] = 0
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Sample clustering
PCA based clustering of samples
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.no.na, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Batch, shape=cogdx, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```

Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Batch", "Sex", "Diagnosis")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E.no.na)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE, fig.height=10, fig.width=20}
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID'))
p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(.~Diagnosis, scale = 'free')
p
```

Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(CQN.GENE_EXPRESSION$E.no.na))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E.no.na, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix
1. Adding Batch and Sex a priori to variable selection
2. Primary variable of interest Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Batch', 'Sex');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, c(postAdjustCovars, primaryVariable))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
#_# ADD DIAGNOSIS HERE TO REGRESS FOR TWAS
postAdjustCovars = c(postAdjustCovars, "Diagnosis", names(which.max(covariatesEffects))) %>% unique()
loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))

  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

  # Estimate voom weights for dispersion control
  cnts = NEW.COUNTS
  cnts[is.na(cnts)] = 0
  VOOM.GENE_EXPRESSION = voom(cnts, 
                              design=DM1$design, 
                              plot=F,
                              na.rm = T)

  # Fit linear model using new weights and new design
  VOOM.ADJUSTED.FIT = lmFit(CQN.GENE_EXPRESSION$E,
                            design = DM1$design,
                            weights = VOOM.GENE_EXPRESSION$weights)

  # Residuals after normalisation
  RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                CQN.GENE_EXPRESSION$E)

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates), c(postAdjustCovars, primaryVariable))

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  expr = RESIDUAL.GENE_EXPRESSION
  expr[is.na(expr)] = 0
  residualSigCovars = runPCAandPlotCorrelations(expr, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  residCovars = setdiff(residualSigCovars$significantCovars, c(postAdjustCovars, primaryVariable))
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}
modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')
tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`

### Sanity check
```{r residual.adj, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(expr, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(expr))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
PCA of residual data
```{r decompse.normalise.data2.1, fig.height=6, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)
# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2)) 
p <- p + geom_point(aes(color=Batch, shape=Diagnosis, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```
Tree based clustering of residual data
```{r decompse.normalise.data2.2, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c('Batch', 'Sex', primaryVariable)])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Adjust data with covariates for Network Analysis
Identified covariates are regressed out from the expression matrix for network analysis
```{r network.adjust}
# Get design matrix
DESIGN.NET = getDesignMatrix(COVARIATES[, postAdjustCovars, drop = F], Intercept = F)
DESIGN.NET = DESIGN.NET$design[,linColumnFinder(DESIGN.NET$design)$indepCols]
# Estimate voom weights for dispersion control
cnts = NEW.COUNTS
cnts[is.na(cnts)] = 0
VOOM.NET.WEIGHTS = voom(cnts, design=DESIGN.NET, plot=F)
# Fit linear model using new weights and new design
VOOM.NET.FIT = lmFit(CQN.GENE_EXPRESSION$E,
                     design = DESIGN.NET,
                     weights = VOOM.NET.WEIGHTS$weights)
# Residuals after normalisation
RESIDUAL.NET.GENE_EXPRESSION = residuals.MArrayLM(VOOM.NET.FIT,
                                                  CQN.GENE_EXPRESSION$E)
```

### SVA Adjustments for eQTL analysis
Conditioned on primary variable (Diagnosis) and identified covariates estimate surrogate variables using SVA package
```{r sva.adjust}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
cnts = NEW.COUNTS
cnts[is.na(cnts)] = 0
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep(primaryVariable[2],colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]
# Calculate the number of surrogate variables (be method)
# NUM.SV = sva::num.sv(CQN.GENE_EXPRESSION$E, MODEL1, method = 'be', seed = 123456, B = 30)
# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(FIT, CQN.GENE_EXPRESSION$E)
# Compute actual variance of all principal components
expr = RESIDUAL.SVA.GENE_EXPRESSION
expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)
# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])
# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.3)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p
# Estimate surrogate variables
SURR.VAR = sva::sva(CQN.GENE_EXPRESSION$E.no.na, 
                    MODEL1, MODEL0, n.sv = NUM.SV, B = 30)$sv
SURR.VAR = data.frame(SURR.VAR)
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)
# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F)
# Fit linear model using new weights and new design
VOOM.SVA.FIT = lmFit(CQN.GENE_EXPRESSION$E, 
                     design = cbind(MODEL1, SURR.VAR),
                     weights = VOOM.SVA.WEIGHTS$weights)
# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$E)
# Add variable of interest back to the residuals
varsToAddIn = grep("Diagnosis", colnames(MODEL1), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```

### Save Environment and SVA.GENE_EXPRESSION
```{r SAVE and Wrap, eval = F}
save.image("ROSMAP.RData")
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = "ROSMAP_SVA_Diagnosis_Corrected.tsv", row.names = T, col.names = T, sep ="\t", quote=F)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn20798011';
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with GRCh37 with CQN normalisation (DLPFC)';
thisFileName <- 'ROSMAP_RegressDiagnosis_4_TWAS.Rmd'
# Github link
thisRepo <- getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with GRCh37 with CQN normalisation (DLPFC)';

# Code WTF is this??
CODE <- Folder(name = "ROSMAP", parentId = parentId)
CODE <- synStore(CODE)

#Set Used SynIDs For Provenance
Syns_Used <- c("syn3382527", "syn4300313", "syn8449369", "syn7116000",
  "syn8691134", "syn3191087", "syn8698240"
)

# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  tissueTypeAbrv	= 'DLPFC', 
  study = 'ROSMAP', 
  organism = 'HomoSapiens',
  consortium	= 'ROSMAP',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh37'
)

# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'ROSMAP_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ = File('ROSMAP_SVA_Diagnosis_Corrected.tsv', 
                  name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'residualGeneExpForeQTL'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'ROSMAP_SVA_Diagnosis_Covariates.tsv', row.names = F, sep = '\t', quote=F)
ENRICH_OBJ = File('ROSMAP_SVA_Diagnosis_Covariates.tsv', 
                  name = 'Covariates', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'covariates'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_Counts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_Counts.tsv', 
                  name = 'Counts (filtered raw)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'filteredCounts'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store Estimated counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_Counts.tsv', 
                  name = 'Counts (estimated)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'estimatedCounts'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_logCPM.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_logCPM.tsv', 
                  name = 'Counts (filtered logCPM)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'filteredLCPM'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_offset.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_offset.tsv', 
                  name = 'Gene length and GC content offset', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'offset'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store design matrix
MODEL1 %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_Design.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_Design.tsv', 
                  name = 'Design Matrix', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'designMatrix'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store surrogate variables
SURR.VAR %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_SurVar.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_SurVar.tsv', 
                  name = 'Surrogate Variables', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'surrogateVariables'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Diagnosis_Regressed_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ = File('ROSMAP_DLPFC_Diagnosis_Regressed_netResidualExpression.tsv', 
                  name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'residualGeneExpForNetAnlz'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

stopCluster(cl)
```

### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')
synapseLogin()
knit2synapse::knitfile2synapse(file = "./ROSMAP_RegressDiagnosis_4_TWAS.Rmd",
                                 owner = "syn20798014")
```