### 2019-09-25 code/Ancestry_PCA.Rmd
Ancestry Update

---
title: "Process Genotypes For Ancestry Analysis"
author: "JKG"
date: "9/18/2019"
output: pdf_document
  chunk_output_type: console
---

```{r libs, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  results='hide')
library(reticulate)
```
##Before running 
Run the following in docker exec -it <CONTAINER> /bin/bash 
chmod 777 /root
mv TWAS/ /home/<USR>/TWAS/
chown -hR <GID>:<USR> /home/<USR>/TWAS/

## Login to Synapse Will need to replace with your credentials
```{bash Login, echo=T, results='hide', eval=FALSE}
source /root/.bashrc
synapse login -u <USR> -p <PSWD> --rememberMe

#Alternativly you can setup a credentials file as such:
touch ~/.synapseConfig
echo "[authentication]" >> ~/.synapseConfig
echo "username = <USR>" >> ~/.synapseConfig
echo "password = <PASWD>" >> ~/.synapseConfig
```
## Pull 1000Genomes down and process
* Pull out reference populations (CEU, YRI, CHB, JPT)
```{bash pull_1000Genomes, echo=T, results='hide', cache = FALSE, echo=FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
#PATH=$PATH:TWAS/bin/
CORES=`expr $(nproc) - 2`

#ID Sample list of needed Ref Populations
wget ftp://ftp-trace.ncbi.nih.gov/1000genomes/ftp/phase1/analysis_results/integrated_call_sets/integrated_call_samples.20101123.ALL.panel

ftp://ftp-trace.ncbi.nih.gov/1000genomes/ftp/phase1/analysis_results/integrated_call_sets/

grep CEU integrated_call_samples.20101123.ALL.panel | cut -f1 > CEU.samples.list
grep YRI integrated_call_samples.20101123.ALL.panel | cut -f1 > YRI.samples.list
grep CHB integrated_call_samples.20101123.ALL.panel | cut -f1 > CHB.samples.list
grep JPT integrated_call_samples.20101123.ALL.panel | cut -f1 > JPT.samples.list

#Make Plink IDV List of all 1000G Ref Samps to pull
cat CEU.samples.list YRI.samples.list CHB.samples.list JPT.samples.list > 1000G_IDV.list
paste 1000G_IDV.list 1000G_IDV.list > foo.txt
mv foo.txt 1000G_IDV.list

##Pull The LDREF Panel SNPS
#Pull LDREF Data
wget https://data.broadinstitute.org/alkesgroup/FUSION/LDREF.tar.bz2
tar xjvf LDREF.tar.bz2

#Merge  LDREF Data
for i in {2..22} 
do 
    echo LDREF/1000G.EUR.$i>> MergedChrs.txt
done 
plink --bfile LDREF/1000G.EUR.1 --threads $CORES --silent --merge-list MergedChrs.txt --make-bed --allow-no-sex --out TotalLDRef

#Write SNP List for LDREF Data
plink --threads $CORES --bfile TotalLDRef --silent --write-snplist --out LDREF_SNPs

#Reformat SNP List for plink filter.txt
awk '{ print $1"\t"$4"\t"$4"\t"$2 }' TotalLDRef.bim > LDREF_SNPs.txt

rm -r LDREF/

#Pull 1000G Data, convert to plink, Filt for INDV and SNPs
for i in {1..22}
do
	wget ftp://ftp-trace.ncbi.nih.gov/1000genomes/ftp/phase1/analysis_results/integrated_call_sets/ALL.chr$i\.integrated_phase1_v3.20101123.snps_indels_svs.genotypes.vcf.gz

    plink --vcf ALL.chr$i\.integrated_phase1_v3.20101123.snps_indels_svs.genotypes.vcf.gz --snps-only --extract LDREF_SNPs.txt --keep 1000G_IDV.list --silent --threads $CORES --make-bed --out 1000Genomes_chr$i
  rm ALL.chr$i\.integrated_phase1_v3.20101123.snps_indels_svs.genotypes.vcf.gz
done

#Merge All Plink Chr Files 
for i in {2..22}
do
  echo 1000Genomes_chr$i >> Merge.list
done

plink --bfile 1000Genomes_chr1 --merge-list Merge.list --silent --make-bed --out 1000Genomes_Merged

#Store 1000G Populations
mkdir 1000G_RefSamps
mv CEU.samples.list 1000G_RefSamps/
mv YRI.samples.list 1000G_RefSamps/
mv JPT.samples.list 1000G_RefSamps/
mv CHB.samples.list 1000G_RefSamps/

#Clean Working Directory
rm 1000Genomes_chr*
rm *.list
rm *.panel
rm LDREF.tar.bz2
rm MergedChrs.txt
```

## Process MSBB WGS Data
* Pull VCFs from Synapse
* Convert to plink
* Merge across chromosomes
* Filter for 1000G SNPs
```{bash ProcessMSBB, echo=T, results='hide', cache = FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`

##Pull VCF Files from Synapse
mkdir VCF
synapse get -r syn11707204 --downloadLocation VCF/

##Convert VCFs to plink
mkdir GWAS_Plink

rm VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_X*
rm VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_Y*
rm VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_others*
rm -r VCF/WGS_JointCalling/
rm VCF/SYNAPSE_METADATA_MANIFEST.tsv

#Convert to PLINK
for i in {22..1}
do
  plink --vcf VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_$i\.recalibrated_variants.MSSM.vcf.gz --threads $CORES --make-bed --silent --allow-no-sex --snps-only --out GWAS_Plink/MSSB_Chr_$i
  rm VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_$i\.recalibrated_variants.MSSM.vcf.*
done
rm -r VCF/

#Replace the "." IDs with a uniq ID
for i in {22..1}
do
  awk '{ print $1"\t"$1"_"$4"\t"$3"\t"$4"\t"$5"\t"$6 }' GWAS_Plink/MSSB_Chr_$i\.bim > foo.txt
  mv foo.txt GWAS_Plink/MSSB_Chr_$i\.bim
done

#Make an LD_Ref SNPList with matching name Schema
awk '{print $1"\t"$2"\t"$3"\t"$1"_"$2 }' LDREF_SNPs.txt > MSBB_LD_SNPS.txt

##Merge Chromsomes in Plink Format 
for i in {2..22} 
do 
    echo GWAS_Plink/MSSB_Chr_$i >> MergedGWAS.txt
done

plink --threads $CORES --bfile GWAS_Plink/MSSB_Chr_1 --silent --merge-list MergedGWAS.txt --extract MSBB_LD_SNPS.txt --make-bed --allow-no-sex --out GWASTotal
#rm -r GWAS_Plink/
rm MergedGWAS.txt

```

### ReName the MSBB SNPS so the can be combined
```{python ReNamer, echo=T, results='hide'}
import sys
import os
import re

InFile = 'TotalLDRef.bim'
InFileb = 'GWASTotal.bim' #LDREF_FiltSNPs.bim

RS_IDs = {}
RS_Pos = {}

F1 = open(InFile)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	RS_IDs[ ''.join([lst[0], '_', lst[3]]) ] = lst[1]
	RS_Pos[ ''.join([lst[0], '_', lst[3]]) ] = lst[2]
F1.close()

#Open Output
OUT = open( ''.join([ InFileb.replace(".bim", ""), 'Renamed.bim']), 'w' )

F1 = open(InFileb)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	Name = RS_IDs[ ''.join([ lst[0], "_", lst[3] ])]
	Post = RS_Pos[ ''.join([ lst[0], "_", lst[3] ])]

	Entry = '\t'.join([ lst[0], Name, Post, lst[3], lst[4], lst[5] ])
	print >>OUT, Entry

F1.close()
OUT.close()
```

### Finalize the MSBB LD SNP-Set Derived from WGS
```{bash FinalizeMSBB, echo=T, results='hide', cache = FALSE  }
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`

#Check the SNP-Set
##These numbers should be equal:
grep 'rs' GWASTotalRenamed.bim | wc -l
cat GWASTotalRenamed.bim | wc -l

##This Number should be zero:
Repod=$(awk '{print $2}' GWASTotalRenamed.bim | grep "\." - | wc -l)
if [ "$Repod" -eq "0" ]; then
   mv GWASTotalRenamed.bim GWASTotal.bim;
fi
if [ "$Repod" -ne "0" ]; then
   echo $Repod" lines could've failed Translation"
fi
plink --bfile GWASTotal --silent --threads $CORES --make-bed --out MSBB_Preliminary_Genos
rm GWASTotal.*
```

## Process MAYO Genotyping Data
```{bash MAYO, echo=T, results='hide', cache = FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`

#Download
mkdir Mayo
synapse get -r syn8650955 --downloadLocation Mayo/

plink --bfile Mayo/MayoRNAseq_RNAseq_Genome-Wide_Genotypes_HRCimputed --threads $CORES --extract LDREF_SNPs.txt --range --make-bed --out Mayo_LDREF_Filt

rm -r Mayo/
#SNPS REMAINING COMPARED TO TOTAL LDREF SET:
##1,187,249 out of 1,190,321 (  99.74% )
```

### ReName the Mayo SNPS so the can be combined
```{python ReNamerMayo, echo=T, results='hide'}
import sys
import os
import re

InFile = 'TotalLDRef.bim'
InFileb = 'Mayo_LDREF_Filt.bim'

RS_IDs = {}
RS_Pos = {}

F1 = open(InFile)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	RS_IDs[ ''.join([lst[0], '_', lst[3]]) ] = lst[1]
	RS_Pos[ ''.join([lst[0], '_', lst[3]]) ] = lst[2]
F1.close()

#Open Output
OUT = open( ''.join([ InFileb.replace(".bim", ""), 'Renamed.bim']), 'w' )

F1 = open(InFileb)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	Name = RS_IDs[ ''.join([ lst[0], "_", lst[3] ])]
	Post = RS_Pos[ ''.join([ lst[0], "_", lst[3] ])]

	Entry = '\t'.join([ lst[0], Name, Post, lst[3], lst[4], lst[5] ])
	print >>OUT, Entry

F1.close()
OUT.close()
```

### Finalize the Mayo LD SNP-Set Derived from imputed genotyping
```{bash CheckMayo, echo=T, results='hide', cache = FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`

#Check the SNP-Set
##These numbers should be equal:
grep 'rs' Mayo_LDREF_FiltRenamed.bim | wc -l
cat Mayo_LDREF_Filt.bim | wc -l

##This Number should be zero:
Repod=$(awk '{print $2}' Mayo_LDREF_FiltRenamed.bim | grep "\." - | wc -l)
if [ "$Repod" -eq "0" ]; then
   mv Mayo_LDREF_FiltRenamed.bim Mayo_LDREF_Filt.bim;
fi
if [ "$Repod" -ne "0" ]; then
   echo $Repod" lines could've failed Translation"
fi
plink --bfile Mayo_LDREF_Filt --silent --threads $CORES --make-bed --out Mayo_Preliminary_Genos
rm Mayo_LDREF_Filt.*
```

## Process ROSMAP Genotyping Data
```{bash DosageToHardCalls, echo=T, results='hide', cache = FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`

#Synapse gets:
declare -A SynArr
arr["1"]=syn5879161
arr["2"]=syn5879456
arr["3"]=syn5879461
arr["4"]=syn5879463
arr["5"]=syn5879470
arr["6"]=syn5879472
arr["7"]=syn5879473
arr["8"]=syn5879559
arr["9"]=syn5879628
arr["10"]=syn5879677
arr["11"]=syn5879748
arr["12"]=syn5879792
arr["13"]=syn5879812
arr["14"]=syn5879827
arr["15"]=syn5879828
arr["16"]=syn5879830
arr["17"]=syn5879832
arr["18"]=syn5879834
arr["19"]=syn5879835
arr["20"]=syn5879836
arr["21"]=syn5879837
arr["22"]=syn5879838

#Pull FAM File
synapse get syn20809809

#Loop Through chromosomes and Process
for i in {1..22}
do

	synapse get ${arr[$i]}

	chr=$i
	echo "1 AMP-AD_ROSMAP_Rush-Broad_AffymetrixGenechip6_Imputed_chr"$chr".dosage.gz" > b.txt

  #Write a Dosage File
	plink --threads $CORES --fam Updated_Actual_Fam.fam --silent --dosage b.txt list noheader format=1 --write-dosage

  #Convert to plink hard calls 
	plink2 --threads $CORES --fam Updated_Actual_Fam.fam --silent --import-dosage plink.out.dosage format=1 --hard-call-threshold 0.1 --make-bed --out AMP-AD_ROSMAP_Rush-Broad_AffymetrixGenechip6_Imputed_chr$chr

	#Clean out dosage to save disk space!
	rm plink.out.dosage
  rm AMP-AD_ROSMAP_Rush-Broad_AffymetrixGenechip6_Imputed_chr$i\.dosage.gz
done
rm b.txt
##Merge Chromsomes in Plink Format 
echo AMP-AD_ROSMAP_Rush-Broad_AffymetrixGenechip6_Imputed_chr2 > MergedGWAS.txt
for i in {3..22} 
do 
    echo AMP-AD_ROSMAP_Rush-Broad_AffymetrixGenechip6_Imputed_chr$i >> MergedGWAS.txt
done

plink --threads $CORES --bfile AMP-AD_ROSMAP_Rush-Broad_AffymetrixGenechip6_Imputed_chr1 --silent --merge-list MergedGWAS.txt  --make-bed --allow-no-sex --out ROSMAP_Temp

rm -r AMP-AD_ROSMAP_Rush-Broad_AffymetrixGenechip6_Imputed_chr*
rm MergedGWAS.txt

grep 'rs' ROSMAP_Temp.bim | awk '{ print $2 }' - | sort - > RSIDS.txt
awk '{ print $4 }' LDREF_SNPs.txt | sort - > RefIDs.txt

comm -12 RefIDs.txt RSIDS.txt > RosMapKeeps.txt

plink --threads $CORES --bfile ROSMAP_Temp --extract RosMapKeeps.txt --silent --make-bed --allow-no-sex --out ROSMAP_temp
rm ROSMAP_Temp.*
rm RSIDS.txt
rm RefIDs.txt
#1705 people pass mind
#1068685 of 1187582 snps pass --geno 0.1
```

### ReName the Mayo SNPS so the can be combined
```{python ReNamerRosMap, echo=T, results='hide'}
import sys
import os
import re

InFile = 'TotalLDRef.bim'
InFileb = 'ROSMAP_temp.bim'

RS_Chr = {}
RS_Pos = {}
RS_BP ={}

F1 = open(InFile)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	RS_Chr[ ''.join([lst[1], '_', lst[4], '_', lst[5]]) ] = lst[0]
	RS_Pos[ ''.join([lst[1], '_', lst[4], '_', lst[5]]) ] = lst[2]
	RS_BP[ ''.join([lst[1], '_', lst[4], '_', lst[5]]) ] = lst[3]
F1.close()

#Open Output
OUT = open( ''.join([ InFileb.replace(".bim", ""), 'Renamed.bim']), 'w' )
OUT2 = open( 'FailedSNPS.txt', 'w' )

F1 = open(InFileb)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')
	LOC = ''.join([ lst[1], '_', lst[4], '_', lst[5] ])
	chrm = RS_Chr.get( LOC, None )
	pos = RS_Pos.get( LOC, None )
	bp = RS_BP.get( LOC, None )

	if( chrm == None ):
	  print >>OUT, LINE
	  print >>OUT2, lst[1]
	else:
	  Entry = '\t'.join([ chrm, lst[1], pos, bp, lst[4], lst[5] ])
	  print >>OUT, Entry

F1.close()
OUT.close()
OUT2.close()
```

##Merge the Genotype Sets and Preform PCA
```{bash MergeAndCall, cache = FALSE, echo=T, results='hide'}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`

mv ROSMAP_tempRenamed.bim ROSMAP_temp.bim
plink --threads $CORES --bfile ROSMAP_temp --silent --exclude FailedSNPS.txt --make-bed --allow-no-sex --out ROSMAP_Temp
#rm ROSMAP_temp.*

#Check ROSMAP SNP-Set
##These numbers should be equal:
grep 'rs' ROSMAP_Temp.bim | wc -l
wc -l ROSMAP_Temp.bim

##This Number should be zero:
Repod=$(awk '{print $2}' ROSMAP_Temp.bim | grep "\." - | wc -l)

if [ "$Repod" -eq "0" ]; then
  plink --threads $CORES --bfile ROSMAP_Temp --silent --make-bed --allow-no-sex --out ROSMAP_Preliminary_Genos
  rm ROSMAP_Temp.*
fi
if [ "$Repod" -ne "0" ]; then
   echo $Repod" lines could've failed Translation"
fi

#Overly Cautious ID of common SNPs as the 1000G Data set is much smaller
##Write & sort SNP List
plink --threads $CORES --silent --bfile ROSMAP_Preliminary_Genos --write-snplist --out RosMap
plink --threads $CORES --silent --bfile MSBB_Preliminary_Genos --write-snplist --out MSBB
plink --threads $CORES --silent --bfile Mayo_Preliminary_Genos --write-snplist --out Mayo
plink --threads $CORES --silent --bfile 1000Genomes_Merged --write-snplist --out 1000G

cat MSBB.snplist | sort > MSBB.srt.snplist
cat RosMap.snplist | sort > RosMap.srt.snplist
cat Mayo.snplist | sort > Mayo.srt.snplist
cat 1000G.snplist | sort > 1000G.srt.snplist

#ID Common SNPS in all datasets
comm -12 MSBB.srt.snplist 1000G.srt.snplist | comm -12 Mayo.srt.snplist - | comm -12 RosMap.srt.snplist - > All_CommonSNPs.snplist
wc -l All_CommonSNPs.snplist

##ID A/T C/G SNPS for removal (Not sure if I need to do this for the TWAS as well..)
#Find and remove A/T C/G SNPS
grep -P "A\tT$" 1000Genomes_Merged.bim | awk '{ print $2 }' - > remove.snplist
grep -P "T\tA$" 1000Genomes_Merged.bim | awk '{ print $2 }' - >> remove.snplist
grep -P "G\tC$" 1000Genomes_Merged.bim | awk '{ print $2 }' - >> remove.snplist
grep -P "C\tG$" 1000Genomes_Merged.bim | awk '{ print $2 }' - >> remove.snplist
wc -l remove.snplist

echo "ROSMAP_Preliminary_Genos" > Merge.list
echo "MSBB_Preliminary_Genos" >> Merge.list
echo "Mayo_Preliminary_Genos" >> Merge.list

plink --threads $CORES --bfile 1000Genomes_Merged --silent --merge-list Merge.list --allow-no-sex --exclude remove.snplist --extract All_CommonSNPs.snplist --make-bed --out Merged_GenoTypes

cat Merged_GenoTypes-merge.missnp remove.snplist > Total_Remove.snplist
wc -l Total_Remove.snplist

##Try removing from all sets independently....
plink --threads $CORES --bfile 1000Genomes_Merged --silent --exclude Total_Remove.snplist --extract All_CommonSNPs.snplist --make-bed --out 1000_4_Merge
plink --threads $CORES --bfile ROSMAP_Preliminary_Genos --silent --exclude Total_Remove.snplist --extract All_CommonSNPs.snplist --make-bed --out ROSMAP_4_Merge
plink --threads $CORES --bfile MSBB_Preliminary_Genos --silent --exclude Total_Remove.snplist --extract All_CommonSNPs.snplist --make-bed --out MSBB_4_Merge
plink --threads $CORES --bfile Mayo_Preliminary_Genos --silent --exclude Total_Remove.snplist --extract All_CommonSNPs.snplist --make-bed --out Mayo_4_Merge

echo "ROSMAP_4_Merge" > Merge.list
echo "MSBB_4_Merge" >> Merge.list
echo "Mayo_4_Merge" >> Merge.list

plink --threads $CORES --bfile 1000_4_Merge --silent --merge-list Merge.list --allow-no-sex --maf 0.01 --mind 0.1 --geno 0.1 --exclude Total_Remove.snplist --extract All_CommonSNPs.snplist --make-bed --out Merged_GenoTypes
rm *_Preliminary_Genos.*

#PCA
plink --bfile Merged_GenoTypes --silent --threads $CORES --maf 0.01 --mind 0.1 --max-maf 0.4 --indep-pairwise 50 5 0.2 --pca var-wts --out Total_PCA
```

## Prep Input for Cluster Based on Ancestry
```{bash CLustPrep, cache = FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`
wget http://www.compgen.pitt.edu/GemTools/GemTools.R

plink --bfile Merged_GenoTypes --threads $CORES --geno 0.0000001 --maf 0.01 --mind 0.1 --hwe 0.00001 --max-maf 0.4  --indep-pairwise 50 5 0.2 --silent --recode12 --compound-genotypes --out AllSnps4GemClust

#Re-Code in 0-1-2 Allele State
cut -d" " -f7- AllSnps4GemClust.ped > peds.txt
perl -pe 's/([^\s]+)\s+([^\s]+)/$1$2/g' peds.txt > PEDS.txt
cut -d" " -f 1 AllSnps4GemClust.ped > FID.txt
cut -d" " -f 2 AllSnps4GemClust.ped > IID.txt
paste -d " " FID.txt IID.txt > Names.txt

sed -i 's/22/2/g' PEDS.txt
sed -i 's/00/3/g' PEDS.txt
sed -i 's/12/1/g' PEDS.txt
sed -i 's/11/0/g' PEDS.txt

paste Names.txt PEDS.txt > GemClustAllGenos.txt
sed -i 's/\t/ /g' GemClustAllGenos.txt
sed -i 's/  */ /g' GemClustAllGenos.txt

rm peds.txt
rm Names.txt
rm PEDS.txt
AllSnps4GemClust.*
```

##Plot PCA -- Needs Work...
```{r PlotPCA, cache = FALSE, echo=FALSE}
library(ggplot2)
library(plotrix)
library(scater)
library(data.table)
source('GemTools.R')

#Load Eigne Vals and Calc Percent Var for first 2 Comps.
Comps <- as.vector(read.table(file="Total_PCA.eigenval"))
PC1_Var_exp <- Comps[1,1]/sum(Comps[,1])
PC2_Var_exp <- Comps[2,1]/sum(Comps[,1])

#Load Populations
CEU <- read.table(file = "1000G_RefSamps/CEU.samples.list", header =F, sep ="\t")
row.names(CEU) <- paste0(CEU[,1], "_",CEU[,1] )
YRI <- read.table(file = "1000G_RefSamps/YRI.samples.list", header =F, sep ="\t")
row.names(YRI) <- paste0(YRI[,1], "_",YRI[,1] )
CHB <- read.table(file = "1000G_RefSamps/CHB.samples.list", header =F, sep ="\t")
row.names(CHB) <- paste0(CHB[,1], "_",CHB[,1] )
JPT <- read.table(file = "1000G_RefSamps/JPT.samples.list", header =F, sep ="\t")
row.names(JPT) <- paste0(JPT[,1], "_",JPT[,1] )

ROSMAP <- read.table(file = "ROSMAP_4_Merge.fam", header =F, sep =" ")
row.names(ROSMAP) <- paste0(ROSMAP[,1], "_",ROSMAP[,2] )
MSBB <- read.table(file = "MSBB_4_Merge.fam", header =F, sep =" ")
row.names(MSBB) <- paste0(MSBB[,1], "_",MSBB[,2] )
Mayo <- read.table(file = "Mayo_4_Merge.fam", header =F, sep =" ")
row.names(Mayo) <- paste0(Mayo[,1], "_",Mayo[,2] )
Ref1000G <- read.table(file = "1000_4_Merge.fam", header =F, sep =" ")
row.names(Ref1000G) <- paste0(Ref1000G[,1], "_",Ref1000G[,2] )

pcs = read.table( file="Total_PCA.eigenvec")
row.names(pcs) <- paste0(pcs$V1, "_", pcs$V2)
pcs <- as.data.frame( pcs[,1:4] )
colnames(pcs) <- c( "FID", "IID", "PC1", "PC2")

RMs <- rep( "ROSMAP", length(row.names(ROSMAP)) )
names(RMs) <- row.names(ROSMAP)
MSBs = rep("MSBB", length(row.names(MSBB)) )
names(MSBs) <- row.names(MSBB)     
Mays <- rep( "MAYO", length(row.names(Mayo)) )
names(Mays) <- row.names(Mayo)
Gs = rep("1000G", length(row.names(Ref1000G)) )
names(Gs) <- row.names(Ref1000G)

CEUs <- rep( 'CEU', length(row.names(CEU)) )
names(CEUs) <- row.names(CEU)
YRIs <- rep( 'YRI', length(row.names(YRI)) )
names(YRIs) <- row.names(YRI)
CHBs <- rep( 'CHB', length(row.names(CHB)) )
names(CHBs) <- row.names(CHB)
JPTs <- rep( 'JPT', length(row.names(JPT)) )
names(JPTs) <- row.names(JPT)

SampleCohort <- c( RMs, MSBs, Mays, CEUs, YRIs, CHBs, JPTs)
pcs$SampleCohort <- as.factor( SampleCohort[ row.names(pcs) ] )
pcs$SampleCohort <- factor(pcs$SampleCohort, levels = c("MAYO", "MSBB", "ROSMAP", "YRI", "CHB", "CEU", "JPT"))

SampleType <- c( RMs, MSBs, Mays, Gs )
pcs$SampleType <- as.factor( SampleType[ row.names(pcs) ] )

P <- ggplot(pcs, aes( PC1, PC2) ) + geom_point( aes(colour = SampleCohort, shape = SampleType), cex=.8, alpha = 4/10) 
P <- P + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
P <- P + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" )) 
P <- P + ggtitle("Ancestry PCA Analysis") + theme(plot.title = element_text(hjust = 0.5))
P

##Plot independent cohorts
p = list()
p[[1]] = ggplot(pcs[pcs$SampleCohort %in% c( 'CEU', 'JPT', 'CHB', 'YRI', 'MSBB' ),], aes( PC1, PC2) ) + geom_point( aes(colour = SampleCohort, shape = SampleType), cex=.8, alpha = 4/10)
p[[1]] = p[[1]] + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
p[[1]] = p[[1]] + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
p[[1]] = p[[1]] + ggtitle("MSBB Ancestry PCA Analysis") + theme(plot.title = element_text(hjust = 0.5))

p[[2]] = ggplot(pcs[pcs$SampleCohort %in% c( 'CEU', 'JPT', 'CHB', 'YRI', 'MAYO' ),], aes( PC1, PC2) ) + geom_point( aes(colour = SampleCohort, shape = SampleType), cex=.8, alpha = 4/10)
p[[2]] = p[[2]] + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
p[[2]] = p[[2]] + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
p[[2]] = p[[2]] + ggtitle("Mayo Ancestry PCA Analysis") + theme(plot.title = element_text(hjust = 0.5))

p[[3]] = ggplot(pcs[pcs$SampleCohort %in% c( 'CEU', 'JPT', 'CHB', 'YRI', 'ROSMAP' ),], aes( PC1, PC2) ) + geom_point( aes(colour = SampleCohort, shape = SampleType), cex=.8, alpha = 4/10)
p[[3]] = p[[3]] + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
p[[3]] = p[[3]] + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
p[[3]] = p[[3]] + ggtitle("ROSMAP Ancestry PCA Analysis") + theme(plot.title = element_text(hjust = 0.5))
#multiplot(plotlist = p, rows = 3)
p[[1]]
p[[2]]
p[[3]]

##Clustering All Samples 
gnt <- as.matrix( fread("PEDS.txt", header = F, sep =" ") )
id <- read.table("Names.txt", header = F, sep =" ")
gnt[gnt <  0 | gnt > 2] = NA
#293,239 SNPS

ids <- paste0( id[,1], "_", id[,2] )
row.names(gnt) <- ids
#pcs_Clust = read.table( file="GemClustAllGenos.txt", sep=" ", header=F)

#Only Consider Comps that explain > 1% of Variance
Upper <- as.numeric(table(Comps[,1]/sum(Comps[,1])>0.01)["TRUE"])
EC_foo.cluster = clusterGem(gnt = gnt, id = ids, min.dim = 3, max.dim = Upper)

Refs <- pcs[ pcs$SampleType == '1000G', ]
Refs$SampleCohort <- paste0( '1000G', '_', Refs$SampleCohort)
CEU_Clusters <- c(names( table(EC_foo.cluster$clusters[ names(EC_foo.cluster$clusters) %in% names(CEUs) == T ]) ))
pcs <- cbind( pcs, Cluster = EC_foo.cluster$clusters)

pcs$Ancestry <- "NonCEU"
pcs[ (pcs$Cluster %in% CEU_Clusters) == T,]$Ancestry <- "CEU"
pcs[ row.names(Refs), ]$Ancestry <- Refs$SampleCohort

P <- ggplot(pcs, aes( PC1, PC2) ) + geom_point( aes(colour = Cluster, shape = SampleType), cex=.8, alpha = 4/10) 
P <- P + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
P <- P + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
P <- P + ggtitle("Ancestry Clustering Analysis") + theme(plot.title = element_text(hjust = 0.5))
P

mydf$task <- factor(mydf$task, levels = c("up", "down", "left", "right", "front", "back"))
P <- ggplot(pcs, aes( PC1, PC2) ) + geom_point( aes(colour = Ancestry, shape = SampleType), cex=.8, alpha = 4/10) 
P <- P + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
P <- P + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
P <- P + ggtitle("Ancestry Cluster Analysis CEU vs NonCEU") + theme(plot.title = element_text(hjust = 0.5))
P

CEUs <- pcs[ ( pcs$Ancestry %in% c('CEU', "1000G_CEU", "1000G_CHB", "1000G_YRI", "1000G_JPT") ) == T, ]
P <- ggplot(CEUs, aes( PC1, PC2) ) + geom_point( aes(colour = Ancestry, shape = SampleType), cex=.8, alpha = 4/10)
P <- P + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
P <- P + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
P <- P + ggtitle("Ancestry Cluster Analysis CEU") + theme(plot.title = element_text(hjust = 0.5))
P

CEUs <- pcs[ ( pcs$Ancestry %in% c('CEU', "1000G_CEU") ) == T, ]
P <- ggplot(CEUs, aes( PC1, PC2) ) + geom_point( aes(colour = Ancestry, shape = SampleType), cex=.8, alpha = 4/10)
P <- P + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
P <- P + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
P <- P + ggtitle("Ancestry Analysis: CEU Clustered Samples") + theme(plot.title = element_text(hjust = 0.5))
P
#Write INDV to Keep to a File:
write.table( file = "CEU_AncestryClustered_INDVs.tsv", CEUs[ CEUs$Ancestry == 'CEU', c(1,2) ], col.names = F, row.names = F, quote = F, sep = "\t")
```

## Filter Genotypes for CEU Ancestry 
```{bash FIlterAncestry, cache=FALSE, echo=TRUE, results='hide'}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
CORES=`expr $(nproc) - 2`

plink --threads $CORES --bfile Merged_GenoTypes --keep CEU_AncestryClustered_INDVs.tsv --make-bed --out CEU_AncestryMatched_Genotypes
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn20798011';
activityName = 'Ancestry Analysis';
activityDescription = 'Processing and Analysis of WGS and Genotype data along with ancestry analysis with 1000 genome\'s CEU, CHB, YRI, and JPT populations';
thisFileName <- 'Ancestry_PCA.Rmd'
# Github link
thisRepo <- getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

```{r Store, cache = FALSE, echo=FALSE, results='hide'}
parentId = 'syn20798011';
activityName = 'Ancestry Analysis';
activityDescription = 'Processing and Analysis of WGS and Genotype data along with ancestry analysis with 1000 genome\'s CEU, CHB, YRI, and JPT populations';

# Code WTF is this??
CODE <- Folder(name = "Ancestry_Analysis", parentId = parentId)
CODE <- synStore(CODE)

#Set Used SynIDs For Provenance
Syns_Used <- c("syn11707204", "syn8650955", "syn5879161", "syn5879456", "syn5879461", 
               "syn5879463", "syn5879470", "syn5879472", "syn5879473", "syn5879559", 
               "syn5879628", "syn5879677", "syn5879748", "syn5879792","syn5879812", 
               "syn5879827", "syn5879828", "syn5879830", "syn5879832", "syn5879834", 
               "syn5879835", "syn5879836", "syn5879837", "syn5879838", "syn20809809")

# Set annotations
all.annotations = list(
  dataType = 'GWAS',
  dataSubType = 'DNAVars',
  summaryLevel = 'SNP',
  assay	 = c('Genotype Array', "WGS"),
  tissueTypeAbrv	= c('TCX', "FP", 'IFG', 'PHG', 'STG', 'DLPFC', 'CBE'), 
  study = c('MAYO', 'ROSMAP', 'MSBB'), 
  organism = 'HomoSapiens',
  consortium	= 'AMPAD',
  normalizationStatus	= TRUE,
  normalizationType	= '',
  rnaquantification = '',
  genomeAssemblyID = 'GRCh37'
)

#STORE
#CEU Indvs
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryClustered_INDVs.tsv', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryClustered_INDVs.tsv', 
                  name = 'Samples Clustered to CEU Ancestry', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'CEU_INDV'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#PCA EginVecs
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'Total_PCA.eigenvec', sep = '\t', quote=F)
ENRICH_OBJ = File('Total_PCA.eigenvec', 
                  name = 'PCA Eigen Vectors', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'EigenVecs'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#PCA EginVecs
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'Total_PCA.eigenval', sep = '\t', quote=F)
ENRICH_OBJ = File('Total_PCA.eigenval', 
                  name = 'PCA Eigen Values', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'EigenValues'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

###Store Genos Seperatly 
CODE <- Folder(name = "Ancestry_Clustered_Genotypes", parentId = parentId)
CODE <- synStore(CODE)

#Geno Bed
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.bed', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.bed', 
                  name = 'Genotype Bed File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno Bim
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.bim', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.bim', 
                  name = 'Genotype Bim File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno Fam
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.fam', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.fam', 
                  name = 'Genotype Fam File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno nosex
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.nosex', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.nosex', 
                  name = 'Genotype nosex File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno nosex
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.log', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.log', 
                  name = 'Genotype log File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)
```
### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapser)
library(knit2synapse) 
#install.packages("reticulate")
library(reticulate)

# get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "Ancestry_PCA.Rmd",
                                          parentId ="syn20798011",
                                          folderName = 'Ancestry_Analysis')

```