### 2019-10-03 code/ProcessData.Rmd
Processed Data Code

---
title: "Process RNA-Seq and Genotype Data for TWAS"
author: "JKG"
date: "9/25/2019"
output: pdf_document
  chunk_output_type: console
---

```{r libs, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  results='hide')
#library(reticulate)
```
##Before running 
Run the following in docker exec -it <CONTAINER> /bin/bash 
chmod 777 /root/
mv TWAS/ /home/<USR>/TWAS/
chown -hR <GID>:<USR> /home/<USR>/TWAS/
Create a ~/.synapseConfig file in the container, but do not push a container containing your credentials file to Docker Hub!!!

## Login to Synapse Will need to replace with your credentials
```{bash Login, echo=T, results='hide', eval=FALSE}
source /root/.bashrc
synapse login -u <USR> -p <PSWD> --rememberMe

#Alternativly you can setup a credentials file as such:
touch ~/.synapseConfig
echo "[authentication]" >> ~/.synapseConfig
echo "username = <USR>" >> ~/.synapseConfig
echo "password = <PASWD>" >> ~/.synapseConfig
```

```{bash PLINKSetup, echo=F, include = T, results='hide', eval=T }
mkdir ~/TWAS/bin
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink1-assets/plink_linux_x86_64_20190617.zip 
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink_linux_x86_64_20190617.zip
wget -q -P ~/TWAS/bin  http://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20191002.zip
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink2_linux_x86_64_20191002.zip

rm ~/TWAS/bin/*.zip
rm ~/TWAS/bin/toy.*
rm ~/TWAS/bin/LICENSE
```

## MSBB RNA-Seq: Average Across Brain Regions, Filter for INDV. w/Genotype Data, Rename Samples with Genotype ID 
```{r MSBBRNA, echo=F, results='hide', cache = FALSE}
library(data.table)

Process_MSBB <- function( dna, exp, id){
  #'@dna synID of Geno samples eg 'syn11384608'
  #'@exp synID of expression data eg 'syn8485026'
  #'@id synID of id translators  eg. 'syn6100548'
  #dna<-'syn11384608'
  #exp<-'syn8485026'
  #id<-'syn6100548'

  DNA <- fread(syn_temp$get(dna)$path)
  Expression <- read.table(syn_temp$get(exp)$path, header=T)
  IDs <- fread(syn_temp$get(id)$path)
  IDs <- IDs[ grepl( 'accepted_hits.sort.coord.bam', IDs$fileName ) == T, ]
  IDs <- IDs[ (IDs$sampleIdentifier %in% colnames(Expression)) == T, ]

  INDVs <- IDs$individualIdentifier[ !duplicated(IDs$individualIdentifier) ]

  if( 'ensembl_gene_id' %in% colnames(Expression) ){
    Averaged <- as.data.frame( Expression$ensembl_gene_id )
  }else{
    Averaged <- as.data.frame( row.names(Expression) )
    colnames(Averaged) <- 'ensembl_gene_id'
  }

  Ave <- function( ID, DF, Exp ){
    #'@ID Sample ID eg. 'AMPAD_MSSM_0000007155'
    #'@DF the target dataframe eg. 'IDs'
    #'@Exp the expression set to filter eg Expression

    Exp <- as.data.frame( Exp )
    COLS <- DF[ DF$individualIdentifier == ID, ]$sampleIdentifier
    Small <- Exp[ , COLS ]

    if( dim(as.matrix(Small))[2] > 1 ){
      eval(parse( text = paste0( 'return( ', ID, ' = rowMeans(Small) )')))
    }else{
      eval(parse( text = paste0( 'return( ', ID, ' = Small)')))
    }
  }

  for( i in INDVs ){
    eval(parse( text = paste0( 'Averaged$', i, ' <- Ave(i, IDs, Expression )') ))
  }

  GWAS_ID <- as.data.frame( DNA[ (DNA$individualIdentifier %in% colnames(Averaged)) == T , ] )
  row.names(GWAS_ID) <- GWAS_ID$individualIdentifier
  if( 'ensembl_gene_id' %in% colnames(Averaged) ){
    Averaged <- Averaged[ , c( 'ensembl_gene_id', row.names(GWAS_ID) )]
  }else{
    Averaged <- Averaged[ , c( 'Expression$ensembl_gene_id', row.names(GWAS_ID) )]
    colnames(Averaged[1]) <- 'ensembl_gene_id'
  }
  colnames(Averaged) <- paste0(GWAS_ID[ colnames(Averaged), ]$sampleIdentifier, "_", GWAS_ID[ colnames(Averaged), ]$sampleIdentifier)
  colnames(Averaged)[1] <- "ensembl_gene_id"
  return( Averaged )
}

Norm_Resid_MSBB <- Process_MSBB('syn11384608', 'syn8485026', 'syn6100548')
dCorrect_Resid_MSBB <- Process_MSBB('syn11384608', 'syn20801798', 'syn6100548')
```

## Mayo RNA-Seq: Remove Cerebellum, Filter for INDV. w/Genotype Data, Rename Samples with Genotype ID 
```{r MayoRNA, echo=T, results='hide', cache = FALSE, echo=FALSE}
#Get Geno Names From FAM File
FAM <- as.data.frame(fread( syn_temp$get('syn20820122')$path))
FAM <- FAM[ (grepl('ROS', FAM[,1])) == F , ]
FAM <- FAM[ (grepl('MAP', FAM[,1])) == F , ]
FAM <- FAM[ (grepl('ROS', FAM[,2])) == F , ]
FAM <- FAM[ (grepl('MAP', FAM[,2])) == F , ]
FAM <- FAM[ FAM[,1] == 0, ]

#Filter Expression Samples by Presence of a Genotype
Mayo_filt <- function( syn, fam, Type ){
  #'@syn A synID for the RNA Seq eg. 'syn20801999' or 'syn20801991'
  #'@fam A fam file eg. FAM
  #'@Type fread = 0 read.table=1

  fam <- FAM
  if( Type == 0){
    RNA <- read.table( syn_temp$get(syn)$path)
    colnames(RNA) <- gsub('_TCX', '_a', colnames(RNA))
    colnames(RNA) <- gsub('X', '', colnames(RNA))
    colnames(RNA) <- gsub('_a', '_TCX', colnames(RNA))
    RNA <- cbind( 'ensembl_gene_id' = row.names(RNA), RNA[,colnames(RNA)] )
  }else{
    RNA <- as.data.frame( fread( syn_temp$get(syn)$path ) )
  }
  RNA <- RNA[ , c('ensembl_gene_id', colnames(RNA)[(grepl('_TCX', colnames(RNA))==T)]) ]
  colnames( RNA ) <- gsub( '_TCX', '', colnames( RNA ) )
  colnames( RNA )[2:length(colnames( RNA ))] <- paste0( '0_', colnames( RNA )[2:length(colnames( RNA ))] )

  RNA <- RNA[ , c( "ensembl_gene_id", colnames(RNA)[ (colnames(RNA) %in% paste0( FAM[,1], "_", FAM[,2] )) == T ]) ]
  return(RNA)
}
Norm_Resid_MAYO <- Mayo_filt( 'syn8466825', FAM, 1 )
dCorrect_Resid_MAYO <- Mayo_filt( 'syn20801991', FAM, 0 )
```

## ROSMAP RNA-Seq: Filter for INDV. w/Genotype Data and Rename Samples with Genotype ID 
```{r ROSMAPoRNA, echo=T, results='hide', cache = FALSE, echo=FALSE}
#SVA: 'syn8456704'
#SVA+Diag: 'syn20800652'
Ros_Filt <- function(Exp, name, Trans, Fam, Var){
  #'@Exp the expression matrix eg. 'syn8456704'
  #'@name another name file eg. 'syn5879839'
  #'@Trans 'syn3382527'
  #'@Fam a .fam file of the ROS samps syn20820122
  #'@Var 1=Nondiagnosis correcte 0=Diagnosis Corrected

  # eg. Ros_Filt( 'syn8456704', 'syn5879839', 'syn3382527','syn20820122')

  FAM <- as.data.frame(fread( syn_temp$get(Fam)$path))
  Fam <- FAM[ (grepl('ROS', FAM[,1])) == T | (grepl('MAP', FAM[,1])) == T | (grepl('ROS', FAM[,2])) == T | (grepl('MAP', FAM[,2])) == T, ]
  InvFam <- FAM[ (grepl('ROS', FAM[,1])) == F & (grepl('MAP', FAM[,1])) == F & (grepl('ROS', FAM[,2])) == F & (grepl('MAP', FAM[,2])) == F, ]
  Fam[,1] <- Fam[,2]

  #Need to assign sex to MSBB and Mayo
  NAMED_FAM <- rbind( InvFam[1:dim(InvFam)[1], ],  Fam[1:dim(Fam)[1], ] )

  if(Var == 1){
    Exp <- as.data.frame( fread(syn_temp$get(Exp)$path) )
  }else{
    if(Var == 0){
        Exp <- as.data.frame( read.table(syn_temp$get('syn20800652')$path) )
        colnames(Exp) <- gsub('X','', colnames(Exp))
        Exp <- cbind(row.names(Exp), Exp[1:dim(Exp)[2]])
        colnames(Exp)[1] <- 'ensembl_gene_id'
    }else{}
  }
  Trans <- fread(syn_temp$get(Trans)$path)
  Trans <- Trans[ Trans$gwas_id != "", ]
  names <- as.data.frame(fread(syn_temp$get(name)$path))
  Trans <- Trans[ (Trans$gwas_id  %in%  names[,2]) == T, ]
  Trans <- Trans[ (Trans$rnaseq_id == "") == F, ]
  Trans <- Trans[ !duplicated(Trans$rnaseq_id), ]

  #Filter for matched RNA-Seq and GWAS samples
  Exp <- Exp[ ,c('ensembl_gene_id', colnames(Exp)[ (colnames(Exp) %in% Trans$rnaseq_id) == T]) ]
  Trans <- as.data.frame( Trans[ (Trans$rnaseq_id %in% colnames(Exp)) == T,  ] )
  row.names(Trans) <- Trans$rnaseq_id

  colnames(Exp) <- c('ensembl_gene_id', paste0( Trans[colnames(Exp)[2:length(colnames(Exp))],]$gwas_id, "_",  Trans[colnames(Exp)[2:length(colnames(Exp))],]$gwas_id))
  return(Exp)
}

Norm_Resid_RM <-Ros_Filt( 'syn8456704', 'syn5879839', 'syn3382527','syn20820122', 1)
dNorm_Resid_RM <-Ros_Filt( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122', 0)
```

## Merge RNA-Seq: Merge and Annotate RNA-Seq
* Remove genes on Chr X, Chr Y, Chr M 
* Check for no Patch Contig Nonsense
* Print Sample ID List
* Filter Genotypes for RNA Samples
```{r Merge RNASeq, echo=T, results='hide', cache = FALSE, echo=FALSE}
#Common Genes:
GNames <- names( table(c( as.character(Norm_Resid_MSBB$ensembl_gene_id), as.character(Norm_Resid_MAYO$ensembl_gene_id), as.character(Norm_Resid_RM$ensembl_gene_id)  ))[ table(c( as.character(Norm_Resid_MSBB$ensembl_gene_id), as.character(Norm_Resid_MAYO$ensembl_gene_id), as.character(Norm_Resid_RM$ensembl_gene_id)  )) == 3] )

#Combine Into One DataFrame
Comb <- function( MSBB, Mayo, RM, Names){
  #MSBB <- Norm_Resid_MSBB
  #Mayo <- Norm_Resid_MAYO
  #RM <- Norm_Resid_RM
  #Names <- GNames

  row.names(MSBB) <- as.character(MSBB$ensembl_gene_id)
  row.names(Mayo) <- as.character(Mayo$ensembl_gene_id)
  row.names(RM) <- as.character(RM$ensembl_gene_id)

  MSBB <- MSBB[ GNames ,2:dim(MSBB)[2] ]
  Mayo <- Mayo[ GNames ,2:dim(Mayo)[2] ]
  RM <- RM[ GNames ,2:dim(RM)[2] ]

  Out <- cbind( MSBB[GNames,][ ,1:dim(MSBB)[2] ],
                Mayo[GNames, ][ ,1:dim(Mayo)[2] ],
                RM[GNames, ][, 1:dim(RM)[2] ]
  )

  return( Out )
}

Norm_Resid <- Comb( Norm_Resid_MSBB, Norm_Resid_MAYO, Norm_Resid_RM, GNames )
Diag_Resid <- Comb( dCorrect_Resid_MSBB,  dCorrect_Resid_MAYO, dNorm_Resid_RM, GNames )

#Annotate for TWAS
hg19 <- as.data.frame(fread(syn_temp$get('syn20826610')$path))
hg19 <- hg19[ hg19[,1] %in% row.names( Norm_Resid ), ]
row.names(hg19) <- hg19[,1]

#Filter out the 231 Genes that don't have coherent annotations between GRCH38 and GRCH 37
Norm_Resid <- Norm_Resid[ (row.names(Norm_Resid) %in% row.names(hg19)) == T, ]
Diag_Resid <- Diag_Resid[ (row.names(Diag_Resid) %in% row.names(hg19)) == T, ]

#Finalize
Norm_Resid <- as.data.frame( cbind( 
               TargetID = row.names(Norm_Resid), 
               Gene_Symbol = row.names(Norm_Resid),
               Chr = hg19[row.names(Norm_Resid), ][,2] , 
               Coord = hg19[ row.names(Norm_Resid), ][,3],
               Norm_Resid[ ,1:dim(Norm_Resid)[2]
             ]) )

Diag_Resid <- as.data.frame(cbind( 
               TargetID = row.names(Diag_Resid), 
               Gene_Symbol = row.names(Diag_Resid),
               Chr = hg19[row.names(Diag_Resid), ][,2] , 
               Coord = hg19[ row.names(Diag_Resid), ][,3],
               Diag_Resid[ ,1:dim(Diag_Resid)[2]
             ]) )

#Filter out the ~ 480 ChrX genes
Norm_Resid <- Norm_Resid[ (Norm_Resid$Chr %in% 'X') == F, ]
Diag_Resid <- Diag_Resid[ (Diag_Resid$Chr %in% 'X') == F, ]

#Order:
Norm_Resid <- Norm_Resid[ with(Norm_Resid, order( Chr, Coord)), ]
Diag_Resid <- Diag_Resid[ with(Diag_Resid, order( Chr, Coord)), ]

#Filter for CEU Ancestry Matched Training Cohort
Tot_Samps <- do.call( rbind, strsplit( colnames(Norm_Resid)[5:length(colnames(Norm_Resid))],  "_" ))
write.table(Tot_Samps, file="ALL_RNASamplesComb.tsv", col.names =F, row.names=F, sep=" ", quote=F)

#Output the ROSMAP INDV CEU Matched Training Cohort
ROS <- colnames(Norm_Resid)[ grepl('ROS', colnames(Norm_Resid)) == T | grepl('MAP', colnames(Norm_Resid)) == T ]
ROS_Norm_Resid <- Norm_Resid[ ,c("TargetID", "Gene_Symbol", "Chr", "Coord", ROS) ]
ROS_Diag_Resid <- Diag_Resid[ ,c("TargetID", "Gene_Symbol", "Chr", "Coord", ROS) ]

ROS_Samps <- do.call( rbind, strsplit( colnames(ROS_Norm_Resid)[5:length(colnames(ROS_Norm_Resid))],  "_" ))
write.table(ROS_Samps, file="ROS_RNASamples.tsv", col.names =F, row.names=F, sep=" ", quote=F)
```

## Fix FamFile
```{r FamFix, cache=TRUE}
FAM <- as.data.frame(fread( syn_temp$get('syn20820122')$path))
MSBB <-  as.data.frame(fread( syn_temp$get('syn11384608')$path))
MSBB_Clin <- as.data.frame(fread( syn_temp$get('syn6101474')$path))
ROSMAP <- as.data.frame(fread( syn_temp$get('syn3191087')$path))
row.names(ROSMAP) <- ROSMAP$projid
ROSMAP_ID <- as.data.frame(fread( syn_temp$get('syn3382527')$path))
MAYO <- as.data.frame(fread( syn_temp$get('syn19164046')$path))
mayo <- as.data.frame(fread( syn_temp$get('syn19164050')$path))

#MSBB 
row.names(MSBB) <- MSBB$sampleIdentifier
name <- FAM[ FAM$V2 %in% row.names(MSBB) == T, ]$V2
MSBB <- MSBB[name,]

row.names(MSBB_Clin) <- MSBB_Clin$individualIdentifier

MSBB_Clin <- cbind(MSBB_Clin[MSBB$individualIdentifier,1:dim(MSBB_Clin)[2]], MSBB[,1:dim(MSBB)[2] ] )
row.names(MSBB_Clin) <- MSBB_Clin$sampleIdentifier

MSBB_Clin$Diagnosis <- -9
for( i in 1:length(row.names(MSBB_Clin))){
  if( is.na(MSBB_Clin[ i,]$CDR) ==  F & is.na(MSBB_Clin[ i,]$bbscore) ==  F & is.na(MSBB_Clin[ i,]$NP.1) ==  F ){
    if( MSBB_Clin[ i,]$CDR >= 1 & MSBB_Clin[ i,]$bbscore > 3 & MSBB_Clin[ i,]$NP.1 > 1 ){
      MSBB_Clin[ i,]$Diagnosis <- 2
    }else{
      if(MSBB_Clin[ i,]$CDR < 1 & MSBB_Clin[ i,]$bbscore < 4 & MSBB_Clin[ i,]$NP.1 < 2){
        MSBB_Clin[ i,]$Diagnosis <- 1
      }else{}
    }
  }
}

Sex<-c(1,2)
names(Sex) <- c('M', 'F')
MSBB$SEX.inferred <- Sex[MSBB$SEX.inferred]
row.names(FAM)<-FAM$V2
FAM[ row.names(MSBB), ]$V5 <- MSBB$SEX.inferred
FAM[ row.names(MSBB_Clin), ]$V6 <- MSBB_Clin$Diagnosis

#Mayo
mayo <- mayo[complete.cases(mayo$Gender),]
MAYO <- MAYO[complete.cases(MAYO$Sex),]

NewMayo <- as.data.frame( rbind( cbind( MAYO$SampleID,MAYO$Sex,as.character(MAYO$Diagnosis)), cbind( mayo$SampleID,mayo$Gender, as.character(mayo$Diagnosis))) )
NewMayo<- NewMayo[!duplicated(NewMayo),]
NewMayo <- as.data.frame( cbind( as.character(NewMayo$V1),as.character(NewMayo$V2), gsub("_CER","",gsub("_TCX","",NewMayo$V1)), as.character(NewMayo$V3)) )
NewMayo$V2 <- Sex[as.character( NewMayo$V2 )]

names <- FAM[ FAM$V1 == 0 ,]$V2
NewMayo <- NewMayo[ !duplicated(NewMayo$V3), ]
row.names(NewMayo) <- NewMayo$V3
NewMayo <- NewMayo[ names, ]

Aff <- c(-9,-9,1,2)
names(Aff) <- c("Pathologic Aging", "PSP", "Control", "AD")
FAM[ names ,]$V6 <- as.numeric(Aff[as.character(NewMayo[ names, ]$V4)])
FAM[ names ,]$V5 <- NewMayo[ names, ]$V2

#Fix the Final
FAM[ '1957',]$V5 <- 2
FAM[ '1957',]$V6 <- -9
FAM[ '1950',]$V5 <- 2
FAM[ '1950',]$V6 <- -9

#Fix ROSMAP Names
FAM[ grepl('ROS', FAM$V2) == T  | grepl('MAP', FAM$V2) == T , ]$V1 <- FAM[ grepl('ROS', FAM$V2) == T  | grepl('MAP', FAM$V2) == T , ]$V2

#Write New FAM to file
write.table(FAM, file='New_Fam.fam', row.names = F, col.names=F, quote=F, sep=" ")
```

## Filter Genotype data
```{bash FilterGenos, include=FALSE, cache=TRUE}
#source /root/.bashrc
PATH=$PATH:~/TWAS/bin/
CORES=`expr $(nproc) - 2`

#Pull Geno-data
synapse get syn20820120
synapse get syn20820121
synapse get syn20820122

#Copy over FAM File with Sex Info:
mv New_Fam.fam CEU_AncestryMatched_Genotypes.fam

#Extract CEU
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --keep ALL_RNASamplesComb.tsv --silent --make-bed --out All_CEU_ToTrainForTWAS

#Extract ROS_Only:
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --keep ROS_RNASamples.tsv --make-bed --out ROSMAPonly_CEU_ToTrainForTWAS
```

```{r sanity check}
FAM <- as.data.frame(fread('All_CEU_ToTrainForTWAS.fam'))
row.names(FAM) <- paste0( FAM[,1], '_', FAM[,2] )
table(row.names(FAM) %in% colnames(Diag_Resid))
Diag_Resid_train <- Diag_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]
Norm_Resid_train <- Norm_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]

FAM <- as.data.frame(fread('ROSMAPonly_CEU_ToTrainForTWAS.fam'))
row.names(FAM) <- paste0( FAM[,1], '_', FAM[,2] )
table(row.names(FAM) %in% colnames(ROS_Diag_Resid))
ROS_Diag_Resid_train <- ROS_Diag_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]
ROS_Norm_Resid_train <- ROS_Norm_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]

```

## Push Results to Synapse
```{r synapse.parameters.head, include=FALSE, cache=TRUE}
parentId = 'syn18936948';
activityName = 'Processed TWAS Training Data';
activityDescription = 'Data from processing and hamonizing CEU ancestry matched Genotypes with available RNA-Seq profiles which were aligned to GRCh38 but annotated with hg19 TSS sites here';
thisFileName <- 'ProcessData.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
activityName = 'Processed TWAS Training Data';
activityDescription = 'Data from processing and hamonizing CEU ancestry matched Genotypes with available RNA-Seq profiles';

CODE <- syn_temp$store(synapseclient$Folder(name = "Processed_TWAS_Training_Data", parentId = parentId))

#Set Used SynIDs For Provenance
Syns_Used <- c("syn5223705", "syn8698211", "syn3817650", "syn8690904", 
"syn6126119", "syn8698214", "syn8449369", "syn6126114", "syn6145639", "syn8690799"
)

# Set annotations
EXP.all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  tissueTypeAbrv	= c('TCX', 'DLPFC', 'PFG', 'STG', 'IFG', 'FP'), 
  study = c('MAYO','MSBB', 'ROSMAP'), 
  organism = 'HomoSapiens',
  consortium	= 'AMPAD',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh38'
)
# Set annotations
Var.all.annotations = list(
  dataType = 'Variant',
  dataSubType = 'SNP',
  summaryLevel = 'Nucleotide',
  assay	 = 'Genotype Array',
  tissueTypeAbrv	= c('Whole Blood'), 
  study = c('MAYO','MSBB', 'ROSMAP'), 
  organism = 'HomoSapiens',
  consortium	= 'AMPAD',
  genomeAssemblyID = 'GRCh37'
)

Basic.annotations = list(
  dataType = 'Induvidual List',
  dataSubType = 'Plink PatientID Filter',
  study = c('MAYO','MSBB', 'ROSMAP'), 
  organism = 'HomoSapiens',
  consortium	= 'AMPAD'
)

## TOTAL SVA-only Exp
write.table(Norm_Resid, file="ALL_RNA_Seq_SVAadusted.tsv", col.names =F, row.names=F, sep="\t", quote=F)
Syns_Used_SVA_EXP <- c('syn11384608', 'syn8485026', 'syn6100548', 'syn8456704', 'syn5879839', 'syn3382527','syn20820122', 'syn8466825', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_RNA_Seq_SVAadusted.tsv', name = 'Total Merged and region Avereaged RNA-Seq with Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_SVA_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
EXP.all.annotations$dataSubType = 'residualGeneExpForeQTL'
syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

## TOTAL Diagnosis Exp
write.table(Diag_Resid, file="ALL_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv", col.names =F, row.names=F, sep="\t", quote=F)
Syns_Used_Diagnosis_EXP <- c('syn20801991', 'syn11384608', 'syn20801798', 'syn6100548',  'syn20800652', 'syn5879839', 'syn3382527','syn20820122', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv', name = 'Total Merged and region Avereaged RNA-Seq with Diagnosis Regressed Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_Diagnosis_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
    EXP.all.annotations$dataSubType = 'residualGeneExpForDiagnosisRegressedeQTL'
    syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

## CEU SVA-only Exp
write.table(Norm_Resid_train, file="ALL_CEU_RNA_Seq_SVAadusted.tsv", col.names =F, row.names=F, sep="\t", quote=F)
Syns_Used_SVA_EXP <- c('syn11384608', 'syn8485026', 'syn6100548', 'syn8456704', 'syn5879839', 'syn3382527','syn20820122', 'syn8466825', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_CEU_RNA_Seq_SVAadusted.tsv', name = 'CEU Merged and region Avereaged RNA-Seq with Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_SVA_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

## CEU Diagnosis Exp
write.table(Diag_Resid_train, file="ALL_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv", col.names =F, row.names=F, sep="\t", quote=F)
Syns_Used_Diagnosis_EXP <- c('syn20801991', 'syn11384608', 'syn20801798', 'syn6100548',  'syn20800652', 'syn5879839', 'syn3382527','syn20820122', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv', name = 'CEU Merged and region Avereaged RNA-Seq with Diagnosis Regressed Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_Diagnosis_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForDiagnosisRegressedeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

#ROSMAP only Expression
  ## CEU SVA-only Exp -- ROSMAP Only
write.table(ROS_Norm_Resid_train, file="ROS_CEU_RNA_Seq_SVAadusted.tsv", col.names =F, row.names=F, sep="\t", quote=F)
Syns_Used_SVA_EXP <- c('syn8456704', 'syn5879839', 'syn3382527','syn20820122' )

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ROS_CEU_RNA_Seq_SVAadusted.tsv', name = 'CEU ROSMAP Only RNA-Seq with Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_SVA_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForeQTL'
  EXP.all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

  ## CEU Diagnosis Exp -- ROSMAP Only
write.table(ROS_Diag_Resid_train, file="ROS_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv", col.names =F, row.names=F, sep="\t", quote=F)
Syns_Used_Diagnosis_EXP <- c( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122')
ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ROS_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv', name = 'CEU ROSMAP Only RNA-Seq with Diagnosis Regressed Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_Diagnosis_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForDiagnosisRegressedeQTL'
  EXP.all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

#ALL INDV
 ##Check that order is correct between file lists
  #table(paste0(Tot_Samps[,1], '_', Tot_Samps[,2]) == colnames(Diag_Resid)[ 5:length(colnames(Diag_Resid))])
Syns <- c( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122')
ENRICH_OBJ_Tot_list <- syn_temp$store( synapseclient$File( path='ALL_RNASamplesComb.tsv', name = 'All Indv with RNA-Seq Profiles in AMP-Ad For Filtering', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  syn_temp$setAnnotations(ENRICH_OBJ_Tot_list, annotations = Basic.annotations)

#ROSMAP INDV
 ##Check that order is correct between file lists
  #table(paste0(ROS_Samps[,1], '_', ROS_Samps[,2]) == colnames(ROS_Norm_Resid)[ 5:length(colnames(ROS_Norm_Resid))])
Syns <- c( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122')
ENRICH_OBJ_RM_List <- syn_temp$store( synapseclient$File( path='ROS_RNASamples.tsv', name = 'ROSMAP Indv with RNA-Seq Profiles in AMP-Ad For Filtering', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  syn_temp$setAnnotations(ENRICH_OBJ_RM_List, annotations = Basic.annotations)

#PLINK ALL -- STORE IN SEPERATE FOLDER INCASE PHI LIMITATIONS APPLY LATER
  CODE <- syn_temp$store(synapseclient$Folder(name = "Ancestry_Clustered_Genotypes", parentId = parentId))
  Syns <- c( 'syn20820120', 'syn20820121', 'syn20820122', ENRICH_OBJ_Tot_list$properties$id  )
#FAM  
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.fam', name = 'All Matched CEU Genotype FAM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BED
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.bed', name = 'All Matched CEU Genotype BED', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#LOG
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.log', name = 'All Matched CEU Genotype LOG', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BIM
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.bim', name = 'All Matched CEU Genotype BIM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#PLINK Rosmap
  Syns <- c( 'syn20800652', 'syn5879839', 'syn3382527', ENRICH_OBJ_RM_List$properties$id)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.fam', name = 'ROSMAP Matched CEU Genotype FAM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BED
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.bed', name = 'ROSMAP Matched CEU Genotype BED', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#LOG
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.log', name = 'ROSMAP Matched CEU Genotype LOG', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BIM
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.bim', name = 'ROSMAP Matched CEU Genotype BIM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

```

##Knit to Synapse
```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

library(data.table)

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")

createAndKnitToFolderEntityClient(file = "ProcessData.Rmd",
                                          parentId ="syn18936948",
                                          folderName = 'Processed_TWAS_Training_Data')
```