### 2019-10-16 code/ProcessData.Rmd
Annotate MSBB By Brain Region

---
title: "Process RNA-Seq and Genotype Data for TWAS"
author: "JKG"
date: "9/25/2019"
output: pdf_document
  chunk_output_type: console
---

```{r libs, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  results='hide')
#library(reticulate)
```
##Before running 
Run the following in docker exec -it <CONTAINER> /bin/bash 
chmod 777 /root/
mv TWAS/ /home/<USR>/TWAS/
chown -hR <GID>:<USR> /home/<USR>/TWAS/
Create a ~/.synapseConfig file in the container, but do not push a container containing your credentials file to Docker Hub!!!

## Login to Synapse Will need to replace with your credentials
```{bash Login, echo=T, results='hide', eval=FALSE}
#source /root/.bashrc
synapse login -u <USR> -p <PSWD> --rememberMe

#Alternativly you can setup a credentials file as such:
touch ~/.synapseConfig
echo "[authentication]" >> ~/.synapseConfig
echo "username = <USR>" >> ~/.synapseConfig
echo "password = <PASWD>" >> ~/.synapseConfig
```

```{bash PLINKSetup, results='hide', eval=F, echo=TRUE}
mkdir ~/TWAS/bin
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink1-assets/plink_linux_x86_64_20190617.zip 
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink_linux_x86_64_20190617.zip
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20191012.zip
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink2_linux_x86_64_20191012.zip

rm ~/TWAS/bin/*.zip
rm ~/TWAS/bin/toy.*
rm ~/TWAS/bin/LICENSE
```

## MSBB RNA-Seq: Label Brain Regions, Filter for INDV. w/Genotype Data, Rename Samples with Genotype ID 
```{r MSBBRNA, results='hide', echo=TRUE, cache = FALSE}
library(data.table)

Process_MSBB <- function( dna, exp, id){
  #'@dna synID of Geno samples eg 'syn11384608'
  #'@exp synID of expression data eg 'syn8485026'
  #'@id synID of RNA id translators  eg. 'syn6100548'

  DNA <- fread(syn_temp$get(dna)$path)
  Expression <- read.table(syn_temp$get(exp)$path, header=T, row.names = 1)
  IDs <- fread(syn_temp$get(id)$path)
  sink <- Expression

  #Make sample IDs the row names
  IDs <- IDs[ IDs$Action != 'Exclude', ]
    ###Convert to Z-Value by Tissue
  Regions <- names( table(IDs$BrodmannArea) ) 
  for( BM in Regions ){
    temp <- Expression[ ,colnames(Expression)[ colnames(Expression) %in% IDs[ IDs$BrodmannArea == BM, ]$sampleIdentifier] ]
    foo <- apply(temp, 1, scale)
    foo <- t(foo)
    colnames(foo) <- colnames(temp)
    Expression[ , colnames(foo)] <- foo [, colnames(foo)]
  }

  IDs <- IDs[ (IDs$sampleIdentifier %in% colnames(Expression)) == T, ]
  Expression <- Expression[ , colnames(Expression)[(colnames(Expression) %in% IDs$sampleIdentifier) == T] ]
  INDVs <- IDs[ IDs$fileType=='fastq' , ]

  DNA <- as.data.frame(DNA)
  DNA[ is.na(DNA$individualIdentifier)==T,]$individualIdentifier <- DNA[ is.na(DNA$individualIdentifier)==T,]$individualIdentifier.inferred
  DNA <- DNA[ complete.cases(DNA$individualIdentifier), ]
  DNA <- DNA[ DNA$Action != "Exclude", ]
  row.names(DNA) <- DNA$individualIdentifier

  INDVs <- IDs[ !duplicated(IDs$sampleIdentifier) ]
  INDVs <- INDVs[ (INDVs$individualIdentifier %in% DNA$individualIdentifier) == T, ]

  DNA <- as.data.frame(DNA)
  DNA <- DNA[ complete.cases(DNA$individualIdentifier), ]
  row.names(DNA) <- DNA$individualIdentifier

  INDVs <- as.data.frame(INDVs)
  INDVs$GenoID <- DNA[ INDVs$individualIdentifier, ]$sampleIdentifier
  INDVs$NewID <- paste0(INDVs$GenoID, '_', INDVs$BrodmannArea )
  row.names(INDVs) <- INDVs$sampleIdentifier
  INDVs[ colnames(Expression), ]$NewID

  Expression <- Expression[ , colnames(Expression)[(colnames(Expression) %in% INDVs$sampleIdentifier) == T] ]
  INDVs <- as.data.frame(INDVs)
  row.names(INDVs) <- INDVs$sampleIdentifier
  colnames(Expression) <- INDVs[ colnames(Expression), ]$NewID

  return( Expression )
}

Norm_Resid_MSBB <- Process_MSBB('syn11384608', 'syn8485026', 'syn6100548')
dCorrect_Resid_MSBB <- Process_MSBB('syn11384608', 'syn20801798', 'syn6100548')
```

## Mayo RNA-Seq: Remove Cerebellum, Filter for INDV. w/Genotype Data, Rename Samples with Genotype ID 
```{r MayoRNA, results='hide', echo=TRUE, cache = FALSE}
#Get Geno Names From FAM File
FAM <- as.data.frame(fread( syn_temp$get('syn20820122')$path))
FAM <- FAM[ (grepl('ROS', FAM[,1])) == F , ]
FAM <- FAM[ (grepl('MAP', FAM[,1])) == F , ]
FAM <- FAM[ (grepl('ROS', FAM[,2])) == F , ]
FAM <- FAM[ (grepl('MAP', FAM[,2])) == F , ]
FAM <- FAM[ FAM[,1] == 0, ]

#Filter Expression Samples by Presence of a Genotype
Mayo_filt <- function( syn, fam, Type ){
  #'@syn A synID for the RNA Seq eg. 'syn20801999' or 'syn20801991'
  #'@fam A fam file eg. FAM
  #'@Type fread = 0 read.table=1

  #syn <- 'syn20801991'
  #fam <- FAM
  #Type <- 1

  fam <- FAM
  if( Type == 0){
    RNA <- read.table( syn_temp$get(syn)$path)
    colnames(RNA) <- gsub('_TCX', '_a', colnames(RNA))
    colnames(RNA) <- gsub('X', '', colnames(RNA))
    colnames(RNA) <- gsub('_a', '_TCX', colnames(RNA))
    RNA <- cbind( 'ensembl_gene_id' = row.names(RNA), RNA[,colnames(RNA)] )
    RNA <- RNA[ , colnames(RNA)[(grepl('_TCX', colnames(RNA))==T)] ]
  }else{
    RNA <- as.data.frame( fread( syn_temp$get(syn)$path ) )
    row.names(RNA) <- RNA$ensembl_gene_id
    RNA <- RNA[ , colnames(RNA)[(grepl('_TCX', colnames(RNA))==T)] ]
  }
  colnames( RNA ) <- gsub( '_TCX', '', colnames( RNA ) )
  colnames( RNA ) <- paste0( '0_', colnames( RNA ) )

  ## Z-Scale
  foo <- as.matrix(RNA)
  foo <- apply(RNA, 1, scale)
  foo <- t(foo)
  colnames(foo) <- colnames(RNA)
  #Expression[ , colnames(foo)] <- foo [, colnames(foo)]

  RNA <- foo[ , colnames(foo)[ (colnames(foo) %in% paste0( FAM[,1], "_", FAM[,2] )) == T ] ]
  return(RNA)
}
Norm_Resid_MAYO <- Mayo_filt( 'syn8466825', FAM, 1 )
dCorrect_Resid_MAYO <- Mayo_filt( 'syn20801991', FAM, 0 )
```

## ROSMAP RNA-Seq: Filter for INDV. w/Genotype Data and Rename Samples with Genotype ID 
```{r ROSMAPoRNA, results='hide', echo=TRUE, cache = FALSE}
#SVA: 'syn8456704'
#SVA+Diag: 'syn20800652'
Ros_Filt <- function(Exp, name, Trans, Fam, Var){
  #'@Exp the expression matrix eg. 'syn8456704'
  #'@name another name file eg. 'syn5879839'
  #'@Trans 'syn3382527'
  #'@Fam a .fam file of the ROS samps syn20820122
  #'@Var 1=Nondiagnosis correcte 0=Diagnosis Corrected
  # eg. Ros_Filt( 'syn8456704', 'syn5879839', 'syn3382527','syn20820122')

  FAM <- as.data.frame(fread( syn_temp$get(Fam)$path))
  Fam <- FAM[ (grepl('ROS', FAM[,1])) == T | (grepl('MAP', FAM[,1])) == T | (grepl('ROS', FAM[,2])) == T | (grepl('MAP', FAM[,2])) == T, ]
  InvFam <- FAM[ (grepl('ROS', FAM[,1])) == F & (grepl('MAP', FAM[,1])) == F & (grepl('ROS', FAM[,2])) == F & (grepl('MAP', FAM[,2])) == F, ]
  Fam[,1] <- Fam[,2]

  #Need to assign sex to MSBB and Mayo
  NAMED_FAM <- rbind( InvFam[1:dim(InvFam)[1], ],  Fam[1:dim(Fam)[1], ] )

  if(Var == 1){
    Exp <- as.data.frame( fread(syn_temp$get(Exp)$path) )
  }else{
    if(Var == 0){
        Exp <- as.data.frame( read.table(syn_temp$get('syn20800652')$path) )
        colnames(Exp) <- gsub('X','', colnames(Exp))
        Exp <- cbind(row.names(Exp), Exp[1:dim(Exp)[2]])
        colnames(Exp)[1] <- 'ensembl_gene_id'
    }else{}
  }
  Trans <- fread(syn_temp$get(Trans)$path)
  Trans <- Trans[ Trans$gwas_id != "", ]
  names <- as.data.frame(fread(syn_temp$get(name)$path))
  Trans <- Trans[ (Trans$gwas_id  %in%  names[,2]) == T, ]
  Trans <- Trans[ (Trans$rnaseq_id == "") == F, ]
  Trans <- Trans[ !duplicated(Trans$rnaseq_id), ]

  ## Z-Scale
  row.names( Exp ) <- Exp$ensembl_gene_id
  Exp <- Exp[ , (colnames( Exp ) %in% "ensembl_gene_id") == F ]
  foo <- as.matrix(Exp)
  foo <- apply(Exp, 1, scale)
  foo <- t(foo)
  colnames(foo) <- colnames(Exp)

  #Filter for matched RNA-Seq and GWAS samples
  Exp <- foo[ , colnames(foo)[ (colnames(foo) %in% Trans$rnaseq_id) == T] ]
  Trans <- as.data.frame( Trans[ (Trans$rnaseq_id %in% colnames(Exp)) == T,  ] )
  row.names(Trans) <- Trans$rnaseq_id

  colnames(Exp) <- c( paste0( Trans[colnames(Exp)[1:length(colnames(Exp))],]$gwas_id, "_",  Trans[colnames(Exp)[1:length(colnames(Exp))],]$gwas_id))
  return(Exp)
}

Norm_Resid_RM <-Ros_Filt( 'syn8456704', 'syn5879839', 'syn3382527','syn20820122', 1)
dNorm_Resid_RM <-Ros_Filt( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122', 0)
```

## Merge RNA-Seq: Merge and Annotate RNA-Seq
* Remove genes on Chr X, Chr Y, Chr M 
* Check for no Patch Contig Nonsense
* Print Sample ID List
* Filter Genotypes for RNA Samples
```{r Merge RNASeq, results='hide', echo=TRUE, cache = FALSE}
#Common Genes:
GNames <- names( table(c( row.names(Norm_Resid_MSBB), row.names(Norm_Resid_MAYO), row.names(Norm_Resid_RM)  ))[ table(c( row.names(Norm_Resid_MSBB), row.names(Norm_Resid_MAYO), row.names(Norm_Resid_RM)  )) == 3] )

#Combine Into One DataFrame
Comb <- function( MSBB, Mayo, RM, Names){
  #'@MSBB The MSBB Expression Set Data Frame
  #'@Mayo The Mayo Expression Set Data Frame
  #'@RM The ROSMAP Expression Set Data Frame
  #'@Names Gene names Vector

  #row.names(MSBB) <- as.character(MSBB$ensembl_gene_id)
  #row.names(Mayo) <- as.character(Mayo$ensembl_gene_id)
  #row.names(RM) <- as.character(RM$ensembl_gene_id)

  MSBB <- MSBB[ GNames ,2:dim(MSBB)[2] ]
  Mayo <- Mayo[ GNames ,2:dim(Mayo)[2] ]
  RM <- RM[ GNames ,2:dim(RM)[2] ]

  Out <- cbind( MSBB[GNames,][ ,1:dim(MSBB)[2] ],
                Mayo[GNames, ][ ,1:dim(Mayo)[2] ],
                RM[GNames, ][, 1:dim(RM)[2] ]
  )

  return( Out )
}

Norm_Resid <- Comb( Norm_Resid_MSBB, Norm_Resid_MAYO, Norm_Resid_RM, GNames )
Diag_Resid <- Comb( dCorrect_Resid_MSBB,  dCorrect_Resid_MAYO, dNorm_Resid_RM, GNames )

#Annotate for TWAS
hg19 <- as.data.frame(fread(syn_temp$get('syn20826610')$path))
hg19 <- hg19[ hg19[,1] %in% row.names( Norm_Resid ), ]
row.names(hg19) <- hg19[,1]

#Filter out the 231 Genes that don't have coherent annotations between GRCH38 and GRCH 37
Norm_Resid <- Norm_Resid[ (row.names(Norm_Resid) %in% row.names(hg19)) == T, ]
Diag_Resid <- Diag_Resid[ (row.names(Diag_Resid) %in% row.names(hg19)) == T, ]

#Finalize
Norm_Resid <- as.data.frame( cbind( 
               TargetID = row.names(Norm_Resid), 
               Gene_Symbol = row.names(Norm_Resid),
               Chr = hg19[row.names(Norm_Resid), ][,2] , 
               Coord = hg19[ row.names(Norm_Resid), ][,3],
               Norm_Resid[ ,1:dim(Norm_Resid)[2]
             ]) )

Diag_Resid <- as.data.frame(cbind( 
               TargetID = row.names(Diag_Resid), 
               Gene_Symbol = row.names(Diag_Resid),
               Chr = hg19[row.names(Diag_Resid), ][,2] , 
               Coord = hg19[ row.names(Diag_Resid), ][,3],
               Diag_Resid[ ,1:dim(Diag_Resid)[2]
             ]) )

#Filter out the ~ ChrX, MT, and Patch genes
Norm_Resid <- Norm_Resid[ (as.character(Norm_Resid$Chr) %in% c('MT', 'X', as.character(Norm_Resid$Chr)[ grepl('PATCH', as.character(Norm_Resid$Chr)) == T])) == F, ]
Diag_Resid <- Diag_Resid[ (as.character(Diag_Resid$Chr) %in% c('MT', 'X', as.character(Diag_Resid$Chr)[ grepl('PATCH', as.character(Diag_Resid$Chr)) == T])) == F, ]

#Order:
Norm_Resid <- Norm_Resid[ with(Norm_Resid, order( Chr, Coord)), ]
Diag_Resid <- Diag_Resid[ with(Diag_Resid, order( Chr, Coord)), ]

#Filter for CEU Ancestry Matched Training Cohort
Tot_Samps <- do.call( rbind, strsplit( colnames(Norm_Resid)[5:length(colnames(Norm_Resid))],  "_" ))
write.table(Tot_Samps, file="ALL_RNASamplesComb.tsv", col.names =F, row.names=F, sep=" ", quote=F)

#Output the ROSMAP INDV CEU Matched Training Cohort
ROS <- colnames(Norm_Resid)[ grepl('ROS', colnames(Norm_Resid)) == T | grepl('MAP', colnames(Norm_Resid)) == T ]
ROS_Norm_Resid <- Norm_Resid[ ,c("TargetID", "Gene_Symbol", "Chr", "Coord", ROS) ]
ROS_Diag_Resid <- Diag_Resid[ ,c("TargetID", "Gene_Symbol", "Chr", "Coord", ROS) ]

ROS_Samps <- do.call( rbind, strsplit( colnames(ROS_Norm_Resid)[5:length(colnames(ROS_Norm_Resid))],  "_" ))
write.table(ROS_Samps, file="ROS_RNASamples.tsv", col.names =F, row.names=F, sep=" ", quote=F)
```

## Fix FamFile
```{r FamFix, results='hide', echo=TRUE, cache=TRUE}
FAM <- as.data.frame(fread( syn_temp$get('syn20820122')$path))
MSBB <-  as.data.frame(fread( syn_temp$get('syn11384608')$path))
MSBB_Clin <- as.data.frame(fread( syn_temp$get('syn6101474')$path))
ROSMAP <- as.data.frame(fread( syn_temp$get('syn3191087')$path))
row.names(ROSMAP) <- ROSMAP$projid
ROSMAP_ID <- as.data.frame(fread( syn_temp$get('syn3382527')$path))
MAYO <- as.data.frame(fread( syn_temp$get('syn19164046')$path))
mayo <- as.data.frame(fread( syn_temp$get('syn19164050')$path))

#MSBB 
row.names(MSBB) <- MSBB$sampleIdentifier
name <- FAM[ FAM$V2 %in% row.names(MSBB) == T, ]$V2
MSBB <- MSBB[name,]

row.names(MSBB_Clin) <- MSBB_Clin$individualIdentifier

MSBB_Clin <- cbind(MSBB_Clin[MSBB$individualIdentifier,1:dim(MSBB_Clin)[2]], MSBB[,1:dim(MSBB)[2] ] )
row.names(MSBB_Clin) <- MSBB_Clin$sampleIdentifier

MSBB_Clin$Diagnosis <- -9
for( i in 1:length(row.names(MSBB_Clin))){
  if( is.na(MSBB_Clin[ i,]$CDR) ==  F & is.na(MSBB_Clin[ i,]$bbscore) ==  F & is.na(MSBB_Clin[ i,]$NP.1) ==  F ){
    if( MSBB_Clin[ i,]$CDR >= 1 & MSBB_Clin[ i,]$bbscore > 3 & MSBB_Clin[ i,]$NP.1 > 1 ){
      MSBB_Clin[ i,]$Diagnosis <- 2
    }else{
      if(MSBB_Clin[ i,]$CDR < 1 & MSBB_Clin[ i,]$bbscore < 4 & MSBB_Clin[ i,]$NP.1 < 2){
        MSBB_Clin[ i,]$Diagnosis <- 1
      }else{}
    }
  }
}

Sex<-c(1,2)
names(Sex) <- c('M', 'F')
MSBB$SEX.inferred <- Sex[MSBB$SEX.inferred]
row.names(FAM)<-FAM$V2
FAM[ row.names(MSBB), ]$V5 <- MSBB$SEX.inferred
FAM[ row.names(MSBB_Clin), ]$V6 <- MSBB_Clin$Diagnosis

#Mayo
mayo <- mayo[complete.cases(mayo$Gender),]
MAYO <- MAYO[complete.cases(MAYO$Sex),]

NewMayo <- as.data.frame( rbind( cbind( MAYO$SampleID,MAYO$Sex,as.character(MAYO$Diagnosis)), cbind( mayo$SampleID,mayo$Gender, as.character(mayo$Diagnosis))) )
NewMayo<- NewMayo[!duplicated(NewMayo),]
NewMayo <- as.data.frame( cbind( as.character(NewMayo$V1),as.character(NewMayo$V2), gsub("_CER","",gsub("_TCX","",NewMayo$V1)), as.character(NewMayo$V3)) )
NewMayo$V2 <- Sex[as.character( NewMayo$V2 )]

names <- FAM[ FAM$V1 == 0 ,]$V2
NewMayo <- NewMayo[ !duplicated(NewMayo$V3), ]
row.names(NewMayo) <- NewMayo$V3
NewMayo <- NewMayo[ names, ]

Aff <- c(-9,-9,1,2)
names(Aff) <- c("Pathologic Aging", "PSP", "Control", "AD")
FAM[ names ,]$V6 <- as.numeric(Aff[as.character(NewMayo[ names, ]$V4)])
FAM[ names ,]$V5 <- NewMayo[ names, ]$V2

#Fix the Final
FAM[ '1957',]$V5 <- 2
FAM[ '1957',]$V6 <- -9
FAM[ '1950',]$V5 <- 2
FAM[ '1950',]$V6 <- -9

#Fix ROSMAP Names
FAM[ grepl('ROS', FAM$V2) == T  | grepl('MAP', FAM$V2) == T , ]$V1 <- FAM[ grepl('ROS', FAM$V2) == T  | grepl('MAP', FAM$V2) == T , ]$V2

#Write New FAM to file
write.table(FAM, file='New_Fam.fam', row.names = F, col.names=F, quote=F, sep=" ")

tail(FAM[ FAM$V1[ grepl('ROS', FAM$V1) == F ] $ FAM$V1[ grepl('MAP', FAM$V1) == F ] ,])
# & FAM$V1[ grepl('ROS', FAM$V1) == F ]
MSBB_List <- FAM[ FAM$V1 != '0', ] 
MSBB_List <- MSBB_List[ MSBB_List$V1[ grepl('MAP', MSBB_List$V1) == F ] ,]
MSBB_List <- MSBB_List[ MSBB_List$V1[ grepl('ROS', MSBB_List$V1) == F ] ,]
write.table( MSBB_List[,1:2], file="MSBB_SamplesToExpand.txt", quote = F, row.names = F, col.names = F, sep = "\t" )
```

## Filter Genotype data
```{bash FilterGenos, results='hide', echo=TRUE, cache=TRUE}
#source /root/.bashrc
export PATH=$PATH:~/TWAS/bin/
CORES=`expr $(nproc) - 2`

#Pull Geno-data
synapse get syn20820120
synapse get syn20820121
synapse get syn20820122

#Copy over FAM File with Sex Info:
mv New_Fam.fam CEU_AncestryMatched_Genotypes.fam

#Extract the MSBB Samples to duplicate and rename to match tissues
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --keep MSBB_SamplesToExpand.txt --recode --out MSBB_Intermediate

#Repeat the rows in the ped file 4 times
awk '{for(i=0;i<4;i++)print}' MSBB_Intermediate.ped > temp

gawk -v m=1 -v n=2 -v el='BM10' 'NR % m == 0 { $n = el } 1' temp > foo
gawk -v m=2 -v n=2 -v el='BM22' 'NR % m == 0 { $n = el } 1' foo > temp
gawk -v m=3 -v n=2 -v el='BM36' 'NR % m == 0 { $n = el } 1' temp > foo
gawk -v m=4 -v n=2 -v el='BM44' 'NR % m == 0 { $n = el } 1' foo > temp
mv temp MSBB_Intermediate.ped
rm foo

plink --threads $CORES --file MSBB_Intermediate --make-bed --out MSBB_Expanded
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --bmerge MSBB_Expanded --make-bed --out Rebuilt
plink --threads $CORES --bfile Rebuilt --make-bed --out CEU_AncestryMatched_Genotypes

rm Rebuilt.*
rm MSBB_Expanded.*

#Extract CEU
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --keep ALL_RNASamplesComb.tsv --silent --make-bed --out All_CEU_ToTrainForTWAS

#Extract ROS_Only:
plink --threads $CORES --bfile CEU_AncestryMatched_Genotypes --keep ROS_RNASamples.tsv --make-bed --out ROSMAPonly_CEU_ToTrainForTWAS
```

```{r sanitycheck, include=TRUE, echo=TRUE}
FAM <- as.data.frame(fread('All_CEU_ToTrainForTWAS.fam'))
row.names(FAM) <- paste0( FAM[,1], '_', FAM[,2] )
table(row.names(FAM) %in% colnames(Diag_Resid))
Diag_Resid_train <- Diag_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]
Norm_Resid_train <- Norm_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]

FAM <- as.data.frame(fread('ROSMAPonly_CEU_ToTrainForTWAS.fam'))
row.names(FAM) <- paste0( FAM[,1], '_', FAM[,2] )
table(row.names(FAM) %in% colnames(ROS_Diag_Resid))
ROS_Diag_Resid_train <- ROS_Diag_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]
ROS_Norm_Resid_train <- ROS_Norm_Resid[ , c("TargetID","Gene_Symbol","Chr","Coord", row.names(FAM) )]

Diag_Resid_train <- Diag_Resid_train[ (Diag_Resid_train$Chr %in% c(1:22)) == T, ]
Norm_Resid_train <- Norm_Resid_train[ (Norm_Resid_train$Chr %in% c(1:22)) == T, ]

ROS_Diag_Resid_train <- ROS_Diag_Resid_train[ (ROS_Diag_Resid_train$Chr %in% c(1:22)) == T, ]
ROS_Norm_Resid_train <- ROS_Norm_Resid_train[ (ROS_Norm_Resid_train$Chr %in% c(1:22)) == T, ]

colnames(Diag_Resid_train) <- c( colnames(Diag_Resid_train[1:4]), do.call( rbind, strsplit( colnames(Diag_Resid_train[ 5:length(colnames(Diag_Resid_train)) ]), "_") )[,2] )

colnames(Norm_Resid_train) <- c( colnames(Norm_Resid_train[1:4]), do.call( rbind, strsplit( colnames(Norm_Resid_train[ 5:length(colnames(Norm_Resid_train)) ]), "_") )[,2] )

colnames(ROS_Diag_Resid_train) <- c( colnames(ROS_Diag_Resid_train[1:4]), do.call( rbind, strsplit( colnames(ROS_Diag_Resid_train[ 5:length(colnames(ROS_Diag_Resid_train)) ]), "_") )[,2] )

colnames(ROS_Norm_Resid_train) <- c( colnames(ROS_Norm_Resid_train[1:4]), do.call( rbind, strsplit( colnames(ROS_Norm_Resid_train[ 5:length(colnames(ROS_Norm_Resid_train)) ]), "_") )[,2] )

```

## Push Results to Synapse -- Contained in ProcessData.Rmd if needed for reference
```{r synapse.parameters.head, include=FALSE, results='hide', echo=FALSE, cache=TRUE}
parentId = 'syn18936948';
activityName = 'Processed TWAS Training Data';
activityDescription = 'Data from processing and hamonizing CEU ancestry matched Genotypes with available RNA-Seq profiles which were aligned to GRCh38 but annotated with hg19 TSS sites here';
thisFileName <- 'ProcessData.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

```{r synapse.parameters, include=FALSE, results='hide', echo=FALSE, cache=TRUE}
activityName = 'Processed TWAS Training Data';
activityDescription = 'Data from processing and hamonizing CEU ancestry matched Genotypes with available RNA-Seq profiles';

CODE <- syn_temp$store(synapseclient$Folder(name = "Processed_TWAS_Training_Data", parentId = parentId))

#Set Used SynIDs For Provenance
Syns_Used <- c("syn5223705", "syn8698211", "syn3817650", "syn8690904", 
"syn6126119", "syn8698214", "syn8449369", "syn6126114", "syn6145639", "syn8690799"
)

# Set annotations
EXP.all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay  = 'RNAseq',
  tissueTypeAbrv  = c('TCX', 'DLPFC', 'PFG', 'STG', 'IFG', 'FP'), 
  study = c('MAYO','MSBB', 'ROSMAP'), 
  organism = 'HomoSapiens',
  consortium  = 'AMPAD',
  normalizationStatus = TRUE,
  normalizationType = 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh38'
)
# Set annotations
Var.all.annotations = list(
  dataType = 'Variant',
  dataSubType = 'SNP',
  summaryLevel = 'Nucleotide',
  assay  = 'Genotype Array',
  tissueTypeAbrv  = c('Whole Blood'), 
  study = c('MAYO','MSBB', 'ROSMAP'), 
  organism = 'HomoSapiens',
  consortium  = 'AMPAD',
  genomeAssemblyID = 'GRCh37'
)

Basic.annotations = list(
  dataType = 'Induvidual List',
  dataSubType = 'Plink PatientID Filter',
  study = c('MAYO','MSBB', 'ROSMAP'), 
  organism = 'HomoSapiens',
  consortium  = 'AMPAD'
)

## TOTAL SVA-only Exp
write.table(Norm_Resid, file="ALL_RNA_Seq_SVAadusted.tsv", col.names =T, row.names=F, sep="\t", quote=F)
Syns_Used_SVA_EXP <- c('syn11384608', 'syn8485026', 'syn6100548', 'syn8456704', 'syn5879839', 'syn3382527','syn20820122', 'syn8466825', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_RNA_Seq_SVAadusted.tsv', name = 'Total Merged and region Averaged RNA-Seq with Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_SVA_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
EXP.all.annotations$dataSubType = 'residualGeneExpForeQTL'
syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

## TOTAL Diagnosis Exp
write.table(Diag_Resid, file="ALL_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv", col.names =T, row.names=F, sep="\t", quote=F)
Syns_Used_Diagnosis_EXP <- c('syn20801991', 'syn11384608', 'syn20801798', 'syn6100548',  'syn20800652', 'syn5879839', 'syn3382527','syn20820122', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv', name = 'Total Merged and region Averaged RNA-Seq with Diagnosis Regressed Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_Diagnosis_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
    EXP.all.annotations$dataSubType = 'residualGeneExpForDiagnosisRegressedeQTL'
    syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

## CEU SVA-only Exp
write.table(Norm_Resid_train, file="ALL_CEU_RNA_Seq_SVAadusted.tsv", col.names =T, row.names=F, sep="\t", quote=F)
Syns_Used_SVA_EXP <- c('syn11384608', 'syn8485026', 'syn6100548', 'syn8456704', 'syn5879839', 'syn3382527','syn20820122', 'syn8466825', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_CEU_RNA_Seq_SVAadusted.tsv', name = 'CEU Merged and region Averaged RNA-Seq with Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_SVA_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

## CEU Diagnosis Exp
write.table(Diag_Resid_train, file="ALL_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv", col.names =T, row.names=F, sep="\t", quote=F)
Syns_Used_Diagnosis_EXP <- c('syn20801991', 'syn11384608', 'syn20801798', 'syn6100548',  'syn20800652', 'syn5879839', 'syn3382527','syn20820122', 'syn20820122')

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ALL_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv', name = 'CEU Merged and region Averaged RNA-Seq with Diagnosis Regressed Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_Diagnosis_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForDiagnosisRegressedeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

#ROSMAP only Expression
  ## CEU SVA-only Exp -- ROSMAP Only
write.table(ROS_Norm_Resid_train, file="ROS_CEU_RNA_Seq_SVAadusted.tsv", col.names =T, row.names=F, sep="\t", quote=F)
Syns_Used_SVA_EXP <- c('syn8456704', 'syn5879839', 'syn3382527','syn20820122' )

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ROS_CEU_RNA_Seq_SVAadusted.tsv', name = 'CEU ROSMAP Only RNA-Seq with Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_SVA_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForeQTL'
  EXP.all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

  ## CEU Diagnosis Exp -- ROSMAP Only
write.table(ROS_Diag_Resid_train, file="ROS_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv", col.names =T, row.names=F, sep="\t", quote=F)
Syns_Used_Diagnosis_EXP <- c( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122')
ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='ROS_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv', name = 'CEU ROSMAP Only RNA-Seq with Diagnosis Regressed Normalised, covariates, surrogate variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used_Diagnosis_EXP, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  EXP.all.annotations$dataSubType = 'residualGeneExpForDiagnosisRegressedeQTL'
  EXP.all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

#ALL INDV
 ##Check that order is correct between file lists
  #table(paste0(Tot_Samps[,1], '_', Tot_Samps[,2]) == colnames(Diag_Resid)[ 5:length(colnames(Diag_Resid))])
Syns <- c( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122')
ENRICH_OBJ_Tot_list <- syn_temp$store( synapseclient$File( path='ALL_RNASamplesComb.tsv', name = 'All Indv with RNA-Seq Profiles in AMP-Ad For Filtering', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  syn_temp$setAnnotations(ENRICH_OBJ_Tot_list, annotations = Basic.annotations)

#ROSMAP INDV
 ##Check that order is correct between file lists
  #table(paste0(ROS_Samps[,1], '_', ROS_Samps[,2]) == colnames(ROS_Norm_Resid)[ 5:length(colnames(ROS_Norm_Resid))])
Syns <- c( 'syn20800652', 'syn5879839', 'syn3382527','syn20820122')
ENRICH_OBJ_RM_List <- syn_temp$store( synapseclient$File( path='ROS_RNASamples.tsv', name = 'ROSMAP Indv with RNA-Seq Profiles in AMP-Ad For Filtering', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  syn_temp$setAnnotations(ENRICH_OBJ_RM_List, annotations = Basic.annotations)

###PLINK ALL -- STORE IN SEPERATE FOLDER INCASE PHI LIMITATIONS APPLY LATER
  CODE <- syn_temp$store(synapseclient$Folder(name = "Ancestry_Clustered_Genotypes", parentId = parentId))
  Syns <- c( 'syn20820120', 'syn20820121', 'syn20820122', ENRICH_OBJ_Tot_list$properties$id  )
#FAM  
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.fam', name = 'All Matched CEU Genotype FAM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BED
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.bed', name = 'All Matched CEU Genotype BED', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#LOG
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.log', name = 'All Matched CEU Genotype LOG', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BIM
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='All_CEU_ToTrainForTWAS.bim', name = 'All Matched CEU Genotype BIM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#PLINK Rosmap
  Syns <- c( 'syn20800652', 'syn5879839', 'syn3382527', ENRICH_OBJ_RM_List$properties$id)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.fam', name = 'ROSMAP Matched CEU Genotype FAM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BED
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.bed', name = 'ROSMAP Matched CEU Genotype BED', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#LOG
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.log', name = 'ROSMAP Matched CEU Genotype LOG', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

#BIM
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAPonly_CEU_ToTrainForTWAS.bim', name = 'ROSMAP Matched CEU Genotype BIM', parentId=CODE$properties$id ), used = Syns, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

```

```{bash Clean, echo=T, results='hide'}
rm -r ~/TWAS/bin/
rm *.bed
rm *.fam
rm *.bim
rm *.log
rm *.tsv
```

### R Source Code
[Github](`r thisFile`)

##Knit to Synapse
```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

library(data.table)

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")

createAndKnitToFolderEntityClient(file = "ProcessData.Rmd",
                                          parentId ="syn18936948",
                                          folderName = 'Processed_TWAS_Training_Data')
```