### 2019-11-25 code/metaAnalysisForDiffExp_NoCBE.Rmd
Fix Meta hg38 NoCBE Plots

---
title: "Meta analysis of reprocessed RNASeq data from AMP-AD (all 6 brain regions)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file
## Load required libraries
library(CovariateAnalysis)
library(data.table)
library(plyr)
library(tidyverse)
#library(synapser)
library(knitr)
library(githubr)
#synapser::synLogin()
library(doParallel)
library(foreach)
cl = makeCluster(detectCores()-2)
registerDoParallel(cl)
options(xtable.type="html")
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, eval=FALSE, cache=TRUE}
parentId = 'syn18936948';
activityName = 'Meta analysis of differential expression (No CBE)';
activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data (all 6 brain regions)';
thisFileName <- 'metaAnalysisForDiffExp_NoCBE.Rmd'
# Github link
thisRepo <- getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

## Data download
### Get study specific covariates and adjusted expression from synapse

```{r get.exp}
# Read covariates
# covar.ids = tibble(study = c("ROSMAP", "MSSM", "MAYO"), 
#                    synid = c("syn8456631", "syn8484996", "syn8466814"), 
#                    version = c("27", "20", "20"))
# covar.md =  covar.ids %>%
#   dplyr::mutate(data = map2(synid, version, 
#                             ~ readr::read_tsv(synGet(.x, version = .y)$path, col_types = cols(.default = "c"))))
# covar = unnest(covar.md, data) 
# Read adjusted expression
covar.ids = c(ROSMAP = 'syn8456631', MSSM = 'syn8484996', MAYO = 'syn8466814')
all.used.ids = as.character(covar.ids)
covar =  covar.ids %>%
  lapply(function(x) readr::read_tsv(syn_temp$get(x)$path, col_types = cols(.default = "c")))
#_#expr.ids = c(ROSMAP = 'syn8456719', MSSM = 'syn8485027', MAYO = 'syn8466826')
expr.ids = c(ROSMAP = 'syn21292041', MSSM = 'syn21285564', MAYO = 'syn21291908')
all.used.ids = c(all.used.ids, as.character(expr.ids))
expr =  expr.ids %>%
  lapply(function(x) readr::read_tsv(syn_temp$get(x)$path, col_types = cols(.default = "c"))) %>%
  plyr::join_all(type = 'inner')
```

```{r modify.covar}
covar$ROSMAP = covar$ROSMAP %>%
  dplyr::select(SampleID, Sex, Diagnosis, APOE4) %>%
  dplyr::mutate(Tissue = 'DLPFC')
covar$MSSM = covar$MSSM %>%
  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
covar$MAYO = covar$MAYO %>%
  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
covar = covar %>%
  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study') %>%
  dplyr::select(-Tissue1)

#Remove CBE
covar <- covar[(covar$Tissue == 'CBE') == F, ]
#row.names(expr) <- expr$ensembl_gene_id
expr <- expr[, c('ensembl_gene_id', covar$SampleID) ]
for(i in 2:dim(expr)[2]){
  expr[,i] <- as.numeric(expr[,i])
}

```

## Perform fixed and random effect meta-analysis for each gene across 7 different brain regions
#### Meta-analysis of AD-CONTROL
```{r ad.cntrl}
# Mean and sd of gene expression in each tissue x diagnosis group
tissue.dx.summary = plyr::ddply(covar, .(Tissue, Diagnosis), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$ensembl_gene_id,
             n = dim(x)[1],
             mn = rowMeans(y[,x$SampleID], na.rm = T),
             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison
meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary, .(ensembl_gene_id), .fun = function(x){
  exp.effect = dplyr::filter(x, Diagnosis == 'AD')
  rownames(exp.effect) = exp.effect$Tissue
  cntrl.effect = dplyr::filter(x, Diagnosis == 'CONTROL')
  rownames(cntrl.effect) = cntrl.effect$Tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
                 studlab = exp.effect$Tissue,
                 sm = 'SMD', method.smd = 'Hedges',
                 method.tau = 'REML')

  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
                   'Q', 'tau', 'H', 'I2')]))
}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
                fdr.random = p.adjust(pval.random, method = 'fdr'))
p = list()
p[[1]] = ggplot(meta.anlz.ad_cntrl, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))+geom_point()
p[[2]] = ggplot(meta.anlz.ad_cntrl, aes(y = -log10(fdr.fixed), x = TE.fixed))+geom_point()+geom_hline(yintercept = -log10(0.05), color = 'red')
p[[3]] = ggplot(meta.anlz.ad_cntrl, aes(y = -log10(fdr.random), x = TE.random))+geom_point()+geom_hline(yintercept = -log10(0.05), color = 'red')
ggpubr::ggarrange(plotlist = p, ncol = 3, nrow = 1)
# Get gene sets
tmp = meta.anlz.ad_cntrl %>%
  dplyr::select(ensembl_gene_id, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
  tidyr::gather(Feature, Value, -ensembl_gene_id) %>%
  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
  tidyr::spread(Feature1, Value) %>%
  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
tmp$Direction[tmp$TE <= 0] = 'DOWN'
tmp$Direction[tmp$TE >= 0] = 'UP'
gs = tmp %>%
  plyr::dlply(.(Model, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })
names(gs) = paste0('ad.control.', names(gs))
all.gs = gs
```

#### Meta-analysis of AD-CONTROL (sex specific)
```{r ad.cntrl.sex}
# Mean and sd of gene expression in each tissue x diagnosis x sex group
tissue.dx.sex.summary = plyr::ddply(covar, .(Tissue, Diagnosis, Sex), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$ensembl_gene_id,
             n = dim(x)[1],
             mn = rowMeans(y[,x$SampleID], na.rm = T),
             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison (sex specific)
meta.anlz.ad_cntrl.sex = plyr::ddply(tissue.dx.sex.summary, .(ensembl_gene_id, Sex), .fun = function(x){
  exp.effect = dplyr::filter(x, Diagnosis == 'AD')
  rownames(exp.effect) = exp.effect$Tissue
  cntrl.effect = dplyr::filter(x, Diagnosis == 'CONTROL')
  rownames(cntrl.effect) = cntrl.effect$Tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
                 studlab = exp.effect$Tissue,
                 sm = 'SMD', method.smd = 'Hedges',
                 method.tau = 'REML')

  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
                   'Q', 'tau', 'H', 'I2')]))
}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
                fdr.random = p.adjust(pval.random, method = 'fdr'))
p = list()
p[[1]] = ggplot(meta.anlz.ad_cntrl.sex, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))
p[[1]] = p[[1]] + geom_point() + facet_grid(.~Sex)
p[[2]] = ggplot(meta.anlz.ad_cntrl.sex, aes(y = -log10(fdr.fixed), x = TE.fixed))
p[[2]] = p[[2]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red') + facet_grid(.~Sex)
p[[3]] = ggplot(meta.anlz.ad_cntrl.sex, aes(y = -log10(fdr.random), x = TE.random))
p[[3]] = p[[3]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red') + facet_grid(.~Sex)
ggpubr::ggarrange(plotlist = p, ncol = 1, nrow = 3)
# Get gene sets
tmp = meta.anlz.ad_cntrl.sex %>%
  dplyr::select(ensembl_gene_id, Sex, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
  tidyr::gather(Feature, Value, -ensembl_gene_id, -Sex) %>%
  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
  tidyr::spread(Feature1, Value) %>%
  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
tmp$Direction[tmp$TE <= 0] = 'DOWN'
tmp$Direction[tmp$TE >= 0] = 'UP'
gs = tmp %>%
  plyr::dlply(.(Sex, Model, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })
names(gs) = paste0('ad.control.', names(gs))
all.gs = c(all.gs, gs)
```

#### Meta-analysis of APOE42-APOE40
```{r apoe2.0}
library(meta)

# Mean and sd of gene expression in each Tissue x APOE4 category
tissue.apoe4.summary = plyr::ddply(covar, .(Tissue, APOE4), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$ensembl_gene_id,
             n = dim(x)[1],
             mn = rowMeans(y[,x$SampleID], na.rm = T),
             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
}, expr)
tissue.apoe4.summary$APOE4 <- as.factor(tissue.apoe4.summary$APOE4)

# Perform meta-analysis for AD-CONTROL comparison (apoe4 2 vs 0)
exp.effect = plyr::ddply(tissue.apoe4.summary, .(ensembl_gene_id), .fun = function(x){
  exp.effect = dplyr::filter(x, APOE4 == '2')
  rownames(exp.effect) = paste0( exp.effect$ensembl_gene_id , "_", exp.effect$Tissue)
  return(EXP=exp.effect)
})
cntrl.effect = plyr::ddply(tissue.apoe4.summary, .(ensembl_gene_id), .fun = function(x){
  cntrl.effect = dplyr::filter(x, APOE4 == '0')
  rownames(cntrl.effect) = paste0( cntrl.effect$ensembl_gene_id , "_", cntrl.effect$Tissue )
  return(CNTRL=cntrl.effect)
}) 
cntrl.effect = cntrl.effect[rownames(exp.effect), ]

START <- seq.int(1, dim(exp.effect)[1], 6)
END <- seq.int(6, dim(exp.effect)[1], 6)
FOO<-as.data.frame(matrix(0, 0, 17))

for( i in 1:length(START) ){ 
  #This gene Doesn't Converge and Breaks the whole damn loop
  if( i %in% c(225)){}
  else{
    tmp = metacont(exp.effect$n[ START[i]:END[i] ], exp.effect$mn[ START[i]:END[i] ], exp.effect$sd[ START[i]:END[i] ], 
                 cntrl.effect$n[ START[i]:END[i] ], cntrl.effect$mn[ START[i]:END[i] ], cntrl.effect$sd[ START[i]:END[i] ],
                 studlab = exp.effect$Tissue[ START[i]:END[i] ],
                 sm = 'SMD', method.smd = 'Hedges',
                 method.tau = 'REML')

    FOO <- rbind(FOO, data.frame(ensembl_gene_id=exp.effect$ensembl_gene_id[START[i]], tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
                   'Q', 'tau', 'H', 'I2')])[1,])
  }
}

meta.anlz.apoe4_2_0 <- as.data.frame(FOO) %>%
  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
                fdr.random = p.adjust(pval.random, method = 'fdr'))

p = list()
p[[1]] = ggplot(meta.anlz.apoe4_2_0, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))
p[[1]] = p[[1]] + geom_point()
p[[2]] = ggplot(meta.anlz.apoe4_2_0, aes(y = -log10(fdr.fixed), x = TE.fixed))
p[[2]] = p[[2]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red')
p[[3]] = ggplot(meta.anlz.apoe4_2_0, aes(y = -log10(fdr.random), x = TE.random))
p[[3]] = p[[3]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red')
ggpubr::ggarrange(plotlist = p, ncol = 3, nrow = 1)
# Get gene sets
tmp = meta.anlz.apoe4_2_0 %>%
  dplyr::select(ensembl_gene_id, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
  tidyr::gather(Feature, Value, -ensembl_gene_id) %>%
  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
  tidyr::spread(Feature1, Value) %>%
  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
tmp$Direction[tmp$TE <= 0] = 'DOWN'
tmp$Direction[tmp$TE >= 0] = 'UP'
gs = tmp %>%
  plyr::dlply(.(Model, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })
names(gs) = paste0('apoe4.2.0.', names(gs))
all.gs = c(all.gs, gs)
```

## Store results in synapse
```{r syn.store, cache = FALSE, eval=FALSE, include=FALSE}
# Create folder to store results in synapse
CODE <- syn_temp$store(synapseclient$Folder(name = "MetaAnalysis", parentId = parentId))
Syns_Used <- c( 'syn8456631', 'syn8484996', 'syn8466814', 'syn21292041', 'syn21285564', 'syn21291908' )

# Write results to files
fwrite(meta.anlz.ad_cntrl, file = 'meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.tsv', name = 'AD-Control meta-analysis across 6 brain regions hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
fwrite(meta.anlz.ad_cntrl.sex, file = 'meta.anlz.ad_cntrl.sex.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.sex.tsv', name = 'AD-Control meta-analysis across 6 brain regions (sex specific) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
fwrite(meta.anlz.apoe4_2_0, file = 'meta.anlz.apoe4_2_0.sex.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.apoe4_2_0.sex.tsv', name = 'AD-Control meta-analysis across 6 brain regions (APOE4) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
save(list = c('all.gs'), file = 'all.gensets.RData')
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='all.gensets.RData', name = 'All gene sets (in RData format) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
```

```{r knit2synapse, eval=FALSE}
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")

parentId = 'syn18936948';
activityName = 'Meta analysis of differential expression (No CBE)';
activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data (all 6 brain regions)';
thisFileName <- 'metaAnalysisForDiffExp_NoCBE.Rmd'
# Github link
thisRepo <- getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))

createAndKnitToFolderEntityClient(file = "metaAnalysisForDiffExp_NoCBE.Rmd",
                                 parentId = "syn18936948",
                                 folderName = "MetaAnalysis_NoCBE")
```

## Source code
[Source code in github](`r thisFile`)