### 2019-12-07 code/GTExPreProcess.Rmd
GTEx Fun

---
title: "Clean GTEx Genotypes for TWAS Expression Imputation"
author: 'JKG'
output: html_notebook
editor_options:
  chunk_output_type: console
---

Date of analysis update: "`r date()`"

| *Synapse ID* | *File Name* |
|  -----------------------------------------  |   ---------                      |
| syn21385083 | | 
| syn20835005 | |
| syn20835007 | |
| syn20834996 | |

```{r libs, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  results='hide')
#library(reticulate)
```
##Before running 
Run the following in docker exec -it <CONTAINER> /bin/bash 
chmod 777 /root/
mv TWAS/ /home/<USR>/TWAS/
chown -hR <GID>:<USR> /home/<USR>/TWAS/
Create a ~/.synapseConfig file in the container, but do not push a container containing your credentials file to Docker Hub!!!

## Login to Synapse Will need to replace with your credentials
```{bash Login, echo=T, results='hide', eval=FALSE}
#source /root/.bashrc
synapse login -u <USR> -p <PSWD> --rememberMe

#Alternativly you can setup a credentials file as such:
touch ~/.synapseConfig
echo "[authentication]" >> ~/.synapseConfig
echo "username = <USR>" >> ~/.synapseConfig
echo "password = <PASWD>" >> ~/.synapseConfig
```

```{bash PLINKSetup, results='hide', eval=F, echo=TRUE}
mkdir ~/TWAS/bin
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink1-assets/plink_linux_x86_64_20190617.zip 
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink_linux_x86_64_20190617.zip
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20191015.zip
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink2_linux_x86_64_20191015.zip

rm ~/TWAS/bin/*.zip
rm ~/TWAS/bin/toy.*
rm ~/TWAS/bin/LICENSE
```

## Pull GTEx Data From Synapse and Prep for Imputation
```{bash Synapse, results='hide', echo=TRUE, cache=TRUE}
synapse get syn21385083
tar -xvf phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1.GRU.tar.gz

rm phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1.GRU.tar.gz
cd phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1/
#tar -xvf phased.vcf.tgz
#tar -xvf info4.maf05.vcf.tgz

#Pull Data with needed SNPS only and fiter GTEx for that SNP List
synapse get syn20835005
synapse get syn20835007
synapse get syn20834996

~/TWAS/bin/plink --threads 10 --bfile All_CEU_ToTrainForTWAS --write-snplist

tar -xvf all.SNPs.tgz
touch merge.txt
touch Log.txt
for i in {22..1}
  do
    tar -xvf chr$i.all.tgz

    rm chr$i.dosage
    rm chr$i.tped
    rm data_imputed.chr$i.gen
    rm data_imputed.chr$i.info
    rm chr$i.info4.exclusion.snplist.txt
    rm chr$i.info4.maf05.exclusion.snplist.txt

    vcftools --vcf chr$i.vcf --plink --out chr$i

    echo $(ls -ltr chr$i.vcf) >> Log.txt
    rm chr$i.vcf

    ~/TWAS/bin/plink --threads 10 --file chr$i --extract plink.snplist -make-bed --out chr$i\_filt
    rm chr$i.ped
    rm chr$i.map
    rm chr$i.log

    echo "chr"$i"_filt.bed chr"$i"_filt.bim chr"$i"_filt.fam" >> merge.txt
  done
grep -v 'chr1_filt.bed' merge.txt > foo.txt
mv foo.txt merge.txt

~/TWAS/bin/plink --threads 10 --bfile chr1_filt --merge-list merge.txt --make-bed --out GTEx_FiltSNP_Data
rm merge.txt
rm Log.txt 
rm chr*_filt.*

#Move files up to code directory and then clean up 
mv GTEx_FiltSNP_Data.* ../
mv All_CEU_ToTrainForTWAS.* ../
mv plink.snplist ../SNPs_for_Impute.snplist
cd ..

rm -r phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1/
```

## Push Results to Synapse -- Contained in ProcessData.Rmd if needed for reference
```{r synapse.parameters.head, include=FALSE, results='hide', echo=FALSE, cache=TRUE}
parentId = 'syn18936948';
activityName = 'Processed GTEx Genotypes TWAS';
activityDescription = 'Proccessed and Filtered GTEx Genotype Data';
thisFileName <- 'GTExPreProcess.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

```{r synapse.parameters, include=FALSE, results='hide', echo=FALSE, cache=TRUE}

activityName = 'Processed GTEx Genotypes TWAS';
activityDescription = 'Proccessed and Filtered GTEx Genotype Data';

CODE <- syn_temp$store(synapseclient$Folder(name = "GTExGenotypes", parentId = parentId))

#Set Used SynIDs For Provenance
Syns_Used <- c("syn21385083", "syn20835005", "syn20835007", "syn20834996")

# Set annotations
Var.all.annotations = list(
  dataType = 'Variant',
  dataSubType = 'SNP',
  summaryLevel = 'Nucleotide',
  assay  = 'Genotype Array',
  tissueTypeAbrv  = c('Mixed Tissues'), 
  study = 'GTEx', 
  organism = 'HomoSapiens',
  consortium  = 'GTEx',
  genomeAssemblyID = 'GRCh37'
)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GTEx_FiltSNP_Data.fam', name = 'Filtered GTEx Genotype FAM file for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GTEx_FiltSNP_Data.bim', name = 'Filtered GTEx Genotype BIM file for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GTEx_FiltSNP_Data.bed', name = 'Filtered GTEx Genotype BED file for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

```

```{bash CleanClean, results='hide', echo=TRUE, cache=TRUE}
rm *.bed
rm *.fam
rm *.bim
rm *.nosex
rm *.log
rm *.snplist
```

### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")

createAndKnitToFolderEntityClient(file = "GTExPreProcess.Rmd",
                                          parentId ="syn18936948",
                                          folderName = 'GTExPreProcess')

```