### 2019-12-11 code/GTExPreProcess.Rmd
GTEx Processing

---
title: "Clean GTEx Genotypes for TWAS Expression Imputation"
author: 'JKG'
output: html_notebook
editor_options:
  chunk_output_type: console
---

Date of analysis update: "`r date()`"

| *Synapse ID* | *File Name* |
|  -----------------------------------------  |   ---------                      |
| syn21385083 | | 
| syn20835005 | |
| syn20835007 | |
| syn20834996 | |

```{r libs, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  results='hide')
library(reticulate)
#BiocManager::install("Rgraphviz")
#BiocManager::install("graph")
#install.packages('CePa')
#BiocManager::install("phantasus")
```
##Before running 
Run the following in docker exec -it <CONTAINER> /bin/bash 
chmod 777 /root/
mv TWAS/ /home/<USR>/TWAS/
chown -hR <GID>:<USR> /home/<USR>/TWAS/
Create a ~/.synapseConfig file in the container, but do not push a container containing your credentials file to Docker Hub!!!

## Login to Synapse Will need to replace with your credentials
```{bash Login, echo=T, results='hide', eval=FALSE}
#source /root/.bashrc
synapse login -u <USR> -p <PSWD> --rememberMe

#Alternativly you can setup a credentials file as such:
touch ~/.synapseConfig
echo "[authentication]" >> ~/.synapseConfig
echo "username = <USR>" >> ~/.synapseConfig
echo "password = <PASWD>" >> ~/.synapseConfig
```

```{bash PLINKSetup, results='hide', eval=F, echo=TRUE}
mkdir ~/TWAS/bin
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink1-assets/plink_linux_x86_64_20190617.zip 
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink_linux_x86_64_20190617.zip
wget -q -P ~/TWAS/bin http://s3.amazonaws.com/plink2-assets/plink2_linux_x86_64_20191015.zip
unzip -d ~/TWAS/bin/ ~/TWAS/bin/plink2_linux_x86_64_20191015.zip

rm ~/TWAS/bin/*.zip
rm ~/TWAS/bin/toy.*
rm ~/TWAS/bin/LICENSE
```

## Pull GTEx Data From Synapse and Prep for Imputation
```{bash Synapse, results='hide', echo=TRUE, cache=TRUE}
synapse get syn21385083
tar -xvf phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1.GRU.tar.gz

rm phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1.GRU.tar.gz
cd phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1/
#tar -xvf phased.vcf.tgz
#tar -xvf info4.maf05.vcf.tgz

#Pull Data with needed SNPS only and fiter GTEx for that SNP List
synapse get syn20835005
synapse get syn20835007
synapse get syn20834996

~/TWAS/bin/plink --threads 12 --bfile All_CEU_ToTrainForTWAS --write-snplist

tar -xvf all.SNPs.tgz
touch merge.txt
touch Log.txt
for i in {22..1}
  do
    tar -xvf chr$i.all.tgz

    rm chr$i.dosage
    rm chr$i.tped
    rm data_imputed.chr$i.gen
    rm data_imputed.chr$i.info
    rm chr$i.info4.exclusion.snplist.txt
    rm chr$i.info4.maf05.exclusion.snplist.txt

    vcftools --vcf chr$i.vcf --plink --out chr$i

    echo $(ls -ltr chr$i.vcf) >> Log.txt
    rm chr$i.vcf

    ~/TWAS/bin/plink --threads 12 --file chr$i --extract plink.snplist -make-bed --out chr$i\_filt
    rm chr$i.ped
    rm chr$i.map
    rm chr$i.log

    echo "chr"$i"_filt.bed chr"$i"_filt.bim chr"$i"_filt.fam" >> merge.txt
  done
grep -v 'chr1_filt.bed' merge.txt > foo.txt
mv foo.txt merge.txt

~/TWAS/bin/plink --threads 12 --bfile chr1_filt --merge-list merge.txt --make-bed --out GTEx_FiltSNP_Data
rm merge.txt
rm Log.txt 
rm chr*_filt.*

#Move files up to code directory and then clean up 
mv GTEx_FiltSNP_Data.* ../
mv All_CEU_ToTrainForTWAS.* ../
mv plink.snplist ../SNPs_for_Impute.snplist
cd ..

rm -r phg000219.v4.GTEx_Pilot_Imputation.genotype-imputed-data.c1/
```

## Pull GTEx WGS Data From Synapse and Prep for Imputation
This will need a bunch of space!
```{bash Pull_WGS, results='hide', echo=TRUE, cache=TRUE}
synapse get syn21417881
gunzip GTEx_Analysis_2016-01-15_v7_WholeGenomeSeq_635Ind_PASS_AB02_GQ20_HETX_MISS15_PLINKQC.vcf.gz

synapse get syn20835007

awk '{ print $1"\t"$4 }' All_CEU_ToTrainForTWAS.bim > Positions.txt

vcftools --gzvcf GTEx_Analysis_2016-01-15_v7_WholeGenomeSeq_635Ind_PASS_AB02_GQ20_HETX_MISS15_PLINKQC.vcf.gz --positions Positions.txt --plink --out Init_Filtered

~/TWAS/bin/plink --threads 12 --file Init_Filtered --make-bed --out temp

```

### ReName the MSBB SNPS so the can be combined
```{python ReNamer_ONE, echo=F, results='hide'}
import sys
import os
import re

InFile = 'All_CEU_ToTrainForTWAS.bim'
InFileb = 'temp.bim' #LDREF_FiltSNPs.bim

RS_IDs = {}
RS_Pos = {}

F1 = open(InFile)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')
	RS_IDs[ ''.join([lst[0], '_', lst[3], ':', lst[4], '-', lst[5]] ) ] = lst[1]
	RS_IDs[ ''.join([lst[0], '_', lst[3], ':', lst[5], '-', lst[4]] ) ] = lst[1]
	#RS_IDs[ ''.join([lst[0], '_', lst[3]] ) ] = lst[1]
	#RS_Pos[ ''.join([lst[0], '_', lst[3], ':', lst[4], '-', lst[5]] ) ] = lst[2]
F1.close()

#Open Output
OUT = open( ''.join([ InFileb.replace(".bim", ""), 'Renamed.bim']), 'w' )
#OUT2 = open( ''.join([ InFileb.replace(".bim", ""), 'WTF.bim']), 'w' )
F1 = open(InFileb)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')
	foo = ''.join([lst[0], '_', lst[3], ':', lst[5], '-', lst[4]] )
	#foo = ''.join([lst[0], '_', lst[3] ] ) 
	if foo in RS_IDs:
	  Name = RS_IDs[  ''.join([lst[0], '_', lst[3], ':', lst[4], '-', lst[5] ] ) ]
	  #Name = RS_IDs[  ''.join([lst[0], '_', lst[3] ] ) ]
	  Entry = '\t'.join([ lst[0], Name, lst[2], lst[3], lst[4], lst[5] ])
	  print >>OUT, Entry
	else:
	  print >>OUT, Entry

F1.close()
OUT.close()
#OUT2.close()
```

```{bash FinalizeGenoType, results='hide', echo=TRUE, cache=TRUE}
mv tempRenamed.bim temp.bim

#Remove all Repeated SNPs 9I have no Idea how imputation will handle them
~/TWAS/bin/plink --threads 12 --bfile All_CEU_ToTrainForTWAS --write-snplist
~/TWAS/bin/plink --threads 12 --bfile temp --extract plink.snplist --list-duplicate-vars --make-bed --out Test
~/TWAS/bin/plink --threads 12 --bfile Test --list-duplicate-vars suppress-first --out Duplicates
~/TWAS/bin/plink --threads 12 --bfile Test --exclude Duplicates.dupvar --make-bed --out UniqueGTEx_GWAS_ForImpute

rm Test*
rm *.snplist
rm temp.*
rm Init_Filtered.*
rm Duplicates.*
rm positions.txt

## Pull MetaData & Expression
#_# wget https://storage.googleapis.com/gtex_analysis_v8/multi_tissue_eqtl_data/GTEx_Analysis_v8.metasoft.txt.gz
#_# gunzip GTEx_Analysis_v8.metasoft.txt.gz

wget https://storage.googleapis.com/gtex_analysis_v8/annotations/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt

wget https://storage.googleapis.com/gtex_analysis_v8/rna_seq_data/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_tpm.gct.gz
gunzip GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_tpm.gct.gz
```

```{r GTExPheno, cache=FALSE, echo=FALSE}
#ID Brain INDV
Meta <- as.data.frame( data.table::fread(file='GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt', header=T) )
table(Meta$SMTS)

TissueTypes <- names( table(Meta[ Meta$SMTS == 'Brain', ]$SMTSD) )
TissueTypes <- TissueTypes[ grepl( "Cerebell", TissueTypes ) == F ]

BrainMeta <- Meta[ Meta$SMTSD %in% TissueTypes, ]
IDs <- BrainMeta$SAMPID

GTEX_IDs <- paste0( do.call( rbind, strsplit( IDs,'-' ))[,1], '-', do.call( rbind, strsplit( IDs,'-' ))[,2] )
uGTEX_IDs <-GTEX_IDs[ !duplicated(GTEX_IDs) ]

FAM <- as.data.frame( data.table::fread(file='UniqueGTEx_GWAS_ForImpute.fam', header=F) )
row.names(FAM) <- FAM$V1

BRAIN_INDV <- FAM[ (FAM$V1 %in% uGTEX_IDs) == T, 1:2 ]
write.table( BRAIN_INDV, file = "GTEX_Plink_Indvs.txt", row.names = F, col.names = F, sep='\t', quote=F)

####################################################################################################
#Get Expression
EXP <- phantasus::read.gct('GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_tpm.gct')

exp <- EXP@assayData$exprs
bMeta <- Meta[ Meta$SAMPID %in% IDs, ]
bexp <- exp[ , (colnames(exp) %in% IDs) ==T ] 
row.names( bMeta ) <- bMeta$SAMPID
bMeta <- bMeta[ colnames(bexp), ]

#Filter for relevant genes
row.names(bexp) <- do.call( rbind, strsplit(row.names(bexp), '\\.'))[,1]
Genes <- as.character(as.data.frame(data.table::fread( syn_temp$get('syn21011180')$path, header = T))[,1])
bexp <- bexp[ Genes[ Genes %in% row.names(bexp) ], ]

#Scale the values
ScalerTable <- table( bMeta$SAMPID, bMeta$SMTSD)
ScalerTable <- as.data.frame(ScalerTable)

Scaled_bexp <- as.data.frame( matrix(0,dim(bexp)[1],0) )
row.names( Scaled_bexp ) <- row.names(bexp)

Scaled_bexp_Log2 <- as.data.frame( matrix(0,dim(bexp)[1],0) )
row.names( Scaled_bexp_Log2 ) <- row.names(bexp)

for( Type in names(table(ScalerTable$Var2)) ){
  #print(head( bexp[ ,as.character(ScalerTable[ ScalerTable$Var2 == Type & ScalerTable$Freq == 1,  ]$Var1) ][,1:5]))
  Scal <- t(scale(t(bexp[ ,as.character(ScalerTable[ ScalerTable$Var2 == Type & ScalerTable$Freq == 1,  ]$Var1)] ))) 
  Scaled_bexp <- cbind(Scaled_bexp, Scal)

  Scal_log <- t(scale(t(log2(bexp[ ,as.character(ScalerTable[ ScalerTable$Var2 == Type & ScalerTable$Freq == 1,  ]$Var1) ]+1)))) 
  Scaled_bexp_Log2 <- cbind(Scaled_bexp_Log2, Scal_log)
}

hist(as.numeric(Scaled_bexp['ENSG00000225630',]), breaks =80)
hist(as.numeric(Scaled_bexp_Log2['ENSG00000225630',]), breaks =80)

write.table(Scaled_bexp, file = "Init_GETx_Scaled_Expression.tsv", sep="\t", row.names = T, col.names = T, quote = F)
write.table(Scaled_bexp_Log2, file = "Init_GETx_ScaledLog2Plus1_Expression.tsv", sep="\t", row.names = T, col.names = T, quote = F)

bMeta$PTID <- paste0( do.call( rbind, strsplit( colnames(Scaled_bexp), '-'))[,1], '-', do.call( rbind, strsplit( colnames(Scaled_bexp), '-'))[,2] )
write.table(bMeta, file = "GETx_BrainMeta.tsv", sep="\t", row.names = T, col.names = T, quote = F)

table(FAM$V1 %in% bMeta$PTID)

```

## Push Results to Synapse -- Contained in ProcessData.Rmd if needed for reference
```{r synapse.parameters.head, include=FALSE, results='hide', echo=FALSE, cache=TRUE}
parentId = 'syn18936948';
activityName = 'Processed GTEx Genotypes TWAS';
activityDescription = 'Proccessed and Filtered GTEx Genotype Data';
thisFileName <- 'GTExPreProcess.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

```{r synapse.parameters, include=FALSE, results='hide', echo=FALSE, cache=TRUE}

activityName = 'Processed GTEx Genotypes TWAS';
activityDescription = 'Proccessed and Filtered GTEx Genotype Data';

CODE <- syn_temp$store(synapseclient$Folder(name = "GTExGenotypes", parentId = parentId))
CODEalt <- syn_temp$store(synapseclient$Folder(name = "GTExPreProcess", parentId = parentId))

#Set Used SynIDs For Provenance
Syns_Used <- c("syn21385083", "syn20835005", "syn20835007", "syn20834996")

# Set annotations
Var.all.annotations = list(
  dataType = 'Variant',
  dataSubType = 'SNP',
  summaryLevel = 'Nucleotide',
  assay  = 'Genotype Array',
  tissueTypeAbrv  = c('Mixed Tissues'), 
  study = 'GTEx', 
  organism = 'HomoSapiens',
  consortium  = 'GTEx',
  genomeAssemblyID = 'GRCh37'
)

# Set annotations
EXP.all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay  = 'RNAseq',
  tissueTypeAbrv  = c(gsub( 'Brain - ', "", names(table(bMeta$SMTSD)))), 
  study = 'GTEx', 
  organism = 'HomoSapiens',
  consortium  = 'GTEx',
  normalizationStatus = TRUE,
  normalizationType = 'Scaled TPM',
  rnaquantification = 'UNK',
  genomeAssemblyID = 'GRCh38'
)

Basic.annotations = list(
  dataType = 'MetaData',
  dataSubType = 'GTEx Sample Meta Data',
  study = 'GTEx', 
  organism = 'HomoSapiens',
  consortium  = 'GTEx'
)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GTEx_FiltSNP_Data.fam', name = 'Filtered GTEx Genotype FAM file for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GTEx_FiltSNP_Data.bim', name = 'Filtered GTEx Genotype BIM file for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GTEx_FiltSNP_Data.bed', name = 'Filtered GTEx Genotype BED file for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='UniqueGTEx_GWAS_ForImpute.fam', name = 'Filtered GTEx Genotype FAM file from GWAS Data for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='UniqueGTEx_GWAS_ForImpute.bed', name = 'Filtered GTEx Genotype BIM file from GWAS Data for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='UniqueGTEx_GWAS_ForImpute.bim', name = 'Filtered GTEx Genotype BED file from GWAS Data for TWAS Imputation', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Var.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='Init_GETx_Scaled_Expression.tsv', name = 'TPM GTEx Brain Expression for all Profiles Z-Scaled within Tissue', parentId=CODEalt$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='Init_GETx_ScaledLog2Plus1_Expression.tsv', name = 'Log2 TPM+1 GTEx Brain Expression for all Profiles Z-Scaled within Tissue', parentId=CODEalt$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = EXP.all.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GETx_BrainMeta.tsv', name = 'Brain Sample Meta Data', parentId=CODEalt$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
syn_temp$setAnnotations(ENRICH_OBJ, annotations = Basic.annotations)

ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='GTEX_Plink_Indvs.txt', name = 'Genotype IIDs with a BrainExpression Profile', parentId=CODEalt$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
```

```{bash CleanClean, results='hide', echo=TRUE, cache=TRUE}
rm *.bed
rm *.fam
rm *.bim
rm *.nosex
rm *.log
rm *.snplist
rm *.tsv
rm *.txt
rm *.gct
rm *.gz
```

### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")

createAndKnitToFolderEntityClient(file = "GTExPreProcess.Rmd",
                                          parentId ="syn18936948",
                                          folderName = 'GTExPreProcess')

```