### 2020-02-07 code/SMR.Rmd

---
  title: "Summary-data-based Mendelian Randomization" author: "Jake Gockley" date: "01/17/2019" header-includes:
  - \usepackage{multicol} - \newcommand{\btwocol}{\begin{multicols}{2}} - \newcommand{\etwocol}{\end{multicols}} output:
  html_document: default pdf_document: default ---

```{r setup, include=FALSE} knitr::opts_chunk$set(echo = F) library(readr) library(snpStats) library(sem) ``` ```{r code, eval=TRUE, cache=FALSE, 
include=TRUE}
#Pull eQTL Data
system( 'synapse get syn16984815' )
#Pull Weights
system('aws s3 cp s3://jkg-s3-synapseencryptedexternalbucket-zszdd03ghnb2/ROSMAP/TWAS_2019_12_02_PanCortex_DiagnosisRegressed_Weights/ All_Data/ALL/ 
--recursive') system('rm All_Data/ALL/*.hsq')
#PullGenos
system('synapse get syn20820120 --downloadLocation Binary_Cleaned/') system('synapse get syn20820121 --downloadLocation Binary_Cleaned/') system('synapse 
get syn20820122 --downloadLocation Binary_Cleaned/')
#GWAS
system('synapse get syn18999109 --downloadLocation .')
#Expression
system('synapse get syn21011186 --downloadLocation .') Exp <- read.table('ALL_CEU_RNA_Seq_SVAadusted_DiagnosisRegressed.tsv', header = T, sep='\t')
#LoadKunkle
Kunk <- read.table('Kunkle_etal_2019_IGAP_Summary_statistics.with_allelefreqs.txt', header =T, sep="")
#Try the MR for one gene
load("All_Data/ALL/ENSG00000130208.wgt.RDat") PLINK <- read.plink(bed = 'Binary_Cleaned/CEU_AncestryMatched_Genotypes.bed' ,
                    bim = 'Binary_Cleaned/CEU_AncestryMatched_Genotypes.bim',
                    fam = 'Binary_Cleaned/CEU_AncestryMatched_Genotypes.fam' )
#Filter Kunkle for relevant SNPS by their IDS
K <- Kunk[ Kunk$MarkerName %in% PLINK$map$snp.name, ] row.names(K) <- K$MarkerName mygwas<-K[,c(3:5,9,6:8)] colnames(mygwas) <- c("SNP", "A1", 'A2', "freq", 
"b", "se", "p" ) mygwas$n <- rep( 2003, dim(mygwas)[1] ) write.table(snps$V2, 'snp.list', row.names=F, col.names=F, quote=F, sep='\t') write.table(mygwas, 
'mygwas.ma', row.names=F, col.names=T, quote=F, sep='\t') system( 'plink -bfile Binary_Cleaned/CEU_AncestryMatched_Genotypes --extract snp.list --make-bed 
--out Temp_plink') system( './smr_Linux --bfile Temp_plink --gwas-summary mygwas.ma --beqtl-summary myeqtl --out mymulti --smr-multi')
#make the eqtl data into BESD format
EQTL <- read.csv( 'Cortex_MetaAnalysis_ROSMAP_CMC_HBCC_Mayo_cis_eQTL_release.csv', header = T ) EQTl <- EQTL[ as.character(EQTL$snpid ) %in% 
as.character(PLINK$map$snp.name) , ] EQtl <- EQTl[ as.character(EQTl$gene ) %in% as.character(Exp$TargetID) , ] EQTL_sink <- EQTL EQTL <- EQtl
#SNP gene beta t-stat p-value FDR
EQTL$geneSymbol <- as.character(EQTL$geneSymbol) EQTL$gene <- as.character(EQTL$gene) EQTL[ as.character(EQTL$geneSymbol) =="",]$geneSymbol <- EQTL[ 
as.character(EQTL$geneSymbol) =="",]$gene EQTL$snpid <- as.character(EQTL$snpid) EQTL$DUP <- paste0( as.character(EQTL$snpid), '_', as.character(EQTL$gene)) 
EQTL_deDup <- EQTL[ !duplicated(EQTL$DUP), ]
##LOAD EXPRESSION AND BIM
###FILTER FOR RELEVANT SNP-GENEs
wExp <- read.table(file = 'WeightList/MaterWeight.lst', sep = "", header = T)
#EQTL_deDup <- EQTL_deDup[ EQTL_deDup$snpid %in% as.character(PLINK$bim[,2]), ] EQTL_deDup <- EQTL_deDup[ EQTL_deDup$gene %in% as.character(Exp[,2]), ]
Repo <- function( n ){
  #'@Mat eg. EQTL_deDup '@n eg a number
  if( EQTL_deDup[n,]$geneSymbol == ""){
    EQTL_deDup[n,]$geneSymbol <- EQTL_deDup[n,]$gene
  }else{}
}
REPS <- row.names( EQTL_deDup[ EQTL_deDup$geneSymbol == "", ] ) eqtl <- EQTL_deDup[ , c('snpid', 'geneSymbol', 'beta', 'statistic', 'pvalue', 'beta') ] 
colnames(eqtl) <- c( 'SNP', 'gene', 'beta', 't-stat', 'p-value', 'FDR' ) write.table( eqtl, file = 'mateQTL_geneN.txt', col.names = T, row.names=F, quote=F, 
sep='\t' ) Eeqtl <- EQTL_deDup[ , c('snpid', 'gene', 'beta', 'statistic', 'pvalue', 'beta') ] colnames(Eeqtl) <- c( 'SNP', 'gene', 'beta', 't-stat', 
'p-value', 'FDR' ) write.table( Eeqtl, file = 'mateQTL_ensgN.txt', col.names = T, row.names=F, quote=F, sep='\t' ) system( './smr_Linux --eqtl-summary 
mateQTL_ensgN.txt --matrix-eqtl-format --make-besd --out mybesd_ensgN' ) system( './smr_Linux --eqtl-summary mateQTL_geneN.txt --matrix-eqtl-format 
--make-besd --out mybesd_geneN' )
##Fix the ,esi and .epi files
AleleSelect <- function( Row ){
  if( Row[4] > 0 ){
    Eff <- as.character(Row[9])
    Unaff <- as.character(c(Row[6],Row[7])[ (c(Row[6],Row[7]) %in% Eff) == F ])
  }else{
    Unaff <- as.character(Row[9])
    Eff <- as.character(c(Row[6],Row[7])[ (c(Row[6],Row[7]) %in% Unaff) == F ])
  }
  if(Eff == Row[7]){
    Freq <- Row[8]
  }else{
    Freq <- 1-as.numeric(Row[8])
  }

  return( c(as.character(Eff),as.character(Unaff),as.character(Freq)) )
}
ESiProcess <- function( Fil ){
  #'@Fil the loaded .esi file (ie File)
  ESi_info <- EQTL_deDup[ !duplicated(EQTL_deDup$snpid), c("chromosome", "snpLocation", "snpid", "statistic", "beta", "A1", "A2", "A2freq", 
"expressionIncreasingAllele")]
  row.names(ESi_info) <- ESi_info$snpid
  Fil$V1 <- as.character(ESi_info[ as.character( Fil$V2 ), ]$chromosome)
  Fil$V4 <- as.character(ESi_info[ as.character( Fil$V2 ), ]$snpLocation)

  ESi_info$A1 <- as.character(ESi_info$A1)
  ESi_info$A2 <- as.character(ESi_info$A2)
  ESi_info$expressionIncreasingAllele <- as.character(ESi_info$expressionIncreasingAllele)

  temp <- t( apply(ESi_info, 1, AleleSelect) )
  Fil[,5:7] <- temp[ as.character(Fil$V2),]
  return(Fil)
}
File <- read.table(file='mybesd_ensgN.esi', header=F, sep='\t') File2 <- ESiProcess( File ) write.table(File2, file='mybesd_ensgN.esi', col.names=F, 
row.names = F, quote=F, sep='\t') File <- read.table(file='mybesd_geneN.esi', header=F, sep='\t') File2 <- ESiProcess( File ) write.table(File2, 
file='mybesd_geneN.esi', col.names=F, row.names = F, quote=F, sep='\t')
#ESPI
row.names(Exp) <- as.character(Exp$TargetID) Exp$TargetID <- as.character(Exp$TargetID) Exp$Gene_Symbol <- as.character(Exp$Gene_Symbol) EQTL_deDup$gene <- 
as.character(EQTL_deDup$gene) EQTL_deDup$geneSymbol <- as.character(EQTL_deDup$geneSymbol) EQTL_deDup$chromosome <- as.character(EQTL_deDup$chromosome) 
EQTL_deDup$strand <- as.character(EQTL_deDup$strand)
#Split EQTL_deDup into special data frames to reference
EQTL_deDup_GN <- EQTL_deDup[ !duplicated(EQTL_deDup$geneSymbol), ] EQTL_deDup_ENSG <- EQTL_deDup[ !duplicated(EQTL_deDup$gene), ]

ESpProcess <- function( Name ){
  #'@Name row with gene name or ENSG in column 2
  if(Name[2] %in% EQTL_deDup_ENSG$gene){
    Row <- EQTL_deDup_ENSG[ EQTL_deDup_ENSG$gene %in% Name[2], ][1,]
  }else{
    Row <- EQTL_deDup_GN[ EQTL_deDup_GN$geneSymbol %in% Name[2], ][1,]
  }
  st <- c('+', '-')
  names(st) <- c("1","-1")

  return( c( as.character(Row[1]),
             as.character(Name[2]),
             as.character("0"),
             as.character(Exp[ as.character(Row[5]), 4]),
             as.character(Name[2]),
             as.character(st[Row[,15]])
             )
          )
}
File <- read.table(file='mybesd_geneN.epi', header=F, sep='\t') File$V2 <- as.character(File$V2) foo <- t( apply(File, 1, ESpProcess ) ) write.table( foo, 
file='mybesd_geneN.epi', row.names=F, col.names=F, quote = F, sep='\t' ) File <- read.table(file='mybesd_ensgN.epi', header=F, sep='\t') File$V2 <- 
as.character(File$V2) foo <- t( apply(File, 1, ESpProcess ) ) write.table( foo, file='mybesd_ensgN.epi', row.names=F, col.names=F, quote = F, sep='\t' ) 
Genes <- c("APOC1", "EED", "CD2AP", "CEACAM19", "CLPTM1", "MTCH2", "TREM2", "KNOP1", "POLR2E", "POLR2G", "ZNF660", "MADD", "PVRL2", "PVR", "TMEM223", 
"NUP160", "PTPRJ", "C1QTNF4") ENSG <- c write.table(EQTL_deDup_GN[ EQTL_deDup_GN$geneSymbol %in% Genes, ]$gene, "myprobe.list", row.names=F, col.names=F, 
quote=F, sep='\t') library(doParallel) library(foreach)
#Start Master File
system( 'echo "probeID ProbeChr Gene Probe_bp topSNP topSNP_chr topSNP_bp A1 A2 Freq b_GWAS se_GWAS tp_GWAS b_eQTL se_eQTL p_eQTL b_SMR se_SMR p_SMR 
p_SMR_multi p_HEIDI nsnp_HEIDI" > Compiled_SMR.tsv') List <- read.table( 'WeightList/MaterWeight.lst', header = T, sep='')
#i <- 1
Runner <- function( i ){
  #Load Gene Info
  load(as.character( List$WGT )[i])
  Probe <- as.character( List$ID )[i]
  SNPs <- snps$V2
  #Save Temp Files
  write.table(Probe, paste0("myprobe_", Probe, ".list"), row.names=F, col.names=F, sep="\t", quote=F)
  write.table(SNPs, paste0("mysnp_", Probe, ".list"), row.names=F, col.names=F, sep="\t", quote=F)
  #Run Model
  system( paste0("./smr_Linux --bfile Binary_Cleaned/CEU_AncestryMatched_Genotypes --gwas-summary mygwas.ma --thread-num 1 --beqtl-summary mybesd_ensgN 
--ld-upper-limit 1 --out temp_Ensg_", Probe, " --extract-snp mysnp_", Probe, ".list --extract-probe myprobe_", Probe, ".list --peqtl-heidi 0 --ld-multi-snp 
1 --smr-multi --peqtl-smr 1 --set-wind 2000 "))
  #Pull Info
  system( paste0("grep -v 'probeID' temp_Ensg_", Probe, ".msmr >> Compiled_SMR.tsv") )
  #remove Temp Info
  system( paste0("rm myprobe_", Probe, ".list") )
  system( paste0("rm mysnp_", Probe, ".list") )
  system( paste0("rm temp_Ensg_", Probe, ".*") )
}
cl <- makeCluster( detectCores()-2 ) registerDoParallel(cl) foo <- Sys.time()
  foreach(i=1:dim(List)[1]) %dopar% Runner( i ) Sys.time()-foo stopCluster(cl) Table<-read.table(file='Compiled_SMR.tsv', sep='\t', header=T) system( 'awk 
-F \',\' \'{ print $5"\t"$6 }\' Cortex_MetaAnalysis_ROSMAP_CMC_HBCC_Mayo_cis_eQTL_release.csv | sort | uniq | grep -v \'geneSymbol\' > Translate.txt') Trans 
<- read.table( file='Translate.txt', sep ="\t", header=F) row.names(Trans) <- Trans$V1 Trans$V1 <- as.character(Trans$V1) Trans$V2 <- as.character(Trans$V2) 
Trans[ is.na(Trans$V2), ]$V2 <- Trans[ is.na(Trans$V2), ]$V1 Trans[ Trans$V2 == "", ]$V2 <- Trans[ Trans$V2 == "", ]$V1 Table$Gene <- 
as.character(Table$Gene) Table$probeID <- as.character(Table$probeID) Table$Gene <- Trans[Table$probeID,]$V2
write.table(Table, file='Final_SMR_TWAS_Validation.tsv', row.names=F, col.names=T, quote=F, sep ='\t') ```

```{r synapse.parameters, include=FALSE, cache=TRUE} 
parentId = 'syn18936948'; activityName = 'Mendelian Randomization'; activityDescription = 'Validation of TWAS 
associations via SMR analysis'; thisFileName <- ''
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master') thisFile <- githubr::getPermlink(repository = thisRepo, 
repositoryPath=paste0('code/',thisFileName)) ```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE} activityName = 'Mendelian Randomization'; activityDescription = 'Validation of TWAS associations 
via SMR analysis'; CODE <- syn_temp$store(synapseclient$Folder(name = "Summary-data-based Mendelian Randomization", parentId = parentId))
# Set annotations
all.annotations = list(
  dataType = 'analysis',
  dataSubType = 'SMR_Validation_Tests',
  summaryLevel = 'gene',
  assay = 'TWAS',
  tissueTypeAbrv = c('NA'),
  study = c('ROSMAP','MSBB', 'Mayo')
  organism = 'HomoSapiens',
  consortium = c('ROSMAP','MSBB', 'Mayo') ,
  normalizationStatus	= TRUE,
  normalizationType	= 'NA',
  rnaquantification = 'NA',
  genomeAssemblyID = 'hg19' )
# Store SMR
ENRICH_OBJ <- syn_temp$store( synapseclient$File( path='Final_SMR_TWAS_Validation.tsv', name = 'SMR Values', parentId=CODE$properties$id ), activityName = 
activityName, executed = thisFile, activityDescription = activityDescription) all.annotations$dataSubType = 'Summary Stats' 
syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations) ```
### R Source Code
[Github](`r thisFile`) ```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE} reticulate::use_python("/usr/bin/python", required = TRUE) synapseclient <- 
reticulate::import("synapseclient") syn_temp <- synapseclient$Synapse() syn_temp$login() setwd("~/TWAS/code/") 
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R") source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R") 
createAndKnitToFolderEntityClient(file = "SMR.Rmd",
                                  parentId ="syn18936948",
                                  folderName = 'Summary-data-based Mendelian Randomization') 
```