### 2020-02-11 code/CMC_Validation.Rmd
CMC Validation

---
title: "CMC Validation"
output: html_document
---

Date of analysis update: "`r date()`"

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, cache=FALSE}
##It is assumed your working directory is "~/ampad-DiffExp/gene_level_analysis"
## Load required libraries
library(ggplot2)
library(stringr)

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")

options(xtable.type="html")
# Source modified utility functions from limma and sva package
source('~/TWAS/utilityFunctions/parallelDuplicateCorrelation.R')
source('~/TWAS/utilityFunctions/irwsva.build.R')
source('~/TWAS/utilityFunctions/f.pvalue.R')
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

## Validation

```{r Val}

################################################################################################################
#CMC DLPFC DATA
Impute <-  as.data.frame( read.table( syn_temp$get('syn21442621')$path, header = T, sep = '\t', row.names = 1 ))
row.names(Impute)  <- Impute$IID
Impute <- Impute[ , colnames(Impute)[ (grepl("ENSG", colnames(Impute))) == T ] ]
Impute <- as.data.frame(t(Impute))

#Exp
Exp <- as.data.frame( read.table( syn_temp$get('syn21470926')$path, header = T, sep = '\t', row.names = 1 ))
rownames(Exp) <- do.call( rbind, strsplit(row.names(Exp),'\\.'))[,1]
Exp_ACC <- Exp[ ,(grepl("_ACC_", colnames(Exp)))==T ]
Exp_PFC <- Exp[ ,(grepl("_PFC_", colnames(Exp)))==T ]

#ID--
ID <- read.csv( syn_temp$get('syn21444814')$path, header = T )
table(colnames(Exp) %in% ID$RNAseq.Sample_RNA_ID )
table(colnames(Exp_ACC) %in% ID$RNAseq.Sample_RNA_ID )
table(colnames(Exp_PFC) %in% ID$RNAseq.Sample_RNA_ID )

ID_GENO <- ID[( ID$Genotypes.Genotyping_Sample_ID) %in% gsub( '0_', '', colnames(Impute) )  == T , ]
dups <- ID[ ID$Individual_ID %in% names(table(ID_GENO$Individual_ID)[table(ID_GENO$Individual_ID)>1]), ]
ID_GENO <- ID_GENO[ (ID_GENO$RNAseq.Sample_RNA_ID %in% dups$RNAseq.Sample_RNA_ID[grepl("_BP_", dups$RNAseq.Sample_RNA_ID)]) == F, ]
row.names(ID_GENO) <- as.character( ID_GENO$Genotypes.Genotyping_Sample_ID )

#Replace columnn names
colnames(Impute) <- as.character( ID_GENO[ gsub( '0_', '', colnames(Impute) ), ]$Individual_ID )

ID_RNA <- ID[ ID$RNAseq.Sample_RNA_ID  %in% colnames(Exp_PFC), ]
row.names(ID_RNA) <- ID_RNA$RNAseq.Sample_RNA_ID
ID_RNA$Individual_ID <- as.character(ID_RNA$Individual_ID)
colnames(Exp_PFC) <- as.character( ID_RNA[ colnames(Exp_PFC), ]$Individual_ID)

FinNames <- colnames(Exp_PFC)[ colnames(Exp_PFC)  %in% colnames(Impute) ]
Impute_comp<-Impute[, FinNames ]
Exp_PFC<-Exp_PFC[, FinNames ]
table( colnames(Impute_comp) %in% colnames(Exp_PFC) )

#Filter Row names
keepers <- row.names(Exp_PFC)[ row.names(Exp_PFC) %in% row.names(Impute_comp) ]
Impute_comp <- Impute_comp[ keepers, ]
Exp_PFC <- Exp_PFC[ keepers, ]
table( row.names(Impute_comp) %in% row.names(Exp_PFC) )

Scaled_PFC_Exp <- t(scale(t(Exp_PFC)))

Correlation <- matrix(0, dim(Scaled_PFC_Exp)[1], 3)
row.names(Correlation) <- row.names(Impute_comp)
colnames(Correlation) <- c( "Z", "Correlation", "P-Val")

for( Gene in row.names(Impute_comp) ){
  foo <- cor.test( as.numeric(Impute_comp[Gene,]), as.numeric(Scaled_PFC_Exp[Gene,]) , method = "kendall")
  Correlation[ Gene, ] <- c( as.numeric(foo$statistic), as.numeric(foo$estimate), as.numeric(foo$p.value) )
}

Correlation <- cbind( Correlation, "Log10_Pval"= -log10(Correlation[,3]))
Correlation <- cbind( Correlation, "FDR"= p.adjust( Correlation[,3], method = "fdr", n = 13650) )
Correlation <- cbind( Correlation, "BH"= p.adjust( Correlation[,3], method = "BH", n = 13650) )
Correlation <- cbind( Correlation, "BF"= p.adjust( Correlation[,3], method = "bonferroni", n = 13650) )

Correlation <- cbind( Correlation, "FDR_W_Only"= p.adjust( Correlation[,3], method = "fdr", n = 6643) )
Correlation <- cbind( Correlation, "BH_W_Only"= p.adjust( Correlation[,3], method = "BH", n = 6643) )
Correlation <- cbind( Correlation, "BF_W_Only"= p.adjust( Correlation[,3], method = "bonferroni", n = 6643) )

Correlation <- as.data.frame(Correlation)
Correlation$Sig <- 'NO'
Correlation[ Correlation$FDR_W_Only < 0.05 & Correlation$Correlation > 0 , ]$Sig <- 'YES'

setEPS()
postscript("DLPFC_Val.eps", onefile = TRUE)
  ggplot( data = Correlation, aes( x = Correlation, y = Log10_Pval, col = Sig ) ) + 
    geom_point() + ggtitle("Log10( PVal) vs. Correlation" ) + 
    theme(plot.title = element_text(hjust = 0.5)) + ylab("-Log( PValue)") 
dev.off()
ggplot( data = Correlation, aes( x = Correlation, y = Log10_Pval, col = Sig ) ) + 
    geom_point() + ggtitle("Log10( PVal) vs. Correlation" ) + 
    theme(plot.title = element_text(hjust = 0.5)) + ylab("-Log( PValue)")

table( Correlation$Sig )
dim(Correlation[ Correlation$FDR_W_Only < 0.05, ] )
table( Correlation$`P-Val` < 0.05 )

VWs <- row.names( Correlation[ Correlation$FDR_W_Only < 0.05 & Correlation$Correlation > 0, ] )
length(VWs)/dim(Correlation)[1]
# 71.08234 % $

setEPS()
postscript("DLPFC_Cor_All.eps", onefile = TRUE)
  hist( Correlation$Correlation, breaks = 100, xlab = 'Correlation', las = 1,
      main="Correlation of All 6643 Matched Genes in CMC DLPFC")
dev.off()

hist( Correlation$Correlation, breaks = 100, xlab = 'Correlation', las = 1,
      main="Correlation of All 6643 Matched Genes in CMC DLPFC")

setEPS()
postscript("DLFPFC_Cor_gt5Perc.eps", onefile = TRUE)
  hist( Correlation[Correlation$FDR_W_Only < 0.05, ]$Correlation, 
      breaks = 100, 
      xlab = 'Correlation', 
      las = 1,
      main="Correlation of Genes with FDR Corrected P-Value < 0.05")
dev.off()
hist( Correlation[Correlation$FDR_W_Only < 0.05, ]$Correlation, 
      breaks = 100, 
      xlab = 'Correlation', 
      las = 1,
      main="Correlation of Genes with FDR Corrected P-Value < 0.05")
################################################################################
#CMC ACC DATA
library(stringr)

Impute <-  as.data.frame( read.table(syn_temp$get('syn21442621')$path, header = T, sep = '\t', row.names = 1 ))
row.names(Impute)  <- Impute$IID
Impute <- Impute[ , colnames(Impute)[ (grepl("ENSG", colnames(Impute))) == T ] ]
Impute <- as.data.frame(t(Impute))
colnames(Impute) <- gsub( "BP_", "", gsub("0_", "CMC_", colnames(Impute) ))

#Exp--
Exp <- as.data.frame( read.table( syn_temp$get('syn21483472')$path, header = T, sep = '\t', row.names = 1 ))
rownames(Exp) <- do.call( rbind, strsplit(row.names(Exp),'\\.'))[,1]
Exp_ACC <- Exp[ ,(grepl("_ACC_", colnames(Exp)))==T ]

#ID--
ID <- read.csv( syn_temp$get('syn21444814')$path, header = T )
table(colnames(Exp) %in% ID$RNAseq.Sample_RNA_ID )

colnames(Exp_ACC) <- paste0( "CMC_", gsub( "RNA_", "", gsub("_RNA_ACC_", "_", gsub( "ACC_BP_", "", colnames(Exp_ACC) ))))
Exp <- Exp_ACC
#Sync Names
Padder <- function( names ){
  #'@names format of 'CMC_PITT_XXXX' ie colnames(Exp)
  foo <- do.call( rbind, strsplit(names, '_'))
  foo[,3] <- str_pad(foo[,3], 5, pad = "0")
  return( paste0(foo[,1], "_", foo[,2], "_", foo[,3] ) )
}
colnames(Exp) <- Padder( colnames(Exp) )
colnames(Impute) <- Padder( colnames(Impute) )

#Align Matices
Impute <- Impute[ ,colnames(Impute)[ colnames(Impute) %in% colnames(Exp) ] ]
Exp <- Exp[ ,colnames(Impute)[ colnames(Impute) %in% colnames(Exp) ] ]
Impute <- Impute[ row.names(Impute)[ row.names(Impute) %in% row.names(Exp) ] ,]
Exp <- Exp[ row.names(Impute)[ row.names(Impute) %in% row.names(Exp) ] ,]

Correlation <- matrix(0, dim(Exp)[1], 3)
row.names(Correlation) <- row.names(Impute)
colnames(Correlation) <- c( "Z", "Correlation", "P-Val")

Exp <- Exp[ , colnames(Impute)[colnames(Impute) %in% colnames(Exp)] ]
Impute <- Impute[ , colnames(Impute)[colnames(Impute) %in% colnames(Exp)] ]

for( Gene in row.names(Impute) ){
  temp <- na.omit(cbind( as.numeric(Impute[Gene,]), as.numeric(Exp[Gene,]) ))
  foo <- cor.test( temp[,1], temp[,2] , method = "kendall")
  Correlation[ Gene, ] <- c( as.numeric(foo$statistic), as.numeric(foo$estimate), as.numeric(foo$p.value) )
}

Correlation <- cbind( Correlation, "Log10_Pval"= -log10(Correlation[,3]))
Correlation <- cbind( Correlation, "FDR"= p.adjust( Correlation[,3], method = "fdr", n = 13650) )
Correlation <- cbind( Correlation, "BH"= p.adjust( Correlation[,3], method = "BH", n = 13650) )
Correlation <- cbind( Correlation, "BF"= p.adjust( Correlation[,3], method = "bonferroni", n = 13650) )

Correlation <- cbind( Correlation, "FDR_W_Only"= p.adjust( Correlation[,3], method = "fdr", n = 6708) )
Correlation <- cbind( Correlation, "BH_W_Only"= p.adjust( Correlation[,3], method = "BH", n = 6708) )
Correlation <- cbind( Correlation, "BF_W_Only"= p.adjust( Correlation[,3], method = "bonferroni", n = 6708) )

Correlation <- as.data.frame(Correlation)
Correlation$Sig <- 'NO'
Correlation[ Correlation$FDR_W_Only < 0.05 & Correlation$Correlation > 0 , ]$Sig <- 'YES'
table(Correlation$Sig)

setEPS()
postscript("ACC_Val.eps", onefile = TRUE)
  ggplot( data = Correlation, aes( x = Correlation, y = Log10_Pval, col = Sig ) ) + 
    geom_point() + ggtitle("Log10( PVal) vs. Correlation" ) + 
    theme(plot.title = element_text(hjust = 0.5)) + ylab("-Log( PValue)") 
dev.off()
ggplot( data = Correlation, aes( x = Correlation, y = Log10_Pval, col = Sig ) ) + 
    geom_point() + ggtitle("Log10( PVal) vs. Correlation" ) + 
    theme(plot.title = element_text(hjust = 0.5)) + ylab("-Log( PValue)")

setEPS()
postscript("ACC_Cor_All.eps", onefile = TRUE)
  hist( Correlation$Correlation, breaks = 100, xlab = 'Correlation', las = 1,
      main="Correlation of All 6643 Matched Genes in CMC ACC")
dev.off()
hist( Correlation$Correlation, breaks = 100, xlab = 'Correlation', las = 1,
      main="Correlation of All 6643 Matched Genes in CMC ACC")

setEPS()
postscript("ACC_Cor_gt5Perc.eps", onefile = TRUE)
  hist( Correlation[Correlation$FDR_W_Only < 0.05, ]$Correlation, 
      breaks = 100, 
      xlab = 'Correlation', 
      las = 1,
      main="Correlation of Genes with FDR Corrected P-Value < 0.05")
dev.off()
hist( Correlation[Correlation$FDR_W_Only < 0.05, ]$Correlation, 
      breaks = 100, 
      xlab = 'Correlation', 
      las = 1,
      main="Correlation of Genes with FDR Corrected P-Value < 0.05")
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn21478397';
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'CMC Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with GRCh38 with CQN normalisation (ACC)';
thisFileName <- 'CMC_DataProcessing.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
activityName = 'ACC Validation';
activityDescription = 'CMC Validation of Imputed Expression';
CODE <- syn_temp$store(synapseclient$Folder(name = "ACC Validation", parentId = parentId))

#Set Used SynIDs For Provenance
#Syns_Used <- c("")

# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'TWAS',
  summaryLevel = 'Correlation',
  assay	 = 'TWAS imputation',
  tissueTypeAbrv	= NULL, 
  study = 'CMC', 
  organism = 'HomoSapiens',
  consortium	= 'CMC',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh37'
)

# Store results
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='DLPFC_Val.eps', name = 'DLPFC Correlation vs PValue', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='DLPFC_Cor_All.eps', name = 'DLPFC Correlation Distributions', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='DLPFC_Cor_gt5Perc.eps', name = 'DLPFC Sig Correlation Distributions', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='DLPFC_Val.eps', name = 'DLPFC Correlation vs PValue', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Cor_All.eps', name = 'ACC Correlation Distributions', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Cor_gt5Perc.eps', name = 'ACC Sig Correlation Distributions', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

```

### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")
createAndKnitToFolderEntityClient(file = "CMC_Validation.Rmd",
                                          parentId ="syn18936948",
                                          folderName = 'ACC Validation')
```