### 2020-05-14 code/GeneSet_Expansion.Rmd
ReDo GeneSet enrichment

---
title: "GeneSet_Expansion"
output: html_document
---

```{r setup, include=FALSE}
# Source modified utility functions from limma and sva package
source("~/TWAS/code/utilityFunctions/Parallel_vbsrBootstrap.R")
library(parallel)
library(doParallel)
library(spike)
library(reshape2)
library(biomaRt)
library(data.table)

source('~/TWAS/utilityFunctions/parallelDuplicateCorrelation.R')
source('~/TWAS/utilityFunctions/irwsva.build.R')
source('~/TWAS/utilityFunctions/f.pvalue.R')
library(biomaRt)

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

## Load in Regions and build genesets from BioMart
```{r Load}
Regions <- read.table( file='../Reference_Files/Gene_Hit_Regions.bed', header=F, sep='\t')

row.names(Regions) <- Regions$V4

Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V2 <- Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V2+500000
Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V3<-Regions[ c("TREM2", "CD2AP", "MTCH2", "EED", "APOC1","CEACAM19","CLPTM1"), ]$V3-500000

mart <- useMart("ensembl", host="grch37.ensembl.org", dataset = "hsapiens_gene_ensembl")
#mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "jul2019.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
#mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl")

#row.names(Regions) <- Regions$V4

Candidates <- list()
for( name in as.character(Regions$V4) ){

  filterlist <- paste0( Regions[ name, ]$V1, ':', Regions[ name, ]$V2, ':', Regions[ name, ]$V3 )

  ## Query biomart
  eval(parse(text= paste0( 'Candidates$', name, ' <- getBM(attributes = c(\'ensembl_gene_id\', 
                                        \'hgnc_symbol\', 
                                        \'chromosome_name\', 
                                        \'start_position\', 
                                        \'end_position\'),
                        filters = c(\'chromosomal_region\'), 
                        values = filterlist,
                        mart = mart)'
  )))
}

for( name in as.character(Regions$V4) ){
  eval(parse(text= paste0( 'Candidates$', name, ' <- Candidates$', name, '[ (Candidates$', name, '$hgnc_symbol == \'\') == F, ]' )))
}
```

## Expand Target List
```{r Expand, echo=FALSE}

reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

#SynIDs of expression data to pull from
ExpressionDS <- c('syn21292041','syn21285564','syn21285564','syn21285564','syn21285564','syn21291908')
names( ExpressionDS ) <- c( 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')

#Study ID Translator
Study <- c( 'RosMap', 'Mayo', 'MSBB', 'MSBB', 'MSBB', 'MSBB')
names(Study) <- c('DLPFC', 'TCX', 'FP', 'IFG', 'PHG', 'STG')

Syn <- list('DLPFC', 'TCX', 'FP', 'IFG', 'PHG', 'STG')

Partials <- list()

library(parallel)
library(doParallel)
#detach("package:doParallel", unload=TRUE)
#detach("package:parallel", unload=TRUE)
cores <- detectCores()-2 
cl <- makePSOCKcluster(cores)
registerDoParallel(cl)

mark <- Sys.time()
for( name in as.character(Regions$V4) ){
 #for( Tissue in names(ExpressionDS)){
 for( Tissue in names(ExpressionDS) ){
  message(paste0("Working on: ", Tissue))
  #Tissue<-'DLPFC'

  #Load expression for tissue
  exp <- read.table(syn_temp$get(as.character(ExpressionDS[Tissue]))$path, header =T, sep ='\t', row.names=1)
  colnames( exp ) <- gsub( "X", "", colnames( exp ) )
  #Seperate exp by tissue
  if( Tissue == 'DLPFC'){
  }else{
    if( Tissue == 'TCX' | Tissue == 'CBE' ){
      if( Tissue == 'CBE' ){
        slec <- 'CER'
      }else{ slec <- 'TCX' }
      colnames( exp ) <- gsub( "TC", "TCX", colnames( exp ) )
      exp <- exp[ , grepl( slec, colnames(exp)) ]
    }else{
      if( Tissue %in% c('FP', 'IFG', 'PHG', 'STG') ){
        Meta <- read.table( syn_temp$get('syn21285520')$path, header =T, sep ='\t', row.names=1 )
        exp <- exp[ colnames(exp) %in% row.names(Meta[grepl( Tissue, Meta$Tissue.Diagnosis),])]
      }else{
        stop(paste0("ERROR: SOURCE=Config.yaml Issue=Tissue: ", Tissue," is improper must be one of: CBE, DLPFC, FP, IFG, PHG, STG, TCX"))
      }
    }
  }

  #Impute svalues for given gene-patient NA values
  exp <- exp[rowSums(is.na(exp))<ncol(exp), ]

  foo <- bcv::impute.svd( t(exp) )
  Exp <- foo$x
  row.names(Exp) <- row.names(t(exp)) 
  colnames(Exp) <- colnames(t(exp))

  #Prep for partial correlation detection
  Final <- data.frame()

  source("~/TWAS/code/utilityFunctions/Parallel_vbsrBootstrap.R")
  #Run Partial correlations for each gene

  RUNNe <- function( i=i, x=Exp ){ 
    library(parallel)
    library(doParallel)
    source("~/TWAS/code/utilityFunctions/Parallel_vbsrBootstrap.R")
    OBS <- i
    y <- as.matrix(x[,OBS]) 
    colnames(y) <- i
    X <- x[,(colnames(x) %in% OBS) == F ]
    att <- pvbsrBootstrap( y=y, x=X, nsamp=100, cores=1 )
    names( att ) <- gsub( "intercept", OBS, names(att) )
    att <- c( 'SeedGene' = OBS, att[ colnames(x) ]) 
    return( att[c( 'SeedGene', colnames(x) )] )
    #return(eval(parse( text = paste( c( OBS,att[colnames(x)]), sep='\t') )))
  }

  LIST <- eval( parse( text= paste0( 'Candidates$',name,'$ensembl_gene_id' )))  
  LIST  <- LIST[ (LIST %in% colnames(Exp) ) ==T ]

  mark <- Sys.time()
    foo <- t( parApply(cl, as.matrix( LIST[1:length(LIST)] ), 1, RUNNe, Exp) )
  message( paste0( Sys.time()-mark ))

  eval(parse(text=paste0( 'Partials$', name, '$', Tissue, '<- foo' )))

  message(paste0( "Finished Tissue: ", Tissue, ' For Gene: ', name ))
 }
  message(paste0( "Completed All Tissues Finished Seed Gene: ", name ))
}
mark-Sys.time()
#sink<-Partials

```

```{r Selected, echo=FALSE}
#Average the partial correlation statistics across the data  frames

GENEs <- NULL
for( tissue in names(Partials$TREM2) ){
  GENEs <- eval(parse( text=paste0('c(GENEs, colnames(Partials$TREM2$',tissue,'))') ))
  GENEs <- GENEs[ !duplicated(GENEs) ]
}

GENEs <- GENEs[ GENEs != 'SeedGene' ]
ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = 'ensembl.org')
ensembl <- useDataset('hsapiens_gene_ensembl', mart = ensembl)

Top50_Cor <- function( ENSG, Targets, Tissues, Part, ensembl ){
  #'@ENSG the target gene ID to collect partials on eg. 'ENSG00000095970'
  #'@Targets a vector of ENSG names to scan eg. GENEs
  #'@Tissues charater vector of tissues to scan eg. c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  )
  #'@Part a partial correlation list  eg. Partials$TREM2
  #'@ensembl ensembl object eg. ensembl

  Examine <- matrix( NA, length(Targets), length(Tissues)+1 )
  Examine <- apply(Examine, 2, as.numeric)
  row.names(Examine) <- Targets
  colnames(Examine) <- c( Tissues, "Avg" )

  for( Tis in Tissues ){
    eval(parse( text= paste0( 'Examine[ colnames(Part$\'', Tis,'\')[ colnames(Part$\'', Tis,'\') !=\'SeedGene\' ],\'', Tis, '\' ]  <- as.numeric( Part$\'', Tis, '\'[ Part$\'', Tis,'\'[ ,\'SeedGene\'] == \'', ENSG, '\' , colnames(Part$\'', Tis,'\') != \'SeedGene\' ])' ) ))
  }

  Examine[, 'Avg' ] <- apply( Examine[ , 1:(dim(Examine)[2]-1) ], 1, mean )
  Examine <- Examine[ order(-Examine[,'Avg']), ]

  #message(paste0("1"))  
  #ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = 'ensembl.org')
  #message(paste0("2"))
  #ensembl <- useDataset('hsapiens_gene_ensembl', mart = ensembl)
  #message(paste0("3"))
  gene_ids <- row.names( Examine )[1:50]

  attrs <- c("ensembl_gene_id", "hgnc_symbol")
  Trans <- getBM(filters = 'ensembl_gene_id',
                            attributes = attrs,
                            values = gene_ids,
                            mart = ensembl )

  return( Trans$hgnc_symbol[ Trans$hgnc_symbol != "" ] )
}

Top50 <- list()
Top50$Trem <- Top50_Cor( 'ENSG00000095970', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$TREM2, ensembl )
Top50$CD2AP <- Top50_Cor( 'ENSG00000198087', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$CD2AP, ensembl )
Top50$MTCH2 <- Top50_Cor( 'ENSG00000109919', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$MTCH2, ensembl )
Top50$EED <- Top50_Cor( 'ENSG00000074266', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$EED, ensembl )
Top50$KNOP1 <- Top50_Cor( 'ENSG00000103550', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$KNOP1, ensembl )
Top50$APOC1 <- Top50_Cor( 'ENSG00000130208', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$APOC1, ensembl )
Top50$CEACAM19 <- Top50_Cor( 'ENSG00000186567', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$CEACAM19 , ensembl )
Top50$CLPTM1 <- Top50_Cor( 'ENSG00000104853', GENEs, c( "FP", "TCX", "DLPFC", "IFG", "PHG", "STG"  ), Partials$CLPTM1, ensembl )
```

```{r SIF_Load, echo=FALSE}
reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login( )

SIFs <- c('syn21914055')
names(SIFs) <- c('All')

#Load and Add names: SIFs
boot <- function( i ){
  foo <- read.table( file=syn_temp$get( SIFs[i] )$path , header = F, sep='\t' )
  foo$Pathway <- as.character( i ) 
  return(foo)
}

Loader <- function( SIFs){
  #'@SIFs An object that has SynID values and named for the database
  ##  Each Syn ID is a SIF format file

  total <- list()
  Trial <- for( i in names(SIFs) ){
    temp <- boot( i )
    #eval(parse(text=paste0( 'total$', as.character(NAMES[i]), ' <- temp' )))
    total[[i]] <- temp
  }

  Total <- do.call(rbind, total)
  Total <- Total[ , c("V1", "V3", "V2", "Pathway") ]
  colnames(Total) <- c("from", "to", "interaction", "pathway")

  #Caluclate how many times this intereactcion was found in all databases:
  Total$Occurance <- paste0( Total$from, '-', Total$to, ':', Total$interaction )
  foo <- paste0( Total$from, '-', Total$to, ':', Total$interaction )
  TabFoo <- table(foo)
  Total$Occurance <-  as.numeric( TabFoo[ Total$Occurance ] )

  Genes <- c( as.character(Total$from), as.character(Total$to) )
  Genes <- Genes[ !duplicated(Genes) ]
  Genes <- as.data.frame( Genes )

  #Make the pathway column into a list object
  #library(dplyr)
  #library(reshape)

  Total$UniqCol <- paste0( as.character(Total$from),':', as.character(Total$to),'-', as.character(Total$interaction) )
  dt <- data.table(Total[, c('UniqCol','pathway')])
  DT <- dt[,lapply(.SD, function(col) paste(col, collapse=", ")), by=.(UniqCol)]

  sinl<-DT

  foo <- as.data.frame( DT )
  poo <- as.list( strsplit(as.character(foo$pathway),',') )
  names(poo) <- foo$UniqCol

  totals <- Total[ !duplicated(Total$UniqCol), ]
  pathways <- poo

  table(names(pathways) == as.character(totals$UniqCol))
  table( as.character(totals$UniqCol) == names(pathways) )

  totals$PATH <- pathways
  totals$PATH <- lapply( totals$PATH,
                 function(x) gsub(" ","", x)
               )

  Total <- totals[,c("from", "to", "interaction", "Occurance", "UniqCol", "PATH")]
  colnames(Total) <- c("from", "to", "interaction", "Occurance", "UniqCol", "pathway")
  return( Total )
}

Total <- Loader(SIFs)
Total$from <- as.character(Total$from)
Total$to <- as.character(Total$to)
Total$UniqCol <- as.character(Total$UniqCol)
dim(Total)
table(table( Total$UniqCol ))
#Filter out the :CHEBI
Total <- Total[ (grepl( ':CHEBI', Total$UniqCol) == F) , ]

#Remove Repeat interactions?
Total$Partners <- paste0(Total$from, ":", Total$to )
Total <- Total[ !duplicated(Total$Partners), ]

```

```{r SIF_Expand, echo=FALSE}
Expand_gt1 <- list()
Expand_gt2 <- list()

#Convert Brain Expressed to Gene names
ensembl = useEnsembl(biomart = "ensembl", dataset = "hsapiens_gene_ensembl" )
GENEsyb <- getBM(attributes = c('hgnc_symbol'), 
      filters = 'ensembl_gene_id', 
      values = GENEs, 
      mart = ensembl)

for( nam in names(Top50) ){
  genes <- eval(parse(text=paste0( 'Top50$', nam )))
  Exp <- c( Total[ Total$from %in% genes | Total$to %in% genes, ]$from, 
            Total[ Total$from %in% genes | Total$to %in% genes, ]$to
          )

  #Filt out molecule intereactions and split into conservative and non conservative
  gt1 <- names( table(Exp)[ table(Exp) > 1 ] )
  gt1 <- gt1[ grepl( 'CHEBI:', gt1) == F ]
  gt2 <- names( table(Exp)[ table(Exp) > 2 ] )
  gt2 <- gt2[ grepl( 'CHEBI:', gt2) == F ]

  #Filt for brain expressed Genese
  eval(parse(text=paste0( 'Expand_gt1$', nam, ' = gt1[ gt1 %in% GENEsyb[,\'hgnc_symbol\'] ] ' )))
  eval(parse(text=paste0( 'Expand_gt2$', nam, ' = gt2[ gt2 %in% GENEsyb[,\'hgnc_symbol\'] ] ' )))
}

library( clusterProfiler )
library(enrichplot)
library(enrichR)
library(org.Hs.eg.db)

parentId = 'syn18936948';
activityName = 'Gene Set Enrichment';
activityDescription = 'Figure 4 & Supplement: gene set enrichment of genes implicated in TWAS regions';
thisFileName <- 'GeneSet_Expansion.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))

CODE <- syn_temp$store(synapseclient$Folder(name = "Gene Set Enrichment Plots", parentId = parentId))

for( name in names(Expand_gt2)){
  eval(parse(text=paste0('set <- Expand_gt2$', name )))
  ego <- enrichGO(gene            = set,
                    universe      = GENEsyb[,'hgnc_symbol'],
                    OrgDb         = org.Hs.eg.db,
                    keyType      = 'SYMBOL',
                    ont           = "ALL",
                    pAdjustMethod = "BH",
                    pvalueCutoff  = 0.01,
                    qvalueCutoff  = 0.05
                  )

  clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30)
  cnetplot(ego, foldChange=NULL)

  Dot <- clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30)
  Net <- cnetplot(ego, foldChange=NULL)

  pdf(file = paste0("EnrichmentDotPlot_", name, "_GeneSetSize_", length(set),"_gt2.pdf"), width = 14, height = 7)
    print( clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30) )
  dev.off()
  ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentDotPlot_", name, "_GeneSetSize_", length(set),"_gt2.pdf"), name = paste0("EnrichmentDotPlot_", name, "_GeneSetSize_", length(set),"_Gene apperars gt 2 times"), parentId=CODE$properties$id ), activityName = activityName, executed = as.character(c(ExpressionDS,thisFile)), activityDescription = activityDescription)

  pdf(file = paste0("EnrichmentNetPlot_", name, "_GeneSetSize_", length(set),"_gt2.pdf"), width = 14, height = 7)
    print( cnetplot(ego, foldChange=NULL) )
  dev.off()
  ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentNetPlot_", name, "_GeneSetSize_", length(set),"_gt2.pdf"), name = paste0("EnrichmentNetPlot_", name, "_GeneSetSize_", length(set),"_Gene apperars gt 2 times"), parentId=CODE$properties$id ), activityName = activityName, executed = as.character(c(ExpressionDS,thisFile)), activityDescription = activityDescription)
}

for( name in names(Expand_gt1)){
  eval(parse(text=paste0('set <- Expand_gt1$', name )))
  ego <- enrichGO(gene            = set,
                    universe      = GENEsyb[,'hgnc_symbol'],
                    OrgDb         = org.Hs.eg.db,
                    keyType      = 'SYMBOL',
                    ont           = "ALL",
                    pAdjustMethod = "BH",
                    pvalueCutoff  = 0.01,
                    qvalueCutoff  = 0.05
                  )

  clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30)
  cnetplot(ego, foldChange=NULL)

  Dot <- clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30)
  Net <- cnetplot(ego, foldChange=NULL)

  pdf(file = paste0("EnrichmentDotPlot_", name, "_GeneSetSize_", length(set),"_gt1.pdf"), width = 14, height = 7)
    print( clusterProfiler::dotplot(ego, x = "GeneRatio", showCategory=30) )
  dev.off()
  ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentDotPlot_", name, "_GeneSetSize_", length(set),"_gt1.pdf"), name = paste0("EnrichmentDotPlot_", name, "_GeneSetSize_", length(set),"_Gene apperars gt 1 times"), parentId=CODE$properties$id ), activityName = activityName, executed = as.character(c(ExpressionDS,thisFile)), activityDescription = activityDescription)

  pdf(file = paste0("EnrichmentNetPlot_", name, "_GeneSetSize_", length(set),"_gt1.pdf"), width = 14, height = 7)
   print( cnetplot(ego, foldChange=NULL) )
  dev.off()
  ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0("EnrichmentNetPlot_", name, "_GeneSetSize_", length(set),"_gt1.pdf"), name = paste0("EnrichmentNetPlot_", name, "_GeneSetSize_", length(set),"_Gene apperars gt 1 times"), parentId=CODE$properties$id ), activityName = activityName, executed = as.character(c(ExpressionDS,thisFile)), activityDescription = activityDescription)

}

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.